### 2022-06-08 Code/PaperMelanoma.Rmd
Initial

---
title: "PaperMelanoma"
author: "Carolien C.H.M. Maas"
date: "3-6-2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Development and validation of a clinial prediction model to predict recurrence and disease-specific mortality in patients with melanoma after sentinel lymph node biopsy

This file includes the code and results for the above mentioned paper.

# Data preparation
### Load packages, functions, and data 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# load libraries
library(foreign)   # Reading external (spps) data
library(mice)      # Multipe imputation
library(survminer) # Kaplan-Meier plot
library(prodlim)   # median follow-up time
library(glmnet)    # LASSO
library(rms)       # rcs
library(Hmisc)     # rcorr.cens and fit.mult.impute
library(mitools)   # MIcombine
library(arsenal)   # descriptives table
library(png)       # reading PNG images
set.seed(100)
# import function to create table of missing values
source("Z:/Project Melanoom/PaperMelanoma/Code/table.missing.values.R") 
# import function to plot calibration
source("Z:/Project Melanoom/PaperMelanoma/Code/val_surv_mi_v3.R")
# helper functions for calibration plots
km <- function(sel, S.km, horizon){
  S.km.sel <- S.km[sel,]
  sf <- survfit(S.km.sel~1)
  year1 <- max(sf$time[sf$time<=horizon])
  1-sf$surv[sf$time==year1]
}
km.lower <- function(sel, S.km, horizon){
  S.km.sel <- S.km[sel,]
  sf <- survfit(S.km.sel~1)
  year1 <- max(sf$time[sf$time<=horizon])
  1-sf$lower[sf$time==year1]
}
km.upper <- function(sel, S.km, horizon){
  S.km.sel <- S.km[sel,]
  sf <- survfit(S.km.sel~1)
  year1 <- max(sf$time[sf$time<=horizon])
  1-sf$upper[sf$time==year1]
}
fun.event <- function(lp, h0)
{
  h <- h0*exp(lp)
  p <- 1-exp(-h)
  return(p)
}

# load data file
original.data <- foreign::read.spss(file="Z:/Project Melanoom/PaperMelanoma/Data/Dataset Predictiemodel Stassen.sav",
                                    to.data.frame=TRUE, origin="01-01-1970")

# Select necessary variables
selected.vars <- c("EXCLUSIE", "DatabaseNR", "Center", "Sex", "Age", "DOB",
                   "Ulceration", "Loc_CAT",            # OLD data set: Loc_CAT2
                   "Histology", "Breslow", "Tot_nr_SNs",
                   "nr_SNs_Field_1", "nr_SNs_Field_2", "nr_SNs_Field_3",
                   "Mitosis",                          # needed for AJCC.7.i
                   "SNdate", "Recurrence", "Date_1st_Rec", "Last_FU",
                   "dead", "Status",
                   "SNstatus",                         # distinguish between positive and negative SN patients
                   "Tot_SNs_pos", "Rdamcrit", "Dewar") # additional for positive SN patients
```

### Exclude patients
Exclude when Exclusion = 2: based on duplicates, urogenital melanoma, etc.
Leave the patients included that have EXCLUSIE = 1 and Dbstatus = 8,
even though diagnosis and operation date are very close.
```{r, eval=TRUE}
sel.data <- original.data[original.data$EXCLUSIE!="Exclusion", selected.vars]
cat("Number of patients excluded:", nrow(original.data)-nrow(sel.data), "\n")
```
Omit patient A0523 - no last follow-up date
```{r, eval=TRUE}
sel.data <- sel.data[sel.data$DatabaseNR!="A0523               ",]
```

### Recode
Set Rdamcrit of negative SN patients equal to 0
```{r, eval=TRUE}
sel.data$Rdamcrit[sel.data$SNstatus=="Negative"] <- 1 # since log(1) = 0
```

Merge single cells with subcap and set Dewar for negative patients to subcap
```{r, eval=TRUE}
sel.data$Dewar[sel.data$Dewar=="single cells"] <- "subcap"
sel.data$Dewar[sel.data$SNstatus=="Negative"] <- "subcap"
```

Merge location other with trunk
```{r, eval=TRUE}
sel.data$Loc_CAT[sel.data$Loc_CAT=="other"] <- "trunk"
```

Merge LMM other with other
```{r, eval=TRUE}
sel.data$Histology[sel.data$Histology=="LMM"] <- "other"
```

Melanoma specific mortality (MSM)
```{r, eval=TRUE}
sel.data$MSM<-1*(sel.data$Status=="DOD")
```

Define multiple fields
```{r, eval=TRUE}
sel.data$multiple.fields <- ifelse(!is.na(sel.data$nr_SNs_Field_2)&sel.data$nr_SNs_Field_2>0,1,0)
```

Create number of negative nodes removed
```{r, eval=TRUE}
sel.data$Tot_SNs_neg <- sel.data$Tot_nr_SNs - sel.data$Tot_SNs_pos
```

Use age at SN instead of age at diagnosis
```{r, eval=TRUE}
sel.data$Age.SN <- floor((sel.data$SNdate-sel.data$DOB)/(24*60*60)/365.25)
```

Define time since diagnosis in years (follow-up time)
```{r, eval=TRUE}
sel.data$FU.time <- (sel.data$Last_FU-sel.data$SNdate)/(24*60*60)/365.25
```

Define recurrence time in years
```{r, eval=TRUE}
sel.data$Rec.time<- (sel.data$Date_1st_Rec-sel.data$SNdate)/(24*60*60)/365.25
```

Omit patients with Rec date before SN date
```{r, eval=TRUE}
old.n <- nrow(sel.data)
sel.data <- sel.data[!(sel.data$Rec.time<0&!is.na(sel.data$Rec.time)),]
cat("Number of patients excluded:", old.n-nrow(sel.data), "\n")
```

Set recurrence time equal to follow-up time for those without recurrence
```{r, eval=TRUE}
sel.data$Rec.time[sel.data$Recurrence=="no recurrence"|is.na(sel.data$SNdate)]<-sel.data$FU.time[sel.data$Recurrence=="no recurrence"|is.na(sel.data$SNdate)]
```

Set Mitosis to Unknown if missing
```{r, eval=TRUE}
sel.data$Mitosis[is.na(sel.data$Mitosis)] <- "Unknown" # mitosis unknown for 80% of patients
```

Replace "Unknown" by NA
```{r, eval=TRUE}
sel.data$Ulceration[sel.data$Ulceration=="unknown"]<-NA
sel.data$Loc_CAT[sel.data$Loc_CAT=="unknown"]<-NA
sel.data$Histology[sel.data$Histology=="unknown"]<-NA
sel.data$Dewar[sel.data$Dewar=="unknown"] <- NA
sel.data$Mitosis[sel.data$Mitosis=="Unknown"] <- NA
```

Delete unused levels
```{r, eval=TRUE}
sel.data <- droplevels(sel.data)
```

# Multiple imputation
```{r, eval=TRUE}
variables.to.be.imputed <- c("Center", "SNstatus", "Sex", "Age.SN", "Ulceration",
                             "Loc_CAT", "Histology", "Breslow", "multiple.fields",
                             "Tot_SNs_neg", "Tot_SNs_pos", "Rdamcrit", "Dewar",
                             "Mitosis",
                             "Recurrence", "dead", "MSM")
data.to.be.imputed <- sel.data[,variables.to.be.imputed]
```
Save distributions of all variables, necessary for Harrell summaries
```{r, eval=TRUE}
dd<-datadist(data.to.be.imputed)
options(datadist='dd')
options(digits=8)
```

Relation between outcome and covariates
```{r, eval=TRUE}
data.to.be.imputed$logRec.time <- log(sel.data$Rec.time+1/365.25)
data.to.be.imputed$logFU.time <- log(sel.data$FU.time+1/365.25)
missings.table(data.to.be.imputed)
save(data.to.be.imputed,file='Z:/Project Melanoom/PaperMelanoma/Data/descr.data.Rdata', compress=TRUE)
```

Number of imputations
```{r, eval=TRUE}
m <- 5
```

Perform multiple imputation
```{r, eval=FALSE}
imputed.data <- mice::mice(data.to.be.imputed, m=m, maxit=5, print=FALSE, seed=500)
print(imputed.data$loggedEvents)
save(imputed.data,file='Z:/Project Melanoom/PaperMelanoma/Data/imputed.data.RData', compress=TRUE)
```

Load imputed data set
```{r, eval=TRUE}
load('Z:/Project Melanoom/PaperMelanoma/Data/imputed.data.RData')
complete.imputed.data <- mice::complete(imputed.data, m)
summary(complete.imputed.data)
missings.table(complete.imputed.data)
```

# Set as survival data
```{r, eval=TRUE}
S.Rec <- survival::Surv(exp(complete.imputed.data$logRec.time), sel.data$Recurrence=="recurred")
S.MSM <- survival::Surv(exp(complete.imputed.data$logFU.time), sel.data$MSM)
```

### Kaplan-Meier plots
```{r, eval=TRUE}
center.Rec.KM <- survival::survfit(S.Rec ~ Center, data=sel.data)
survminer::ggsurvplot(center.Rec.KM, conf.int=TRUE, risk.table=TRUE)
center.MSM.KM <- survival::survfit(S.MSM ~ Center, data=sel.data)
survminer::ggsurvplot(center.MSM.KM, conf.int=TRUE, risk.table=TRUE)
```

# Median follow-up time
```{r, eval=TRUE}
median.FU.time.total <- stats::quantile(prodlim::prodlim(Hist(exp(logFU.time), MSM)~1, data=complete.imputed.data, reverse=TRUE))
tab.FU.time <- c("total", paste(sprintf("%.2f", median.FU.time.total$quantiles.survival[3, "quantile"]),
                 paste0("[", sprintf("%.2f", median.FU.time.total$quantiles.survival[3, "lower"]), "; ",
                        sprintf("%.2f", median.FU.time.total$quantiles.survival[3, "upper"]), "]")))
median.FU.time <- stats::quantile(prodlim::prodlim(Hist(exp(logFU.time), MSM)~Center, data=complete.imputed.data, reverse=TRUE))
for (center in sort(as.character(unique(sel.data$Center)))){
  med.FU.table <- eval(parse(text=paste0("median.FU.time$`Center=", center, "`")))
  med.FU <- med.FU.table[3,"quantile"]
  lower.FU <- med.FU.table[3,"lower"]
  upper.FU <- med.FU.table[3,"upper"]
  tab.FU.time <- rbind(tab.FU.time, c(center, paste(sprintf("%.2f", med.FU), paste0("[", sprintf("%.2f", lower.FU), "; ", sprintf("%.2f", upper.FU), "]"))))
}
write.table(as.data.frame(tab.FU.time),file='Z:/Project Melanoom/PaperMelanoma/Results/med.FU.table.txt',
            row.names=FALSE)
tab.FU.time
```

# Recurrence at 5 years
```{r, eval=TRUE}
Rec.table <- c("Total", sum(sel.data$Recurrence=="recurred"), sprintf("%.1f", sum(sel.data$Recurrence=="recurred")/nrow(sel.data)*100))
Rec.table <- rbind(Rec.table, c("Positive", sum(sel.data$Recurrence=="recurred"&sel.data$SNstatus=="Positive"),
                                sprintf("%.1f", sum(sel.data$Recurrence=="recurred"&sel.data$SNstatus=="Positive")/sum(sel.data$SNstatus=="Positive")*100)))
Rec.table <- rbind(Rec.table, c("Negative", sum(sel.data$Recurrence=="recurred"&sel.data$SNstatus=="Negative"),
                                sprintf("%.1f", sum(sel.data$Recurrence=="recurred"&sel.data$SNstatus=="Negative")/sum(sel.data$SNstatus=="Negative")*100)))
utils::write.table(as.data.frame(Rec.table),file='Z:/Project Melanoom/PaperMelanoma/Results/Rec.table.txt',
                   row.names=FALSE)
Rec.table
```

# Melanoma Specific Mortality at 5 years
```{r, eval=TRUE}
MSM.table <- c("Total", sum(sel.data$MSM==1), sprintf("%.1f", sum(sel.data$MSM==1)/nrow(sel.data)*100))
MSM.table <- rbind(MSM.table, c("Positive", sum(sel.data$MSM==1&sel.data$SNstatus=="Positive"),
                                sprintf("%.1f", sum(sel.data$MSM==1&sel.data$SNstatus=="Positive")/sum(sel.data$SNstatus=="Positive")*100)))
MSM.table <- rbind(MSM.table, c("Negative", sum(sel.data$MSM==1&sel.data$SNstatus=="Negative"),
                                sprintf("%.1f", sum(sel.data$MSM==1&sel.data$SNstatus=="Negative")/sum(sel.data$SNstatus=="Negative")*100)))
utils::write.table(as.data.frame(MSM.table),file='Z:/Project Melanoom/PaperMelanoma/Results/MSM.table.txt',
                   row.names=FALSE)
MSM.table
```

### Restrict survival at 5 years
```{r, eval=TRUE}
S.Rec[S.Rec[,1]>5, 1] <- 5 # set recurrence time to 5
S.Rec[S.Rec[,1]>5, 2] <- 0 # set censoring
S.MSM[S.MSM[,1]>5, 1] <- 5 # set follow-up time to 5
S.MSM[S.MSM[,1]>5, 2] <- 0 # set censoring
```

# Obtain C-index of AJCC7 and AJCC8
```{r, eval=TRUE}
TNM.7 <- c()
AJCC.7 <- c()
cindex.AJCC.7 <- rep(0, m)
cindex.AJCC.7.se <- rep(0, m)
cindex.AJCC.7.pos <- rep(0, m)
cindex.AJCC.7.se.pos <- rep(0, m)
cindex.AJCC.7.neg <- rep(0, m)
cindex.AJCC.7.se.neg <- rep(0, m)
for (i in 1:m){
  imputation.i <- mice::complete(imputed.data,i)
  TNM.7.i <- rep("", nrow(imputation.i))
  AJCC.7.i <- rep(0, nrow(imputation.i))
  for (patient in 1:nrow(imputation.i)){
    breslow.patient <- imputation.i[patient, "Breslow"]
    ulceration.patient <- as.character(imputation.i[patient, "Ulceration"])
    mitosis.patient <- as.character(imputation.i[patient, "Mitosis"])
    nodes.patient <- imputation.i[patient, "Tot_SNs_pos"]

    snstatus.patient <- imputation.i[patient, "SNstatus"]

    # T stage
    if (breslow.patient <= 1){
      if (ulceration.patient=="No" & mitosis.patient=="No"){
        TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T1a")
      } else if (ulceration.patient=="Yes" | mitosis.patient=="Yes"){
        TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T1b")
      }
    } else if (breslow.patient>1 & breslow.patient <= 2){
      if (ulceration.patient=="No"){
        TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T2a")
      } else if (ulceration.patient=="Yes"){
        TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T2b")
      }
    } else if (breslow.patient>2 & breslow.patient <= 4){
      if (ulceration.patient=="No"){
        TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T3a")
      } else if (ulceration.patient=="Yes"){
        TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T3b")
      }
    } else if (breslow.patient>4){
      if (ulceration.patient=="No"){
        TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T4a")
      } else if (ulceration.patient=="Yes"){
        TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T4b")
      }
    }

    # N stage
    if (nodes.patient == 0){
      TNM.7.i[patient] <- paste0(TNM.7.i[patient], "N0")
    } else if (nodes.patient==1){
      TNM.7.i[patient] <- paste0(TNM.7.i[patient], "N1a")
    } else if (nodes.patient==2 | nodes.patient==3){
      TNM.7.i[patient] <- paste0(TNM.7.i[patient], "N2a")
    } else if (nodes.patient>3){
      TNM.7.i[patient] <- paste0(TNM.7.i[patient], "N3a")
    }

    # M stage for all patients in our data set
    TNM.7.i[patient] <- paste0(TNM.7.i[patient], "M0")

    # determine AJCC.7.i
    if (TNM.7.i[patient]=="T1aN0M0"){
      AJCC.7.i[patient] <- 1.1
    } else if (TNM.7.i[patient]=="T1bN0M0" | TNM.7.i[patient]=="T2aN0M0"){
      AJCC.7.i[patient] <- 1.2
    } else if (TNM.7.i[patient]=="T2bN0M0" | TNM.7.i[patient]=="T3aN0M0"){
      AJCC.7.i[patient] <- 2.1
    } else if (TNM.7.i[patient]=="T3bN0M0" | TNM.7.i[patient]=="T4aN0M0"){
      AJCC.7.i[patient] <- 2.2
    } else if (TNM.7.i[patient]=="T4bN0M0"){
      AJCC.7.i[patient] <- 2.3
    } else if (TNM.7.i[patient]=="T1aN1aM0" | TNM.7.i[patient]=="T2aN1aM0" | # N1a
               TNM.7.i[patient]=="T3aN1aM0" | TNM.7.i[patient]=="T4aN1aM0" |
               TNM.7.i[patient]=="T1aN2aM0" | TNM.7.i[patient]=="T2aN2aM0" | # N2a
               TNM.7.i[patient]=="T3aN2aM0" | TNM.7.i[patient]=="T4aN2aM0"){
      AJCC.7.i[patient] <- 3.1
    } else if (TNM.7.i[patient]=="T1bN1aM0" | TNM.7.i[patient]=="T2bN1aM0" | # N1a
               TNM.7.i[patient]=="T3bN1aM0" | TNM.7.i[patient]=="T4bN1aM0" |
               TNM.7.i[patient]=="T1bN2aM0" | TNM.7.i[patient]=="T2bN2aM0" | # N2a
               TNM.7.i[patient]=="T3bN2aM0" | TNM.7.i[patient]=="T4bN2aM0" |
               TNM.7.i[patient]=="T1aN1bM0" | TNM.7.i[patient]=="T2aN1bM0" | # N1b
               TNM.7.i[patient]=="T3aN1bM0" | TNM.7.i[patient]=="T4aN1bM0" |
               TNM.7.i[patient]=="T1aN2bM0" | TNM.7.i[patient]=="T2aN2bM0" | # N2b
               TNM.7.i[patient]=="T3aN2bM0" | TNM.7.i[patient]=="T4aN2bM0" |
               TNM.7.i[patient]=="T1aN2cM0" | TNM.7.i[patient]=="T2aN2cM0" | # N2c
               TNM.7.i[patient]=="T3aN2cM0" | TNM.7.i[patient]=="T4aN2cM0"){
      AJCC.7.i[patient] <- 3.2
    } else {
      AJCC.7.i[patient] <- 3.3
    }
  }

  # save TNM.7 and AJCC.7 for each imputated data set
  TNM.7 <- cbind(TNM.7, TNM.7.i)
  AJCC.7 <- cbind(AJCC.7, AJCC.7.i)

  # compute C-index
  rc <- Hmisc::rcorr.cens(-AJCC.7.i, S.Rec) # add a minus to revert class order: from worst to best
  cindex.AJCC.7[i]<-rc["C Index"]
  cindex.AJCC.7.se[i]<-rc["S.D."]/2

  # compute C-index for only positive sentinel node patients
  rc.pos <- Hmisc::rcorr.cens(-AJCC.7.i[imputation.i$SNstatus=="Positive"], S.Rec[imputation.i$SNstatus=="Positive",])
  cindex.AJCC.7.pos[i]<-rc.pos["C Index"]
  cindex.AJCC.7.se.pos[i]<-rc.pos["S.D."]/2

  # compute C-index for only negative sentinel node patients
  rc.neg <- Hmisc::rcorr.cens(-AJCC.7.i[imputation.i$SNstatus=="Negative"], S.Rec[imputation.i$SNstatus=="Negative",])
  cindex.AJCC.7.neg[i]<-rc.neg["C Index"]
  cindex.AJCC.7.se.neg[i]<-rc.neg["S.D."]/2
}
# combine C-index for each imputed data set
cindex.mi.AJCC.7<-Rubin.combine(cindex.AJCC.7,cindex.AJCC.7.se)
cindex.mi.AJCC.7.pos<-Rubin.combine(cindex.AJCC.7.pos,cindex.AJCC.7.se.pos)
cindex.mi.AJCC.7.neg<-Rubin.combine(cindex.AJCC.7.neg,cindex.AJCC.7.se.neg)

# save C-index and confidence interval
utils::write.table(rbind(c("All patients:", sprintf('%.2f', cindex.mi.AJCC.7$est),
              paste0("[", sprintf('%.2f', cindex.mi.AJCC.7$est+qnorm(.025)*cindex.mi.AJCC.7$se),
                      "; ", sprintf('%.2f', cindex.mi.AJCC.7$est+qnorm(.975)*cindex.mi.AJCC.7$se), "]")),
            c("Positive SN patients:", sprintf('%.2f', cindex.mi.AJCC.7.pos$est),
                  paste0("[", sprintf('%.2f', cindex.mi.AJCC.7.pos$est+qnorm(.025)*cindex.mi.AJCC.7.pos$se),
                         "; ", sprintf('%.2f', cindex.mi.AJCC.7.pos$est+qnorm(.975)*cindex.mi.AJCC.7.pos$se), "]")),
            c("Negative SN patients:", sprintf('%.2f', cindex.mi.AJCC.7.neg$est),
                  paste0("[", sprintf('%.2f', cindex.mi.AJCC.7.neg$est+qnorm(.025)*cindex.mi.AJCC.7.neg$se),
                         "; ", sprintf('%.2f', cindex.mi.AJCC.7.neg$est+qnorm(.975)*cindex.mi.AJCC.7.neg$se), "]"))),
            file='Z:/Project Melanoom/PaperMelanoma/Results/AJCC.7.Cindex.txt', row.names=FALSE, col.names=FALSE)

#####
##### COMPARE MODEL WITH AJCC.8 CLASIFCATION
#####
TNM.8 <- c()
AJCC.8 <- c()
cindex.AJCC.8 <- rep(0, m)
cindex.AJCC.8.se <- rep(0, m)
cindex.AJCC.8.pos <- rep(0, m)
cindex.AJCC.8.se.pos <- rep(0, m)
cindex.AJCC.8.neg <- rep(0, m)
cindex.AJCC.8.se.neg <- rep(0, m)
for (i in 1:m){
  imputation.i <- complete(imputed.data,i)
  TNM.8.i <- rep("", nrow(imputation.i))
  AJCC.8.i <- rep(0, nrow(imputation.i))
  for (patient in 1:nrow(imputation.i)){
    breslow.patient <- imputation.i[patient, "Breslow"]
    ulceration.patient <- as.character(imputation.i[patient, "Ulceration"])
    mitosis.patient <- as.character(imputation.i[patient, "Mitosis"])
    nodes.patient <- imputation.i[patient, "Tot_SNs_pos"]

    # T stage
    if (breslow.patient <= 1){
      if (ulceration.patient=="No" & breslow.patient < 0.8){
        TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T1a")
      } else {
        TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T1b")
      }
    } else if (breslow.patient>1 & breslow.patient <= 2){
      if (ulceration.patient=="No"){
        TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T2a")
      } else if (ulceration.patient=="Yes"){
        TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T2b")
      }
    } else if (breslow.patient>2 & breslow.patient <= 4){
      if (ulceration.patient=="No"){
        TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T3a")
      } else if (ulceration.patient=="Yes"){
        TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T3b")
      }
    } else if (breslow.patient>4){
      if (ulceration.patient=="No"){
        TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T4a")
      } else if (ulceration.patient=="Yes"){
        TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T4b")
      }
    }

    # N stage
    if (nodes.patient == 0){
      TNM.8.i[patient] <- paste0(TNM.8.i[patient], "N0")
    } else if (nodes.patient==1){
      TNM.8.i[patient] <- paste0(TNM.8.i[patient], "N1a")
    } else if (nodes.patient==2 | nodes.patient==3){
      TNM.8.i[patient] <- paste0(TNM.8.i[patient], "N2a")
    } else if (nodes.patient>3){
      TNM.8.i[patient] <- paste0(TNM.8.i[patient], "N3a")
    }

    # M stage for all patients in our data set
    TNM.8.i[patient] <- paste0(TNM.8.i[patient], "M0")

    # determine AJCC.8.i
    if (TNM.8.i[patient]=="T1aN0M0"){
      AJCC.8.i[patient] <- 1.1
    } else if (TNM.8.i[patient]=="T1bN0M0" | TNM.8.i[patient]=="T2aN0M0"){
      AJCC.8.i[patient] <- 1.2
    } else if (TNM.8.i[patient]=="T2bN0M0" | TNM.8.i[patient]=="T3aN0M0"){
      AJCC.8.i[patient] <- 2.1
    } else if (TNM.8.i[patient]=="T3bN0M0" | TNM.8.i[patient]=="T4aN0M0"){
      AJCC.8.i[patient] <- 2.2
    } else if (TNM.8.i[patient]=="T4bN0M0"){
      AJCC.8.i[patient] <- 2.3
    } else if (TNM.8.i[patient]=="T1aN1aM0" | TNM.8.i[patient]=="T1bN1aM0" | TNM.8.i[patient]=="T2aN1aM0"
               | TNM.8.i[patient]=="T1aN2aM0" | TNM.8.i[patient]=="T1bN2aM0" | TNM.8.i[patient]=="T2aN2aM0"){
      AJCC.8.i[patient] <- 3.1
    } else if (TNM.8.i[patient]=="T2bN1aM0" | TNM.8.i[patient]=="T3aN1aM0"){
      AJCC.8.i[patient] <- 3.2
    } else if (TNM.8.i[patient]=="T4bN3aM0"){
      AJCC.8.i[patient] <- 3.4
    } else {
      AJCC.8.i[patient] <- 3.3
    }

  }

  # save TNM.8 and AJCC.8 for each imputated data set
  TNM.8 <- cbind(TNM.8, TNM.8.i)
  AJCC.8 <- cbind(AJCC.8, AJCC.8.i)

  # compute C-index
  rc <- Hmisc::rcorr.cens(-AJCC.8.i, S.Rec) # add a minus to revert class order: from worst to best
  cindex.AJCC.8[i]<-rc["C Index"]
  cindex.AJCC.8.se[i]<-rc["S.D."]/2

  # compute C-index for only positive sentinel node patients
  rc.pos <- Hmisc::rcorr.cens(-AJCC.8.i[imputation.i$SNstatus=="Positive"], S.Rec[imputation.i$SNstatus=="Positive",])
  cindex.AJCC.8.pos[i]<-rc.pos["C Index"]
  cindex.AJCC.8.se.pos[i]<-rc.pos["S.D."]/2

  # compute C-index for only negative sentinel node patients
  rc.neg <- Hmisc::rcorr.cens(-AJCC.8.i[imputation.i$SNstatus=="Negative"], S.Rec[imputation.i$SNstatus=="Negative",])
  cindex.AJCC.8.neg[i]<-rc.neg["C Index"]
  cindex.AJCC.8.se.neg[i]<-rc.neg["S.D."]/2
}
# combine C-index for each imputed data set
cindex.mi.AJCC.8<-Rubin.combine(cindex.AJCC.8,cindex.AJCC.8.se)
cindex.mi.AJCC.8.pos<-Rubin.combine(cindex.AJCC.8.pos,cindex.AJCC.8.se.pos)
cindex.mi.AJCC.8.neg<-Rubin.combine(cindex.AJCC.8.neg,cindex.AJCC.8.se.neg)

# save C-index and confidence interval
utils::write.table(rbind(c("All patients:", sprintf('%.2f', cindex.mi.AJCC.8$est),
                        paste0("[", sprintf('%.2f', cindex.mi.AJCC.8$est+qnorm(.025)*cindex.mi.AJCC.8$se),
                               "; ", sprintf('%.2f', cindex.mi.AJCC.8$est+qnorm(.975)*cindex.mi.AJCC.8$se), "]")),
                  c("Positive SN patients:", sprintf('%.2f', cindex.mi.AJCC.8.pos$est),
                        paste0("[", sprintf('%.2f', cindex.mi.AJCC.8.pos$est+qnorm(.025)*cindex.mi.AJCC.8.pos$se),
                               "; ", sprintf('%.2f', cindex.mi.AJCC.8.pos$est+qnorm(.975)*cindex.mi.AJCC.8.pos$se), "]")),
                  c("Negative SN patients", sprintf('%.2f', cindex.mi.AJCC.8.neg$est),
                        paste0("[", sprintf('%.2f', cindex.mi.AJCC.8.neg$est+qnorm(.025)*cindex.mi.AJCC.8.neg$se),
                               "; ", sprintf('%.2f', cindex.mi.AJCC.8.neg$est+qnorm(.975)*cindex.mi.AJCC.8.neg$se), "]"))),
            file='Z:/Project Melanoom/PaperMelanoma/Results/AJCC.8.Cindex.txt', row.names=FALSE, col.names=FALSE)

# table of number of AJCC.7 and AJCC.8
table.AJCC.7 <- c()
table.AJCC.8 <- c()
for (i in 1:m){
  table.AJCC.7 <- cbind(table.AJCC.7, as.numeric(table(AJCC.7[,i])))
  table.AJCC.8 <- cbind(table.AJCC.8, as.numeric(table(AJCC.8[,i])))
}
table.AJCC.7.8 <- cbind(as.numeric(names(table(AJCC.8[,1]))),
                        c(round(rowMeans(table.AJCC.7), 0), 0),
                        round(rowMeans(table.AJCC.8), 0))
utils::write.table(table.AJCC.7.8,
            file='Z:/Project Melanoom/PaperMelanoma/Results/table.AJCC.7.and.8.txt', row.names=FALSE, col.names=FALSE)
table.AJCC.7.8
```

# Full model for recurrence
```{r, eval=TRUE}
form.full.Rec <- S.Rec ~
  SNstatus +
  Sex +
  log(Age.SN) +
  Ulceration +
  Loc_CAT +
  Histology +
  log(Breslow) +
  multiple.fields +
  Tot_SNs_neg +
  Tot_SNs_pos +
  log(Rdamcrit) +
  Dewar +
  SNstatus:Sex +
  SNstatus:log(Age.SN) +
  SNstatus:Ulceration +
  SNstatus:Loc_CAT +
  SNstatus:Histology +
  SNstatus:log(Breslow) +
  SNstatus:multiple.fields
f.mi.full.Rec <- Hmisc::fit.mult.impute(form.full.Rec, cph, xtrans=imputed.data,
                                 data=data.to.be.imputed,n.impute=m, pr=FALSE,
                                 fit.reps=TRUE, y=TRUE, x=TRUE, se.fit=TRUE)
f.mi.full.Rec
stats::anova(f.mi.full.Rec)
summary(f.mi.full.Rec)
```

### Test proportional hazards assumption
```{r, eval=TRUE}
test.prop<-survival::cox.zph(f.mi.full.Rec$fits[[1]])
test.prop
plot(test.prop)
```

# Final model for recurrence selected by backward selection with p = 0.01
```{r, eval=TRUE}
c.Breslow <- mean(log(complete.imputed.data$Breslow))
form.BS.Rec <- S.Rec ~
  SNstatus +
  log(Age.SN) +
  Ulceration +
  Loc_CAT +
  I(log(Breslow)-c.Breslow) +
  log(Rdamcrit) +
  SNstatus:I(log(Breslow)-c.Breslow)
f.mi.BS.Rec <- Hmisc::fit.mult.impute(form.BS.Rec,cph,xtrans=imputed.data,data=data.to.be.imputed,n.impute=m,
                           pr=FALSE,fit.reps=TRUE,y=TRUE,x=TRUE,se.fit=TRUE)
round(stats::anova(f.mi.BS.Rec)[order(stats::anova(f.mi.BS.Rec)[, 'P']),], 3)
```

# Refitted full model for melanoma specific mortality
```{r, eval=TRUE}
form.full.MSM.refit <- S.MSM ~
  SNstatus +
  Sex +
  log(Age.SN) +
  Ulceration +
  Loc_CAT +
  Histology +
  log(Breslow) +
  multiple.fields +
  Tot_SNs_neg +
  Tot_SNs_pos +
  log(Rdamcrit) +
  Dewar +
  SNstatus:Sex +
  SNstatus:log(Age.SN) +
  SNstatus:Ulceration +
  SNstatus:Loc_CAT +
  SNstatus:Histology +
  SNstatus:log(Breslow) +
  SNstatus:multiple.fields
f.mi.full.MSM.refit <- Hmisc::fit.mult.impute(form.full.MSM.refit,cph,xtrans=imputed.data,data=data.to.be.imputed,n.impute=m,
                                 pr=FALSE,fit.reps=TRUE,y=TRUE,x=TRUE,se.fit=TRUE)
f.mi.full.MSM.refit
summary(f.mi.full.MSM.refit)
stats::anova(f.mi.full.MSM.refit)
```

# Refitted final model for melanoma specific mortality
```{r, eval=TRUE}
form.BS.MSM.refit <- S.MSM ~
  SNstatus +
  log(Age.SN) +
  Ulceration +
  Loc_CAT +
  I(log(Breslow)-c.Breslow) +
  log(Rdamcrit) +
  SNstatus:I(log(Breslow)-c.Breslow)
f.mi.BS.MSM.refit <- Hmisc::fit.mult.impute(form.BS.MSM.refit,cph,xtrans=imputed.data,data=data.to.be.imputed,n.impute=m,
                                  pr=FALSE,fit.reps=TRUE,y=TRUE,x=TRUE,se.fit=TRUE)
round(anova(f.mi.BS.MSM.refit)[order(anova(f.mi.BS.MSM.refit)[, 'P']),], 3)
f.mi.BS.MSM.refit
summary(f.mi.BS.MSM.refit)
stats::anova(f.mi.BS.MSM.refit)
```

# Calibrated final model for melanoma specific mortality
Fit disease specific mortality to the linear predictor of recurrence
```{r, eval=TRUE}
lp.BS<-f.mi.BS.Rec$linear
f.mi.BS.MSM<-rms::cph(S.MSM~lp.BS,y=TRUE,x=TRUE)
utils::write.table(paste0(sprintf("%.2f", stats::coef(f.mi.BS.MSM)), " [",
            sprintf("%.2f", as.numeric(stats::confint(f.mi.BS.MSM))[1]), "; ",
            sprintf("%.2f", as.numeric(stats::confint(f.mi.BS.MSM))[2]), "]"),
            file='Z:/Project Melanoom/PaperMelanoma/Results/calibration.MSS.txt', row.names=FALSE, col.names=FALSE)
MSM.cal.fact <- stats::coef(f.mi.BS.MSM)
stats::coef(f.mi.BS.MSM)
stats::confint(f.mi.BS.MSM)
```

### Create variables with results
##### Full model
```{r, eval=TRUE}
for (model in c('Rec', 'MSM.refit')){
  f.mi.full <- eval(parse(text=paste0('f.mi.full.', model)))

  CI.lower.neg <- exp(stats::confint(f.mi.full)[,1])
  CI.upper.neg <- exp(stats::confint(f.mi.full)[,2])
  CI.neg <- paste0("[", sprintf("%.2f", CI.lower.neg), "; ", sprintf("%.2f", CI.upper.neg), "]")
  CI.lower.pos <- c(CI.lower.neg[2:12]*CI.lower.neg[20:30], CI.lower.neg[14:19])
  CI.upper.pos <- c(CI.upper.neg[2:12]*CI.upper.neg[20:30], CI.upper.neg[14:19])
  CI.pos <- paste0("[", sprintf("%.2f", CI.lower.pos), "; ", sprintf("%.2f", CI.upper.pos), "]")

  anova.full <- stats::anova(f.mi.full)[,1]

  full.table <- cbind(c(NA, sprintf("%.2f", exp(f.mi.full$coefficients[2:13])), rep(NA, 5)), # COEF NEG
                      c(NA, CI.neg[2:13], rep(NA, 5)), # CI NEG
                      sprintf("%.2f", c(exp(f.mi.full$coefficients[1]), # COEF POS
                        exp(f.mi.full$coefficients[2:12])*exp(f.mi.full$coefficients[20:30]),
                        exp(f.mi.full$coefficients[14]), exp(f.mi.full$coefficients[15:19]))),
                      c(CI.neg[1], CI.pos), # CI POS
                      c(sprintf("%.1f", anova.full[c(1, 3, 5, 7, 9)]), rep(NA, 2), # CHI2 NEG
                            sprintf("%.1f", anova.full[11]), rep(NA, 2),
                            sprintf("%.1f", anova.full[c(13, 15, 17, 19, 20)]), rep(NA, 3)),
                      c(sprintf("%.1f", anova.full[c(2, 21, 22, 23, 24)]), rep(NA, 2), # CHI2 POS
                            sprintf("%.1f", anova.full[25]), rep(NA, 2),
                            sprintf("%.1f", anova.full[c(26:27, 18)]), rep(NA, 5)))

  ncols <- 6
  final.table <- rbind(full.table[1:4,],
                       rep(NA, ncols), # location
                       c("(ref)", NA, "(ref)", rep(NA, 3)), # ref arm
                       full.table[5:7,], # other locations
                       rep(NA, ncols), # histology
                       c("(ref)", NA, "(ref)", rep(NA, 3)), # ref SMM
                       full.table[8:14,],
                       rep(NA, ncols), # Dewar
                       c(rep(NA, 2), "(ref)", rep(NA, 3)), # ref subcap
                       full.table[15:18,]) # other Dewar

  colnames(final.table) <- c("HR.neg", "CI.neg", "HR.pos", "CI.pos", "Chi2.neg", "Chi2.pos")
  assign(paste0('final.full.table.', model), as.data.frame(final.table))
}
```

##### Final model
```{r, eval=TRUE}
for (model in c('Rec', 'MSM', 'MSM.refit')){
  if (model=="MSM"){
    mult.factor <- MSM.cal.fact # if MSM, multiply by MSM.cal.fact
    f.mi.BS <- eval(parse(text=paste0('f.mi.BS.Rec'))) # use recurrence model if calibrated MSM
  }
  else{
    mult.factor <- 1 # if Rec or MSM.refit, don't multiply
    f.mi.BS <- eval(parse(text=paste0('f.mi.BS.', model)))
  }

  CI.lower.neg <- exp(mult.factor*confint(f.mi.BS)[,1])
  CI.upper.neg <- exp(mult.factor*confint(f.mi.BS)[,2])
  CI.neg <- paste0("[", sprintf("%.2f", CI.lower.neg), "; ", sprintf("%.2f", CI.upper.neg), "]")

  anova.BS <- anova(f.mi.BS)[,1]

  CI.pos <- paste0("[", sprintf("%.2f", exp(mult.factor*confint(f.mi.BS)[7,1])*exp(mult.factor*confint(f.mi.BS)[9,1])),
                   "; ", sprintf("%.2f", exp(mult.factor*confint(f.mi.BS)[7,2])*exp(mult.factor*confint(f.mi.BS)[9,2])), "]")

  BS.table <- cbind(c(NA, sprintf("%.2f", exp(mult.factor*f.mi.BS$coefficients[2:7])), NA), # COEF NEG
                    c(NA, CI.neg[2:7], NA), # CI NEG
                    sprintf("%.2f", exp(mult.factor*f.mi.BS$coefficients[1:8])*c(rep(1, 6), exp(mult.factor*f.mi.BS$coefficients[9]), 1)), #COEF POS
                    c(CI.neg[1:6], CI.pos, CI.neg[8]), # CI POS
                    c(NA, sprintf("%.1f", anova.BS[3:5]), rep(NA, 2), sprintf("%.1f", anova.BS[c(6, 8)])), # CHI2 NEG
                    c(sprintf("%.1f", anova.BS[1]), rep(NA, 5), # CHI2 POS
                      sprintf("%.1f", anova.BS[9]), NA))

  BS.table <- rbind(BS.table[1:3,], rep(NA, 6), c("(ref)", NA, "(ref)", rep(NA, 3)), BS.table[4:8,])

  colnames(BS.table) <- c("HR.neg", "CI.neg", "HR.pos", "CI.pos", "Chi2.neg", "Chi2.pos")

  assign(paste0('final.BS.table.', model), as.data.frame(BS.table))
}
```

# Plot calibration of all centers combined and all centers seperate for recurrence and melanoma specific mortality refit and write results to Excel
```{r, eval=TRUE}
horizon <- 5
for (model in list('Rec', 'MSM.refit')){
  if (model == "Rec"){
    # plot predict
    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/plot.predict.Rec.png"))
    plot.predict <- plot(Predict(f.mi.full.Rec))
    show(plot.predict)
    grDevices::dev.off()

    # plot predict Breslow
    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/plot.predict.Breslow.Rec.png"))
    plot.Breslow <- plot(Predict(f.mi.full.Rec, Breslow=0:20, SNstatus), ~ Breslow, groups="SNstatus",
                         adj.subtitle=FALSE, lwd=2, labels=FALSE,
                         par.settings=list(superpose.line=list(lty=c(1, 2), col=c("red", "blue"))),
                         key=list(cex=.8,corner = c(0.1, .9),lines=list(col=c("red","blue"), lty=c(1,2), lwd=2),text=list(c("Negative","Positive"))))
    show(plot.Breslow)
    grDevices::dev.off()
  }

  # discrimination full and backward selected model
  for (type in c("full", "BS")){
    f.mi <- eval(parse(text=paste0('f.mi.', type, '.', model)))
    if (model=="Rec"){
      S <- S.Rec
    }
    else if (model=="MSM.refit"){
      S <- S.MSM
    }
    p <- matrix(NA, nrow(sel.data), m)
    for (i in 1:m){
      # predicted lp for imputation i
      lp.i <- predict(f.mi,newdata=complete(imputed.data,i), type="lp")
      # baseline hazard
      f.basehaz <- survival::basehaz(f.mi,TRUE)
      h0.i <- f.basehaz$hazard[f.basehaz$time==max(f.basehaz$time[f.basehaz$time<=horizon])]
      # probabilities
      p.i <- fun.event(lp=lp.i, h0=h0.i)
      p[, i] <- p.i
    }

    # omit the observation with predicted probability = 1
    omit <- unique(c(which(p[,1]==1), which(p[,2]==1), which(p[,3]==1), which(p[,4]==1), which(p[,5]==1)))
    if (length(omit)>0){
      cat('Omit:', omit, '\n')
      p <- p[-omit,]
      S <- S[-omit,]
    }

    out <- val.surv.mi(p,S,g=5,main="All centers",
                       file_path=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/all.centers.", model, '.', type, ".png"),
                       save_plots=TRUE,time=NULL,lim=c(0,1),dist=TRUE)

    final.full.table <- eval(parse(text=paste0('final.', type, '.table.', model)))
    final.full.table["C-index", "HR.neg"] <- sprintf('%.2f', out$cindex)
    final.full.table["C-index", "CI.neg"] <- paste0("[", sprintf('%.2f', out$cindex.lower),
                                                   "; ", sprintf('%.2f', out$cindex.upper), "]")

    assign(paste0('final.', type, '.table.', model), final.full.table)
  }
}

# MSM recalibration
for (type in c("full", "BS")){
  f.mi <- eval(parse(text=paste0('f.mi.', type, '.Rec')))
  cindex.MSM.refit <- rep(0, m)
  cindex.se.MSM.refit <- rep(0, m)
  for (i in 1:m){
    rc <- Hmisc::rcorr.cens(-predict(f.mi,newdata=complete(imputed.data,i), type="lp"), S.MSM)
    cindex.MSM.refit[i]<-rc["C Index"]
    cindex.se.MSM.refit[i]<-rc["S.D."]/2
  }
  cindex.mi.MSM.refit<-Rubin.combine(cindex.MSM.refit,cindex.se.MSM.refit)

  final.full.table <- eval(parse(text=paste0('final.', type, '.table.Rec')))
  final.full.table["C-index MSM", "HR.neg"] <- sprintf('%.2f', cindex.mi.MSM.refit$est)
  final.full.table["C-index MSM", "CI.neg"] <- paste0("[", sprintf('%.2f', cindex.mi.MSM.refit$est+qnorm(.025)*cindex.mi.MSM.refit$se),
                                                 "; ", sprintf('%.2f', cindex.mi.MSM.refit$est+qnorm(.975)*cindex.mi.MSM.refit$se), "]")

  assign(paste0('final.', type, '.table.Rec'), final.full.table)
}

# replace NA by blank
final.full.table.Rec[is.na(final.full.table.Rec)] <- ""
final.full.table.MSM.refit[is.na(final.full.table.MSM.refit)] <- ""
final.BS.table.Rec[is.na(final.BS.table.Rec)] <- ""
final.BS.table.MSM[is.na(final.BS.table.MSM)] <- ""
final.BS.table.MSM["C-index MSM",] <- final.BS.table.Rec[nrow(final.BS.table.Rec),]
final.BS.table.MSM.refit[is.na(final.BS.table.MSM.refit)] <- ""

# save tables
write.table(final.full.table.Rec,file='Z:/Project Melanoom/PaperMelanoma/Results/final.full.table.Rec.txt')
write.table(final.full.table.MSM.refit,file='Z:/Project Melanoom/PaperMelanoma/Results/final.full.table.MSM.refit.txt')
write.table(final.BS.table.Rec,file='Z:/Project Melanoom/PaperMelanoma/Results/final.BS.table.Rec.txt')
write.table(final.BS.table.MSM,file='Z:/Project Melanoom/PaperMelanoma/Results/final.BS.table.MSM.txt')
write.table(final.BS.table.MSM.refit,file='Z:/Project Melanoom/PaperMelanoma/Results/final.BS.table.MSM.refit.txt')
final.full.table.Rec
final.full.table.MSM.refit
final.BS.table.Rec
final.BS.table.MSM
final.BS.table.MSM.refit
```

# Cross-validation recurrence 5-years
```{r, eval=TRUE}
Centers <- sort(as.character(unique(sel.data$Center)))
Centers.anonimized <- c("A", "B", "C", "D")
cuts <- 5
for (model in c('Rec', 'MSM')){
  if (model == 'Rec'){
    S.model <- S.Rec
  }
  if (model == 'MSM'){
    S.model <- S.MSM
  }

  for (type in c('full', 'BS')){
    for (j in 1:4){
      S.j<-S.model[data.to.be.imputed$Center==Centers[j],]
      S.notj<-S.model[data.to.be.imputed$Center!=Centers[j],]
      S.Rec.notj<-S.Rec[data.to.be.imputed$Center!=Centers[j],]

      form.model<-eval(parse(text=paste0('form.', type, '.Rec')))
      form.notj<-update(form.model, S.Rec.notj  ~ . )
      f.mi.Rec.notj<-Hmisc::fit.mult.impute(form.notj,cph,xtrans=imputed.data,data=data.to.be.imputed,
                                 n.impute=m,pr=FALSE,fit.reps=TRUE,y=TRUE,x=TRUE,se.fit=TRUE,
                                 sub=data.to.be.imputed$Center!=Centers[j])

      if (model=="Rec"){
        f.basehaz.notj<-survival::basehaz(f.mi.Rec.notj,TRUE)
        h0.notj<-f.basehaz.notj$hazard[f.basehaz.notj$time==max(f.basehaz.notj$time[f.basehaz.notj$time<=horizon])]
      } else if (model=="MSM"){
        lp.notj<-f.mi.Rec.notj$linear
        f.MSM.notj<-cph(S.notj~lp.notj,y=TRUE,x=TRUE)

        f.basehaz.notj<-basehaz(f.MSM.notj,TRUE)
        h0.notj<-f.basehaz.notj$hazard[f.basehaz.notj$time==max(f.basehaz.notj$time[f.basehaz.notj$time<=horizon])]
      }

      for (select.value in c("Positive", "Negative", "All")){
        # identify those patients that are selected
        if (select.value == "All"){
          select.patients <- TRUE
        } else if (select.value != "All"){
          select.patients <- data.to.be.imputed[data.to.be.imputed$Center==Centers[j],]$SNstatus==select.value
        }

        # calibration with metrics
        S <- S.j[select.patients,]
        p <- matrix(NA, nrow(S), m)
        for (i in 1:m){
          lp.i<-predict(f.mi.Rec.notj,newdata=complete(imputed.data,i)[data.to.be.imputed$Center==Centers[j],], type="lp")

          if (model=="MSM"){
            lp.i<-predict(f.MSM.notj,newdata=lp.i)
          }

          if (select.value != "All"){
            lp.i <- lp.i[select.patients]
          }

          # baseline hazard
          if (model=="Rec"){
            f.basehaz<-basehaz(f.mi.Rec.notj,TRUE)
          }
          else if (model=="MSM"){
            f.basehaz<-basehaz(f.MSM.notj,TRUE)
          }
          h0.i <- f.basehaz$hazard[f.basehaz$time==max(f.basehaz$time[f.basehaz$time<=horizon])]

          # probabilities
          p.i <- fun.event(lp=lp.i, h0=h0.i)
          p[, i] <- p.i
        }

        # omit the observation with predicted probability = 1
        omit <- unique(c(which(p[,1]==1), which(p[,2]==1), which(p[,3]==1), which(p[,4]==1), which(p[,5]==1)))
        if (length(omit)>0){
          cat('Omit:', omit, '\n')
          p <- p[-omit,]
          S <- S[-omit,]
        }

        out <- val.surv.mi(p, S, g=5, main=paste("Center", Centers.anonimized[j]),
                           file_path=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/", model, '.', select.value, '.', type, '.', Centers[j], ".png"),
                           save_plots=TRUE,time=NULL,lim=c(0,1),dist=TRUE)
      }
    }
  }
}

# combine calibration plots into one
for (select.value in c("All", "Positive", "Negative")){
  for (y in c("Rec", "MSM")){
    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/", y, ".", select.value, ".png"), width=16, height=16, units="cm", res=300)
    par(mar=rep(0, 4))
    layout(matrix(1:4, ncol=2, byrow=TRUE))
    Centers.val <- rep(1:4, 3)
    for (i in 1:4){
      plot(NA, xlim=0:1, ylim=0:1, xaxt="n", yaxt="n", bty="n")
      img <- readPNG(paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/", y, ".", select.value, ".BS.", Centers[Centers.val[i]], ".png"))
      rasterImage(img, 0, 0, 1, 1)
    }
    grDevices::dev.off()
  }
}
```

# Nomogram
```{r, eval=TRUE}
# Limits of Age, Breslow, Rdamcrit
Age.class <- seq(20, 80, 20)
Breslow.class <- c(seq(0.1, 0.9, 0.2), 1:7)
Rdamcrit.class <- c(seq(1, 10, 2)/100, seq(1, 10, 2)/10, seq(1, 10, 2))

# initial nomogram
nom.0 <- nomogram(f.mi.BS.Rec,maxscale=10,
                Breslow=Breslow.class, Rdamcrit=Rdamcrit.class, Age.SN=Age.class)

# Baseline hazard
rc.Rec <- (1/attributes(nom.0)$info$sc)
int.Rec <- attributes(nom.0)$info$Intercept
f.Rec.basehaz <- basehaz(f.mi.BS.Rec,TRUE)
h0.Rec <- f.Rec.basehaz$hazard[f.Rec.basehaz$time==max(f.Rec.basehaz$time[f.Rec.basehaz$time<=horizon])]
lp.range <- int.Rec+rc.Rec*(0:10)*2
par(mar = rep(0, 4))
nom <- nomogram(f.mi.BS.Rec, maxscale=10,
              Breslow=Breslow.class, Rdamcrit=Rdamcrit.class, Age.SN=Age.class,
              lp=TRUE, fun.at=fun.event(lp=lp.range, h0=h0.Rec))
attributes(nom)$names[attributes(nom)$names=="Age.SN"]<-"Age"
attributes(nom)$names[attributes(nom)$names=="Loc_CAT"]<-"Location"
attributes(nom)$names[attributes(nom)$names=="Rdamcrit"]<-"Tumor burden"
attributes(nom)$names[attributes(nom)$names=="Breslow (SNstatus=Positive)"]<-"Breslow + positive SN"
attributes(nom)$names[attributes(nom)$names=="Breslow (SNstatus=Negative)"]<-"Breslow + negative SN"
plot(nom, total.sep.page=TRUE, col.grid=gray(c(0.8, 0.95)))
nom.0
nom

# Slope and intercept of recurrence and MSM
int.Rec
# check that int is same as this:
log(20)*f.mi.BS.Rec$coef["Age.SN"]+log(0.1)*f.mi.BS.Rec$coef["Breslow"]+log(0.01)*f.mi.BS.Rec$coef["Rdamcrit"]-f.mi.BS.Rec$center
```

# Risk score distribution
```{r, eval=TRUE}
rc.MSM <- f.mi.BS.MSM$coef[1]*rc.Rec
int.MSM <- f.mi.BS.MSM$coef[1]*int.Rec

max.range <- 16
points.range <- 0:max.range
lp.points.Rec <- int.Rec+rc.Rec*points.range
lp.points.MSM <- int.MSM+rc.MSM*points.range

f.MSM.basehaz <- basehaz(f.mi.BS.MSM,TRUE)
h0.MSM <- f.MSM.basehaz$hazard[f.MSM.basehaz$time==max(f.MSM.basehaz$time[f.MSM.basehaz$time<=horizon])]
lp.MSM <- f.mi.BS.MSM$linear

Breslow.trunc <- complete.imputed.data$Breslow
Breslow.trunc[complete.imputed.data$Breslow<0.1] <- .1
Breslow.trunc[complete.imputed.data$Breslow>7] <- 7
lp.score.Rec <- predict(f.mi.BS.Rec,newdata=data.frame(SNstatus=complete.imputed.data$SNstatus,
                                                Age.SN=complete.imputed.data$Age.SN,
                                                Ulceration=complete.imputed.data$Ulceration,
                                                Loc_CAT=complete.imputed.data$Loc_CAT,
                                                Breslow=Breslow.trunc,
                                                Rdamcrit=complete.imputed.data$Rdamcrit))
mean(lp.score.Rec)
score.Rec <- round((lp.score.Rec-int.Rec)/rc.Rec,0)
max.breaks <- max.range+1
h.Rec <- hist(score.Rec,plot=FALSE,breaks=0:max.breaks,right=FALSE)

# score table
data.frame(table(score.Rec)/length(score.Rec))
```

# Check mean event rates of risk distribution
```{r, eval=TRUE}
# Recurrence
sum(h.Rec$density*fun.event(lp=lp.points.Rec, h0=h0.Rec))
# equal to the 5-year survival probability, kaplanmeier op 5 jaar, survfit op 5 jaar
SF<-survfit(S.Rec~1,data=sel.data)
100*(1-SF$surv[SF$time==max(SF$time[SF$time<=horizon])])
mean(fun.event(lp=lp.score.Rec, h0=h0.Rec))

# MSM
sum(h.Rec$density*fun.event(lp=lp.points.MSM, h0=h0.MSM))
SF<-survfit(S.MSM~1,data=sel.data)
100*(1-SF$surv[SF$time==max(SF$time[SF$time<=horizon])])
mean(fun.event(lp=lp.MSM, h0=h0.MSM))
```

# Plot of risk score distribution
```{r, eval=TRUE, message=FALSE}
h.plot <- h.Rec
h.plot$density <- 100*h.Rec$density
p.Rec <- fun.event(lp=lp.points.Rec, h0=h0.Rec)
p.MSM <- fun.event(lp=lp.points.MSM, h0=h0.MSM)
x.lim <- c(0, max.range+1)
y.lim <- c(0, 100)
y.lim.hist <- c(0, 25)
png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Risk.distribution.png"))
par(mar = c(5,5,2,5))
plot(h.plot,freq=FALSE,axes=FALSE, xlab=NA, xlim=x.lim, ylab=NA,ylim=y.lim.hist,main=NA,col='white')
abline(h=(0:30)*1,col="light grey",lty=2)
axis(side = 4)
mtext(side = 4, line = 3, 'Risk score distribution (%)')
par(new = TRUE)
plot(h.plot,freq=FALSE,axes=FALSE, xlab=NA, xlim=x.lim, ylab=NA,ylim=y.lim.hist,main=NA,col="white")
par(new = TRUE)
plot(points.range,p.Rec*100,type="b",lwd=1,xlim=x.lim,ylim=y.lim,pch=15, xlab="Risk score",ylab="Predicted 5-year risk (%)")
points(points.range,p.MSM*100,type="b",lwd=1,xlim=x.lim,ylim=y.lim,pch=16)
legend(x=0,y=100,legend=c("Recurrence","Melanoma-specific mortality"),pch=c(15,16),cex=.9,bg='white')
grDevices::dev.off()
# table of risk distribution
cbind(points.range, round(p.Rec*100, 1), round(p.MSM*100, 1))[c(1, 9, 10, 12, 13),]

# save for Rshiny app
coef.Rec <- coef(f.mi.BS.Rec)
center.Rec <- f.mi.BS.Rec$center
save(h0.Rec, h0.MSM, coef.Rec, c.Breslow, center.Rec,
     file=paste0('Z:/Project Melanoom/PaperMelanoma/Results/h0.Rec.MSM.Rdata'), compress=TRUE)
load(file=paste0('Z:/Project Melanoom/PaperMelanoma/Results/h0.Rec.MSM.Rdata'))
```

# Example to check probabilities with app
```{r, eval=TRUE}
patient.nr <- 115 # 395 and 396 = Negative, 397 and 398 = Positive, 30 = Head & neck, 115 has large Breslow
data.to.be.imputed[patient.nr, c("SNstatus", "Age.SN", "Ulceration", "Loc_CAT", "Breslow", "Rdamcrit")]
lp.score.Rec.true <- predict(f.mi.BS.Rec)[patient.nr]
lp.score.Rec.try <- predict(f.mi.BS.Rec, newdata=data.frame(SNstatus=data.to.be.imputed[patient.nr,"SNstatus"],
                                                            Age.SN=data.to.be.imputed[patient.nr,"Age.SN"],
                                                            Ulceration=data.to.be.imputed[patient.nr,"Ulceration"],
                                                            Loc_CAT=data.to.be.imputed[patient.nr,"Loc_CAT"],
                                                            Breslow=data.to.be.imputed[patient.nr,"Breslow"],
                                                            Rdamcrit=data.to.be.imputed[patient.nr,"Rdamcrit"]),
                                                            type="lp")
p.Rec.example <- fun.event(lp=lp.score.Rec.try, h0=h0.Rec)
p.MSM.example <- fun.event(lp=lp.score.Rec.try, h0=h0.MSM)

# score
round((lp.score.Rec.try-int.Rec)/rc.Rec,0)

# manual calculation
coef.Rec <- coef(f.mi.BS.Rec)
lp.manual <- coef.Rec["SNstatus=Positive"]*(as.numeric(data.to.be.imputed[patient.nr,"SNstatus"])-1)+
              coef.Rec["Age.SN"]*log(data.to.be.imputed[patient.nr,"Age.SN"])+
              coef.Rec["Ulceration=Yes"]*(as.numeric(data.to.be.imputed[patient.nr,"Ulceration"])-1)+
              coef.Rec["Loc_CAT=leg"]*as.numeric(data.to.be.imputed[patient.nr,"Loc_CAT"]=="leg")+
              coef.Rec["Loc_CAT=trunk"]*as.numeric(data.to.be.imputed[patient.nr,"Loc_CAT"]=="trunk")+
              coef.Rec["Loc_CAT=head & neck"]*as.numeric(data.to.be.imputed[patient.nr,"Loc_CAT"]=="head & neck")+
              coef.Rec["Breslow"]*(log(data.to.be.imputed[patient.nr,"Breslow"])-c.Breslow)+
              coef.Rec["Rdamcrit"]*log(data.to.be.imputed[patient.nr,"Rdamcrit"])+
              coef.Rec["SNstatus=Positive * Breslow"]*(log(data.to.be.imputed[patient.nr,"Breslow"])-c.Breslow)*(as.numeric(data.to.be.imputed[patient.nr,"SNstatus"])-1)
center.Rec <- f.mi.BS.Rec$center

# lp
lp.score.Rec.true
lp.score.Rec.try
lp.manual-center.Rec

# probability of recurrence
fun.event(lp=lp.score.Rec.try, h0=h0.Rec)
1-exp(-h0.Rec*exp(lp.manual-center.Rec))

# check MSM
lp.score.MSM.true <- predict(f.mi.BS.MSM)[patient.nr]
lp.score.MSM.true
lp.MSM.try <- MSM.cal.fact*(lp.manual-center.Rec)
lp.MSM.try
fun.event(lp=lp.MSM.try, h0=h0.MSM)
fun.event(lp=lp.MSM.try, h0=h0.MSM)
1-exp(-h0.MSM*exp(MSM.cal.fact*(lp.manual-center.Rec)))
```