### 2022-11-29 Code/0_PrepareTrainingData.Rmd
Validation data added

---
title: "Prepare Training Data"
author: "Carolien C.H.M. Maas"
date: "14-11-2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Training data preparation

### Load packages, functions, and data 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# erase memory
rm(list = ls(all.names = TRUE))

# for reproducibility
set.seed(100)

# load libraries
library(foreign)
library(rms)
source("Z:/Project Melanoom/PaperMelanoma/Code/PaperMelanomaFunctions.R")

# load data file
original.train.data <- foreign::read.spss(file="Z:/Project Melanoom/PaperMelanoma/Data/training/Dataset Predictiemodel.sav",
                                    to.data.frame=TRUE, origin="01-01-1970")

# Select necessary variables
selected.vars <- c("EXCLUSIE", "DatabaseNR", "Center", "Sex", "Age", "DOB",
                   "Ulceration", "Loc_CAT",            # OLD data set: Loc_CAT2
                   "Histology", "Breslow", "Tot_nr_SNs",
                   "nr_SNs_Field_1", "nr_SNs_Field_2", "nr_SNs_Field_3",
                   "Mitosis",                          # needed for AJCC.7.i
                   "SNdate", "Recurrence", "Date_1st_Rec", "Last_FU",
                   "dead", "Status",
                   "SNstatus",                         # distinguish between positive and negative SN patients
                   "Tot_SNs_pos", "Rdamcrit", "Dewar") # additional for positive SN patients
```

### Inclusion criteria
Exclude when Exclusion = 2: based on duplicates, urogenital melanoma, etc. Note:
Leave the patients included that have EXCLUSIE = 1 and Dbstatus = 8 (Dbstatus=0: 
negative, Dbstatus=1: positive), even though diagnosis and operation date are 
very close.
```{r, eval=TRUE}
sel.train.data <- original.train.data[original.train.data$EXCLUSIE!="Exclusion", selected.vars]
cat("Number of patients excluded:", nrow(original.train.data)-nrow(sel.train.data), "\n")
```
Omit one patient that has no last follow-up date
```{r, eval=TRUE, echo=FALSE}
sel.train.data <- sel.train.data[!is.na(sel.train.data$Last_FU),]
```

### Recode
#### Use age at SN instead of age at diagnosis
```{r, eval=TRUE}
sel.train.data$Age.SN <- floor((sel.train.data$SNdate-sel.train.data$DOB)/(24*60*60)/365.25)
```

#### Location
Merge location other with trunk
```{r, eval=TRUE}
sel.train.data$Loc_CAT[sel.train.data$Loc_CAT=="other"] <- "trunk"
```

Merge LMM other with other
```{r, eval=TRUE}
sel.train.data$Histology[sel.train.data$Histology=="LMM"] <- "other"
```

#### Define multiple fields
```{r, eval=TRUE}
sel.train.data$multiple.fields <- ifelse(!is.na(sel.train.data$nr_SNs_Field_2)&sel.train.data$nr_SNs_Field_2>0,1,0)
```

#### Number of positive and negative nodes
Check: there are no negative patients that have positive nodes
```{r, eval=TRUE}
sum(!is.na(sel.train.data$Tot_SNs_pos)&sel.train.data$Tot_SNs_pos>0&sel.train.data$SNstatus=="Negative")
```

Create number of negative nodes removed
```{r, eval=TRUE}
sel.train.data$Tot_SNs_neg <- sel.train.data$Tot_nr_SNs - sel.train.data$Tot_SNs_pos
```

467 positive SN patients have negative nodes
```{r, eval=TRUE}
sum(!is.na(sel.train.data$Tot_SNs_neg)&sel.train.data$Tot_SNs_neg>0&sel.train.data$SNstatus=="Positive")
```

#### SN tumour burden
11 patients with positive SN have SN tumour burden of exactly 1
```{r, eval=TRUE}
sum(sel.train.data$Rdamcrit==1&sel.train.data$SNstatus=="Positive"&!is.na(sel.train.data$Rdamcrit))
```

Set Rdamcrit of negative SN patients equal to 0
```{r, eval=TRUE}
sel.train.data$Rdamcrit[sel.train.data$SNstatus=="Negative"] <- 1 # since log(1) = 0
```

#### Location of metastasis
Set Dewar for negative patients to subcap
```{r, eval=TRUE}
sum(sel.train.data$SNstatus=="Positive"&sel.train.data$Dewar=="subcap"&!is.na(sel.train.data$Dewar))
sel.train.data$Dewar[sel.train.data$SNstatus=="Negative"] <- "subcap"
```

Merge single cells with subcap 
```{r, eval=TRUE}
sel.train.data$Dewar[sel.train.data$Dewar=="single cells"] <- "subcap"
```

#### Recurrence 
Define time since diagnosis in years (follow-up time)
```{r, eval=TRUE}
sel.train.data$FU.time <- (sel.train.data$Last_FU-sel.train.data$SNdate)/(24*60*60)/365.25
```

Define recurrence time in years
```{r, eval=TRUE}
sel.train.data$Rec.time <- (sel.train.data$Date_1st_Rec-sel.train.data$SNdate)/(24*60*60)/365.25
```

Omit five patients with Rec date before SN date
```{r, eval=TRUE}
old.n <- nrow(sel.train.data)
sel.train.data <- sel.train.data[!(sel.train.data$Rec.time<0&!is.na(sel.train.data$Rec.time)),]
cat("Number of patients excluded:", old.n-nrow(sel.train.data), "\n")
```

Set recurrence time equal to follow-up time for those without recurrence
```{r, eval=TRUE}
sel.train.data$Rec.time[sel.train.data$Recurrence=="no recurrence"|is.na(sel.train.data$SNdate)]<-sel.train.data$FU.time[sel.train.data$Recurrence=="no recurrence"|is.na(sel.train.data$SNdate)]
table(sel.train.data$Status, sel.train.data$Recurrence)
```

Impute recurrence time for 45 patients with missing Rec.time but Recurrence=="recurred"
```{r, eval=TRUE}
cat('Number of patients dead with recurrence without Date_1st_Rec:', sum(is.na(sel.train.data$Date_1st_Rec)&sel.train.data$Recurrence=="recurred"&sel.train.data$dead=="Yes"), '\n')
cat('Number of patients alive with recurrence without Date_1st_Rec:', sum(is.na(sel.train.data$Date_1st_Rec)&sel.train.data$Recurrence=="recurred"&sel.train.data$dead=="No"))
```

#### Melanoma specific mortality (MSM)
```{r, eval=TRUE}
# 504 patients died due to melanoma, 102 patients died from other causes
table(sel.train.data$Status, sel.train.data$dead)
sel.train.data$MSM<-1*(sel.train.data$Status=="DOD")
```

#### Mitosis
Set Mitosis to Unknown if missing
```{r, eval=TRUE}
sel.train.data$Mitosis[is.na(sel.train.data$Mitosis)] <- "Unknown" # mitosis unknown for 80% of patients
```

#### Replace "Unknown" by NA
```{r, eval=TRUE}
sel.train.data$Ulceration[sel.train.data$Ulceration=="unknown"]<-NA
sel.train.data$Loc_CAT[sel.train.data$Loc_CAT=="unknown"]<-NA
sel.train.data$Histology[sel.train.data$Histology=="unknown"]<-NA
sel.train.data$Dewar[sel.train.data$Dewar=="unknown"] <- NA
sel.train.data$Mitosis[sel.train.data$Mitosis=="Unknown"] <- NA
```

#### Delete unused levels
```{r, eval=TRUE}
sel.train.data <- droplevels(sel.train.data)

# set levels for multiple.fields
sel.train.data$multiple.fields <- factor(sel.train.data$multiple.fields, labels=c("Other", "Multiple SN fields"))
sel.train.data$MSM <- factor(sel.train.data$MSM, labels=c("Other", "Death by melanoma"))
```

# Multiple imputation
```{r, eval=TRUE}
variables.to.be.imputed <- c("Center", "SNstatus", "Sex", "Age.SN", "Ulceration",
                             "Loc_CAT", "Histology", "Breslow", "multiple.fields",
                             "Tot_SNs_neg", "Tot_SNs_pos", "Rdamcrit", "Dewar",
                             "Mitosis",
                             "Recurrence", "dead", "MSM")
training.data.to.be.imputed <- sel.train.data[, variables.to.be.imputed]
```

Save distributions of all variables, necessary for Harrell summaries
```{r, eval=TRUE}
dd.training<-rms::datadist(training.data.to.be.imputed)
options(datadist='dd.training')
options(digits=8)
```

Relation between outcome and covariates
```{r, eval=TRUE}
training.data.to.be.imputed$logRec.time <- log(sel.train.data$Rec.time+1/365.25)
training.data.to.be.imputed$logFU.time <- log(sel.train.data$FU.time+1/365.25)
missings.table(training.data.to.be.imputed)
save(list=c("dd.training", "training.data.to.be.imputed"),file='Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.Rdata', compress=TRUE)
```