### 2022-11-29 Code/0_PrepareValidationData.Rmd

---
title: "Prepare Validation Data"
author: "Carolien C.H.M. Maas"
date: "14-11-2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Validation data preparation

### Load packages, functions, and data 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# erase memory
rm(list = ls(all.names = TRUE))

# for reproducibility
set.seed(100)

# load libraries
library(foreign)
library(rms)
source("Z:/Project Melanoom/PaperMelanoma/Code/PaperMelanomaFunctions.R")

# load data file
original.validation.data <- foreign::read.spss(file="Z:/Project Melanoom/PaperMelanoma/Data/validation/Validation_cohort_MIA.sav",
                                               to.data.frame=TRUE, origin="01-01-1970")

# remove space after Negative # Bespreken met Robert
levels(original.validation.data$SNstatus)[levels(original.validation.data$SNstatus)=="Negative "] <- "Negative"

# all patients may be included
sel.val.data <- original.validation.data
```

### Recode
#### Use age at SN instead of age at diagnosis
```{r, eval=TRUE}
sel.val.data$Age.SN <- floor((sel.val.data$SNdate-sel.val.data$DOB)/(24*60*60)/365.25)
```

#### Location
Merge location other with trunk
```{r, eval=TRUE}
sel.val.data$Loc_CAT[sel.val.data$Loc_CAT=="other"] <- "trunk"
```

Merge LMM other with other
```{r, eval=TRUE}
sel.val.data$Histology[sel.val.data$Histology=="LLM"] <- "other" # Bespreken met Robert, is LLM and LMM the same?
```

#### Define multiple fields
```{r, eval=TRUE}
sel.val.data$multiple.fields <- as.numeric(!is.na(sel.val.data$Tot_nr_SNs)&sel.val.data$Tot_nr_SNs>1) # Bespreken met ask Robert
```

#### Number of positive and negative nodes
Check: there are no negative patients that have positive nodes
```{r, eval=TRUE}
sum(!is.na(sel.val.data$Tot_SNs_pos)&sel.val.data$Tot_SNs_pos>0&sel.val.data$SNstatus=="Negative")
```

Create number of negative nodes removed
```{r, eval=TRUE}
sel.val.data$Tot_nr_SNs <- sel.val.data$Tot_SNs_neg + sel.val.data$Tot_SNs_pos
```

607 positive SN patients have negative nodes
```{r, eval=TRUE}
sum(!is.na(sel.val.data$Tot_SNs_neg)&sel.val.data$Tot_SNs_neg>0&sel.val.data$SNstatus=="Positive")
```

#### SN tumour burden
43 patients with positive SN have SN tumour burden of exactly 1
```{r, eval=TRUE}
sum(sel.val.data$Rdamcrit==1&sel.val.data$SNstatus=="Positive"&!is.na(sel.val.data$Rdamcrit))
```

Set Rdamcrit of negative SN patients equal to 0, SNstatus will capture the effect of these patients
```{r, eval=TRUE}
sel.val.data$Rdamcrit[sel.val.data$SNstatus=="Negative"] <- 1 # since log(1) = 0
```

#### Location of metastasis
Set Dewar for negative patients to subcap to merge them with the reference category, SN status will capture the effect of Negative patients
```{r, eval=TRUE}
sel.val.data$Dewar <- sel.val.data$Subcap
sel.val.data$Dewar[sel.val.data$SNstatus=="Negative"] <- "subcap"

# 339 patients have subcap with positive SN
sum(sel.val.data$SNstatus=="Positive"&sel.val.data$Dewar=="subcap"&!is.na(sel.val.data$Dewar))
```
#### Recurrence 
Define time since diagnosis in years (follow-up time)
```{r, eval=TRUE}
sel.val.data$FU.time <- (sel.val.data$Last_FU-sel.val.data$SNdate)/(24*60*60)/365.25
```

Define recurrence time in years
```{r, eval=TRUE}
sel.val.data$Rec.time <- (sel.val.data$Date_1st_Rec-sel.val.data$SNdate)/(24*60*60)/365.25
```

Check: no patients with Rec date before SN date
```{r, eval=TRUE}
sum(!(!(sel.val.data$Rec.time<0&!is.na(sel.val.data$Rec.time))))
```

Impute recurrence time for 45 patients with missing Rec.time but Recurrence=="recurred"
```{r, eval=TRUE}
cat('Number of patients dead with recurrence without Date_1st_Rec:', sum(is.na(sel.val.data$Date_1st_Rec)&sel.val.data$Status=="Dead, melanoma       "), '\n')
cat('Number of patients alive with recurrence without Date_1st_Rec:', sum(is.na(sel.val.data$Date_1st_Rec)&sel.val.data$Status=="Alive with Melanoma  "))
```

Set recurrence indicator
```{r, eval=TRUE}
Rec.date <- as.numeric(!is.na(sel.val.data$Date_1st_Rec)) # n=1294
Rec.date <- factor(Rec.date, labels=c("no recurrence", "recurred"))
table(sel.val.data$Status, Rec.date) # Bespreken met Robert

cat("Number of alive patients with Date_1st_Rec, but status NSR:", sum(!is.na(sel.val.data$Date_1st_Rec)&sel.val.data$Status=="Alive NSR            "), "\n")
cat("Number of alive patients with Date_1st_Rec, but status unknown:", sum(!is.na(sel.val.data$Date_1st_Rec)&sel.val.data$Status=="Alive, status unknown"), "\n")
cat("Number of dead patients with Date_1st_Rec, but status unknown:", sum(!is.na(sel.val.data$Date_1st_Rec)&sel.val.data$Status=="Dead, cause unknown  "), "\n")
cat("Number of dead patients with Date_1st_Rec, but status not melanoma:", sum(!is.na(sel.val.data$Date_1st_Rec)&sel.val.data$Status=="Dead, not melanoma   "), "\n")

# Recurrence if patient has Date_1st_Rec, Alive with melanoma, or dead with melanoma
sel.val.data$Recurrence <- as.numeric(sel.val.data$Status=="Alive with Melanoma  " | sel.val.data$Status=="Dead, melanoma       " | !is.na(sel.val.data$Date_1st_Rec)) # n=1352
sum(sel.val.data$Recurrence)
```

#### Melanoma specific mortality (MSM)
```{r, eval=TRUE}
sel.val.data$dead <- as.numeric(stringr::str_detect(sel.val.data$Status, "Dead"))

# 504 patients died due to melanoma, 102 patients died from other causes
table(sel.val.data$Status, sel.val.data$dead)

sel.val.data$MSM <- as.numeric(sel.val.data$Status=="Dead, melanoma       ")
```
#### Mitosis
Define Mitosis as Yes/No
```{r, eval=TRUE}
sel.val.data$Mitosis <- as.numeric(!is.na(sel.val.data$Mitosis))
```

#### Delete unused levels
```{r, eval=TRUE}
sel.val.data <- droplevels(sel.val.data)
sel.val.data$Center <- "Australia"
```

#### Set levels
```{r, eval=TRUE}
sel.val.data$Ulceration <- factor(sel.val.data$Ulceration, labels=c("No", "Yes"))
sel.val.data$Mitosis <- factor(sel.val.data$Mitosis, labels=c("Yes", "No"))

sel.val.data$multiple.fields <- factor(sel.val.data$multiple.fields, labels=c("Multiple SN fields", "Other"))

sel.val.data$Recurrence <- factor(sel.val.data$Recurrence, labels=c("no recurrence", "recurred"))
sel.val.data$dead <- factor(sel.val.data$dead, labels=c("No", "Yes"))
sel.val.data$MSM <- factor(sel.val.data$MSM, labels=c("Other", "Death by melanoma"))
```

# Multiple imputation
```{r, eval=TRUE}
variables.to.be.imputed <- c("Center", "SNstatus", "Sex", "Age.SN", "Ulceration",
                             "Loc_CAT", "Histology", "Breslow", "multiple.fields",
                             "Tot_SNs_neg", "Tot_SNs_pos", "Rdamcrit", "Dewar",
                             "Mitosis",
                             "Recurrence", "dead", "MSM")
validation.data.to.be.imputed <- sel.val.data[, variables.to.be.imputed]
```

Save distributions of all variables, necessary for Harrell summaries
```{r, eval=TRUE}
dd.validation<-rms::datadist(validation.data.to.be.imputed)
options(datadist='dd.validation')
options(digits=8)
```

Relation between outcome and covariates
```{r, eval=TRUE}
validation.data.to.be.imputed$logRec.time <- log(sel.val.data$Rec.time+1/365.25)
validation.data.to.be.imputed$logFU.time <- log(sel.val.data$FU.time+1/365.25)
missings.table(validation.data.to.be.imputed)
save(list=c("dd.validation", "validation.data.to.be.imputed"),file='Z:/Project Melanoom/PaperMelanoma/Data/validation/validation.data.Rdata', compress=TRUE)
```