### 2022-11-29 Code/12_Validation.Rmd

---
title: "Create histogram with risk scores"
author: "Carolien C.H.M. Maas"
date: "14-11-2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Create calibration plots

### Load packages, functions, and data 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# erase memory
rm(list = ls(all.names = TRUE))

# for reproducibility
set.seed(100)

# load libraries
library(mice)
library(survival)         # needed to use
# remotes::install_github("CHMMaas/PredictionTools")
library(PredictionTools)  # to plot calibration
# import helper functions
source("Z:/Project Melanoom/PaperMelanoma/Code/PaperMelanomaFunctions.R")
```

# Validation
# Compute predictions for each multiple imputation and make calibration plots with metrics
```{r, eval=TRUE}
# laod data
load('Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.RData')
options(datadist='dd.training')
options(digits=8)
load('Z:/Project Melanoom/PaperMelanoma/Data/validation/validation.data.RData')
# load model
load('Z:/Project Melanoom/PaperMelanoma/Results/h0.Rec.MSM.Rdata')
p.Rec.val <- matrix(NA, nrow(validation.data.to.be.imputed), m)
p.MSM.val <- matrix(NA, nrow(validation.data.to.be.imputed), m)
for (i in 1:m){
  # Linear predictor
  lp.val.i <- predict(f.mi.BS.Rec.5,
                      newdata=mice::complete(imputed.validation.data, i),
                      type="lp")

  # Recurrence
  p.Rec.val[, i] <- 1-exp(-h0.Rec*exp(lp.val.i))

  # MSM
  p.MSM.val[, i] <- 1-exp(-h0.MSM*exp(MSM.cal.fact*lp.val.i))
}

# calibration curve for recurrence, TODO: based on dd.training?
png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/val.Rec.png"))
PredictionTools::val.surv.mi(p.Rec.val,S.Rec.5.validation,g=5,main="Calibration plot of recurrence for all patients",
                             time=NULL,lim=c(0,1),dist=TRUE)
grDevices::dev.off()

# calibration curve for MSM
png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/val.MSM.png"))
PredictionTools::val.surv.mi(p.MSM.val,S.MSM.5.validation,g=5,main="Calibration plot of MSM for all patients",
                             time=NULL,lim=c(0,1),dist=TRUE)
grDevices::dev.off()

```