### 2022-11-29 Code/2_ImputeData.Rmd

---
title: "Impute missing values of training and validation data using multiple imputation"
author: "Carolien C.H.M. Maas"
date: "14-11-2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Multiple imputation

### Load packages, functions, and data 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# erase memory
rm(list = ls(all.names = TRUE))

# for reproducibility
set.seed(100)

# load libraries
library(mice)
```

Function to perform multiple imputation for training or validation data
```{r, eval=TRUE}
impute.missing.values <- function(type=NULL, m=1, new.imputation=FALSE){
  # load data
  load(paste0('Z:/Project Melanoom/PaperMelanoma/Data/', type, '/', type, '.data.Rdata'))

  if (new.imputation){
    # impute data
    data.to.be.imputed <- eval(parse(text=paste0(type, '.data.to.be.imputed')))
    imputed.data <- mice::mice(data.to.be.imputed, m=m, maxit=5, print=FALSE, seed=500)
    assign(paste0('imputed.', type, '.data'), imputed.data)
    save(list=c(paste0('imputed.', type, '.data')),
         file=paste0('Z:/Project Melanoom/PaperMelanoma/Data/', type, '/imputed.', type, '.data.Rdata'), compress=TRUE)
  }
  load(paste0('Z:/Project Melanoom/PaperMelanoma/Data/', type, '/imputed.', type, '.data.Rdata'))
  save(list=c(paste0('dd.', type), paste0(type, '.data.to.be.imputed'), 'm', paste0('imputed.', type, '.data')),
       file=paste0('Z:/Project Melanoom/PaperMelanoma/Data/', type, '/', type, '.data.Rdata'), compress=TRUE)
}
```

Perform multiple imputation
```{r, eval=TRUE}
# erase memory
m <- 5
new.imputation <- FALSE
impute.missing.values(type='training', m=m, new.imputation=new.imputation)
impute.missing.values(type='validation', m=m, new.imputation=new.imputation)

# training data gives warning messages for Mitosis
# validation data gives multicolinearity of logRectime and recurrence
# print(imputed.data$loggedEvents)
```