### 2022-11-29 Code/3_Descriptives.Rmd

---
title: "Multiple imputation"
author: "Carolien C.H.M. Maas"
date: "14-11-2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Describe data

### Load packages, functions, and data 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# erase memory
rm(list = ls(all.names = TRUE))

# for reproducibility
set.seed(100)

# load libraries
library(mice)
library(survival)
library(survminer)
library(stats)
library(prodlim)
library(utils)
source("Z:/Project Melanoom/PaperMelanoma/Code/PaperMelanomaFunctions.R")
```

```{r, eval=TRUE}
prepare.data <- function(type){
  # Load imputed data set 
  load(paste0('Z:/Project Melanoom/PaperMelanoma/Data/', type, '/', type, '.data.RData'))
  imputed.data <- eval(parse(text=paste0("imputed.", type, ".data")))
  last.imputed.data.set <- mice::complete(imputed.data, m)

  assign(paste0('S.Rec.', type), survival::Surv(exp(last.imputed.data.set$logRec.time), last.imputed.data.set$Recurrence=="recurred"))
  assign(paste0('S.MSM.', type), survival::Surv(exp(last.imputed.data.set$logFU.time), last.imputed.data.set$MSM=="Death by melanoma"))

  assign(paste0("imputed.", type, ".data"), imputed.data)
  assign(paste0("last.imputed.", type, ".data.set"), last.imputed.data.set)

  save(list=c(paste0("dd.", type), "m", 
              paste0(type, ".data.to.be.imputed"), 
              paste0("imputed.", type, ".data"), 
              paste0("last.imputed.", type, ".data.set"), 
              paste0("S.Rec.", type), 
              paste0("S.MSM.", type)),
       file=paste0('Z:/Project Melanoom/PaperMelanoma/Data/', type, '/', type, '.data.Rdata'), compress=TRUE)
}
prepare.data(type='training')
prepare.data(type='validation')
```

### Kaplan-Meier plots
```{r, eval=TRUE}
load(paste0('Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.RData'))
survminer::ggsurvplot(survival::survfit(S.Rec.training ~ Center, data=last.imputed.training.data.set), conf.int=TRUE, risk.table=TRUE)
survminer::ggsurvplot(survival::survfit(S.MSM.training ~ Center, data=last.imputed.training.data.set), conf.int=TRUE, risk.table=TRUE)
load(paste0('Z:/Project Melanoom/PaperMelanoma/Data/validation/validation.data.RData'))
survminer::ggsurvplot(survival::survfit(S.Rec.validation ~ Center, data=last.imputed.validation.data.set), conf.int=TRUE, risk.table=TRUE)
survminer::ggsurvplot(survival::survfit(S.MSM.validation ~ Center, data=last.imputed.validation.data.set), conf.int=TRUE, risk.table=TRUE)
```

# Median follow-up time
```{r, eval=TRUE}
median.FU.time <- function(data=NA){
  median.FU.time.total <- stats::quantile(prodlim::prodlim(prodlim::Hist(exp(logFU.time), MSM=="Death by melanoma")~1, data=data[data$Center!="Australia",], reverse=TRUE))
  tab.FU.time <- c("Total Development", paste(sprintf("%.2f", median.FU.time.total$quantiles.survival[3, "quantile"]),
                   paste0("[", sprintf("%.2f", median.FU.time.total$quantiles.survival[3, "lower"]), "; ",
                          sprintf("%.2f", median.FU.time.total$quantiles.survival[3, "upper"]), "]")))
  median.FU.time <- stats::quantile(prodlim::prodlim(prodlim::Hist(exp(logFU.time), MSM=="Death by melanoma")~Center, data=data, reverse=TRUE))
  for (center in c("Amsterdam", "Berlin", "Rotterdam", "Warsaw", "Australia")){
    med.FU.table <- eval(parse(text=paste0("median.FU.time$`Center=", center, "`")))
    med.FU <- med.FU.table[3,"quantile"]
    lower.FU <- med.FU.table[3,"lower"]
    upper.FU <- med.FU.table[3,"upper"]
    tab.FU.time <- rbind(tab.FU.time, c(center, paste(sprintf("%.2f", med.FU), paste0("[", sprintf("%.2f", lower.FU), "; ", sprintf("%.2f", upper.FU), "]"))))
  }
  write.table(as.data.frame(tab.FU.time),
              file=paste0('Z:/Project Melanoom/PaperMelanoma/Results/med.FU.table.txt'),
              row.names=FALSE, sep=",")
  return(tab.FU.time)
}
median.FU.time(data=rbind(training.data.to.be.imputed[, c("logFU.time", "MSM", "Center")], 
                          validation.data.to.be.imputed[, c("logFU.time", "MSM", "Center")]))
```

# Recurrence at 5 years
```{r, eval=TRUE}
Rec.MSM.table <- function(horizon=5, type=NULL, data.to.be.imputed=NULL){
  # all patients, Recurrence
  S.Rec <- eval(parse(text=paste0('S.Rec.', type)))
  SF.Rec.all<-survfit(S.Rec~1, data=data.to.be.imputed)
  Rec.all <- 100*(1-SF.Rec.all$surv[SF.Rec.all$time==max(SF.Rec.all$time[SF.Rec.all$time<=horizon])])

  # only negative patients, Recurrence
  SF.Rec.neg<-survfit(S.Rec[data.to.be.imputed$SNstatus=="Negative"]~1, data=data.to.be.imputed)
  Rec.neg <- 100*(1-SF.Rec.neg$surv[SF.Rec.neg$time==max(SF.Rec.neg$time[SF.Rec.neg$time<=horizon])])

  # only positive patients, Recurrence
  SF.Rec.pos<-survfit(S.Rec[data.to.be.imputed$SNstatus=="Positive"]~1, data=data.to.be.imputed)
  Rec.pos <- 100*(1-SF.Rec.pos$surv[SF.Rec.pos$time==max(SF.Rec.pos$time[SF.Rec.pos$time<=horizon])])

  # all patients, MSM
  S.MSM <- eval(parse(text=paste0('S.MSM.', type)))
  SF.MSM.all<-survfit(S.MSM~1, data=data.to.be.imputed)
  MSM.all <- 100*(1-SF.MSM.all$surv[SF.MSM.all$time==max(SF.MSM.all$time[SF.MSM.all$time<=horizon])])

  # negative patients, MSM
  SF.MSM.neg<-survfit(S.MSM[data.to.be.imputed$SNstatus=="Negative"]~1, data=data.to.be.imputed)
  MSM.neg <- 100*(1-SF.MSM.neg$surv[SF.MSM.neg$time==max(SF.MSM.neg$time[SF.MSM.neg$time<=horizon])])

  # positive patients, MSM
  SF.MSM.pos<-survfit(S.MSM[data.to.be.imputed$SNstatus=="Positive"]~1, data=data.to.be.imputed)
  MSM.pos <- 100*(1-SF.MSM.pos$surv[SF.MSM.pos$time==max(SF.MSM.pos$time[SF.MSM.pos$time<=horizon])])

  # merge in one table
  Rec.MSM.table <- data.frame(n=c(nrow(data.to.be.imputed),
                                  sum(data.to.be.imputed$SNstatus=="Negative"),
                                  sum(data.to.be.imputed$SNstatus=="Positive")),
                              perc=c(100,
                                     sprintf("%.1f", sum(data.to.be.imputed$SNstatus=="Negative")/nrow(data.to.be.imputed)*100),
                                     sprintf("%.1f", sum(data.to.be.imputed$SNstatus=="Positive")/nrow(data.to.be.imputed)*100)),
                              Rec=sprintf("%.1f", c(Rec.all, Rec.neg, Rec.pos)),
                              MSM=sprintf("%.1f", c(MSM.all, MSM.neg, MSM.pos)))
  row.names(Rec.MSM.table) <- c("All", "Negative", "Positive")
  utils::write.table(as.data.frame(Rec.MSM.table),file=paste0('Z:/Project Melanoom/PaperMelanoma/Results/Rec.MSM.table.', type, '.txt'),
                     row.names=TRUE, sep=",")
  return(Rec.MSM.table)
}

horizon <- 5
Rec.MSM.table(horizon=horizon, type='training', data.to.be.imputed=training.data.to.be.imputed)
Rec.MSM.table(horizon=horizon, type='validation', data.to.be.imputed=validation.data.to.be.imputed)
```

# Restrict survival at 5 years
```{r, eval=TRUE}
for (type in c('training',  'validation')){
  load(paste0('Z:/Project Melanoom/PaperMelanoma/Data/', type, '/', type, '.data.Rdata'))
  for (outcome in c('Rec', 'MSM')){
    S <- eval(parse(text=paste0('S.', outcome, '.', type)))
    S.5 <- S
    S.5[S[,1]>5, 1] <- 5 # set time to 5
    S.5[S[,1]>5, 2] <- 0 # set censoring

    assign(paste0('S.', outcome, '.5.', type), S.5)
  }
  save(list=c(paste0("dd.", type), "m", "horizon", 
              paste0(type, ".data.to.be.imputed"),
              paste0("imputed.", type, ".data"), 
              paste0("last.imputed.", type, ".data.set"),
              paste0("S.Rec.", type), 
              paste0("S.MSM.", type),
              paste0("S.Rec.5.", type), 
              paste0("S.MSM.5.", type)),
     file=paste0('Z:/Project Melanoom/PaperMelanoma/Data/', type, '/', type, '.data.Rdata'), compress=TRUE)
}
```