### 2022-11-29 Code/4_AJCC.Rmd

---
title: "Calculate AJCC"
author: "Carolien C.H.M. Maas"
date: "14-11-2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Calculate AJCC

### Load packages, functions, and data 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# erase memory
rm(list = ls(all.names = TRUE))

# for reproducibility
set.seed(100)

# load libraries
# remotes::install_github("CHMMaas/PredictionTools")
library(PredictionTools)  # to plot calibration
```
# Obtain C-index of AJCC7 and AJCC8
```{r, eval=TRUE}
calculate.AJCC.7.8 <- function(imputed.data=NULL, type=NULL, S.Rec.5=NULL, S.MSM.5=NULL){
  ####
  #### AJCC 7
  ####
  TNM.7 <- c()
  AJCC.7 <- c()
  cindex.AJCC.7 <- rep(0, m)
  cindex.AJCC.7.se <- rep(0, m)
  cindex.AJCC.7.pos <- rep(0, m)
  cindex.AJCC.7.se.pos <- rep(0, m)
  cindex.AJCC.7.neg <- rep(0, m)
  cindex.AJCC.7.se.neg <- rep(0, m)
  for (i in 1:m){
    imputation.i <- mice::complete(imputed.data,i)
    TNM.7.i <- rep("", nrow(imputation.i))
    AJCC.7.i <- rep(0, nrow(imputation.i))
    for (patient in 1:nrow(imputation.i)){
      breslow.patient <- imputation.i[patient, "Breslow"]
      ulceration.patient <- as.character(imputation.i[patient, "Ulceration"])
      mitosis.patient <- as.character(imputation.i[patient, "Mitosis"])
      nodes.patient <- imputation.i[patient, "Tot_SNs_pos"]

      snstatus.patient <- imputation.i[patient, "SNstatus"]

      # T stage
      if (breslow.patient <= 1){
        if (ulceration.patient=="No" & mitosis.patient=="No"){
          TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T1a")
        } else if (ulceration.patient=="Yes" | mitosis.patient=="Yes"){
          TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T1b")
        }
      } else if (breslow.patient>1 & breslow.patient <= 2){
        if (ulceration.patient=="No"){
          TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T2a")
        } else if (ulceration.patient=="Yes"){
          TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T2b")
        }
      } else if (breslow.patient>2 & breslow.patient <= 4){
        if (ulceration.patient=="No"){
          TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T3a")
        } else if (ulceration.patient=="Yes"){
          TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T3b")
        }
      } else if (breslow.patient>4){
        if (ulceration.patient=="No"){
          TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T4a")
        } else if (ulceration.patient=="Yes"){
          TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T4b")
        }
      }

      # N stage
      if (nodes.patient == 0){
        TNM.7.i[patient] <- paste0(TNM.7.i[patient], "N0")
      } else if (nodes.patient==1){
        TNM.7.i[patient] <- paste0(TNM.7.i[patient], "N1a")
      } else if (nodes.patient==2 | nodes.patient==3){
        TNM.7.i[patient] <- paste0(TNM.7.i[patient], "N2a")
      } else if (nodes.patient>3){
        TNM.7.i[patient] <- paste0(TNM.7.i[patient], "N3a")
      }

      # M stage for all patients in our data set
      TNM.7.i[patient] <- paste0(TNM.7.i[patient], "M0")

      # determine AJCC.7.i
      if (TNM.7.i[patient]=="T1aN0M0"){
        AJCC.7.i[patient] <- 1.1
      } else if (TNM.7.i[patient]=="T1bN0M0" | TNM.7.i[patient]=="T2aN0M0"){
        AJCC.7.i[patient] <- 1.2
      } else if (TNM.7.i[patient]=="T2bN0M0" | TNM.7.i[patient]=="T3aN0M0"){
        AJCC.7.i[patient] <- 2.1
      } else if (TNM.7.i[patient]=="T3bN0M0" | TNM.7.i[patient]=="T4aN0M0"){
        AJCC.7.i[patient] <- 2.2
      } else if (TNM.7.i[patient]=="T4bN0M0"){
        AJCC.7.i[patient] <- 2.3
      } else if (TNM.7.i[patient]=="T1aN1aM0" | TNM.7.i[patient]=="T2aN1aM0" | # N1a
                 TNM.7.i[patient]=="T3aN1aM0" | TNM.7.i[patient]=="T4aN1aM0" |
                 TNM.7.i[patient]=="T1aN2aM0" | TNM.7.i[patient]=="T2aN2aM0" | # N2a
                 TNM.7.i[patient]=="T3aN2aM0" | TNM.7.i[patient]=="T4aN2aM0"){
        AJCC.7.i[patient] <- 3.1
      } else if (TNM.7.i[patient]=="T1bN1aM0" | TNM.7.i[patient]=="T2bN1aM0" | # N1a
                 TNM.7.i[patient]=="T3bN1aM0" | TNM.7.i[patient]=="T4bN1aM0" |
                 TNM.7.i[patient]=="T1bN2aM0" | TNM.7.i[patient]=="T2bN2aM0" | # N2a
                 TNM.7.i[patient]=="T3bN2aM0" | TNM.7.i[patient]=="T4bN2aM0" |
                 TNM.7.i[patient]=="T1aN1bM0" | TNM.7.i[patient]=="T2aN1bM0" | # N1b
                 TNM.7.i[patient]=="T3aN1bM0" | TNM.7.i[patient]=="T4aN1bM0" |
                 TNM.7.i[patient]=="T1aN2bM0" | TNM.7.i[patient]=="T2aN2bM0" | # N2b
                 TNM.7.i[patient]=="T3aN2bM0" | TNM.7.i[patient]=="T4aN2bM0" |
                 TNM.7.i[patient]=="T1aN2cM0" | TNM.7.i[patient]=="T2aN2cM0" | # N2c
                 TNM.7.i[patient]=="T3aN2cM0" | TNM.7.i[patient]=="T4aN2cM0"){
        AJCC.7.i[patient] <- 3.2
      } else {
        AJCC.7.i[patient] <- 3.3
      }
    }

    # save TNM.7 and AJCC.7 for each imputated data set
    TNM.7 <- cbind(TNM.7, TNM.7.i)
    AJCC.7 <- cbind(AJCC.7, AJCC.7.i)

    # compute C-index
    rc <- Hmisc::rcorr.cens(-AJCC.7.i, S.Rec.5) # add a minus to revert class order: from worst to best
    cindex.AJCC.7[i]<-rc["C Index"]
    cindex.AJCC.7.se[i]<-rc["S.D."]/2

    # compute C-index for only positive sentinel node patients
    rc.pos <- Hmisc::rcorr.cens(-AJCC.7.i[imputation.i$SNstatus=="Positive"], S.Rec.5[imputation.i$SNstatus=="Positive",])
    cindex.AJCC.7.pos[i]<-rc.pos["C Index"]
    cindex.AJCC.7.se.pos[i]<-rc.pos["S.D."]/2

    # compute C-index for only negative sentinel node patients
    rc.neg <- Hmisc::rcorr.cens(-AJCC.7.i[imputation.i$SNstatus=="Negative"], S.Rec.5[imputation.i$SNstatus=="Negative",])
    cindex.AJCC.7.neg[i]<-rc.neg["C Index"]
    cindex.AJCC.7.se.neg[i]<-rc.neg["S.D."]/2
  }
  # combine C-index for each imputed data set
  cindex.mi.AJCC.7<-Rubin.combine(cindex.AJCC.7,cindex.AJCC.7.se)
  cindex.mi.AJCC.7.pos<-Rubin.combine(cindex.AJCC.7.pos,cindex.AJCC.7.se.pos)
  cindex.mi.AJCC.7.neg<-Rubin.combine(cindex.AJCC.7.neg,cindex.AJCC.7.se.neg)

  # save C-index and confidence interval
  utils::write.table(rbind(c("All patients:", sprintf('%.2f', cindex.mi.AJCC.7$est),
                paste0("[", sprintf('%.2f', cindex.mi.AJCC.7$est+qnorm(.025)*cindex.mi.AJCC.7$se),
                        "; ", sprintf('%.2f', cindex.mi.AJCC.7$est+qnorm(.975)*cindex.mi.AJCC.7$se), "]")),
              c("Positive SN patients:", sprintf('%.2f', cindex.mi.AJCC.7.pos$est),
                    paste0("[", sprintf('%.2f', cindex.mi.AJCC.7.pos$est+qnorm(.025)*cindex.mi.AJCC.7.pos$se),
                           "; ", sprintf('%.2f', cindex.mi.AJCC.7.pos$est+qnorm(.975)*cindex.mi.AJCC.7.pos$se), "]")),
              c("Negative SN patients:", sprintf('%.2f', cindex.mi.AJCC.7.neg$est),
                    paste0("[", sprintf('%.2f', cindex.mi.AJCC.7.neg$est+qnorm(.025)*cindex.mi.AJCC.7.neg$se),
                           "; ", sprintf('%.2f', cindex.mi.AJCC.7.neg$est+qnorm(.975)*cindex.mi.AJCC.7.neg$se), "]"))),
              file=paste0('Z:/Project Melanoom/PaperMelanoma/Results/AJCC.7.Cindex.', type, '.txt'), 
              row.names=FALSE, col.names=FALSE, sep=",")

  ####
  #### AJCC8
  ####
  TNM.8 <- c()
  AJCC.8 <- c()
  cindex.AJCC.8 <- rep(0, m)
  cindex.AJCC.8.se <- rep(0, m)
  cindex.AJCC.8.pos <- rep(0, m)
  cindex.AJCC.8.se.pos <- rep(0, m)
  cindex.AJCC.8.neg <- rep(0, m)
  cindex.AJCC.8.se.neg <- rep(0, m)
  for (i in 1:m){
    imputation.i <- mice::complete(imputed.data,i)
    TNM.8.i <- rep("", nrow(imputation.i))
    AJCC.8.i <- rep(0, nrow(imputation.i))
    for (patient in 1:nrow(imputation.i)){
      breslow.patient <- imputation.i[patient, "Breslow"]
      ulceration.patient <- as.character(imputation.i[patient, "Ulceration"])
      mitosis.patient <- as.character(imputation.i[patient, "Mitosis"])
      nodes.patient <- imputation.i[patient, "Tot_SNs_pos"]

      # T stage
      if (breslow.patient <= 1){
        if (ulceration.patient=="No" & breslow.patient < 0.8){
          TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T1a")
        } else {
          TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T1b")
        }
      } else if (breslow.patient>1 & breslow.patient <= 2){
        if (ulceration.patient=="No"){
          TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T2a")
        } else if (ulceration.patient=="Yes"){
          TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T2b")
        }
      } else if (breslow.patient>2 & breslow.patient <= 4){
        if (ulceration.patient=="No"){
          TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T3a")
        } else if (ulceration.patient=="Yes"){
          TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T3b")
        }
      } else if (breslow.patient>4){
        if (ulceration.patient=="No"){
          TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T4a")
        } else if (ulceration.patient=="Yes"){
          TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T4b")
        }
      }

      # N stage
      if (nodes.patient == 0){
        TNM.8.i[patient] <- paste0(TNM.8.i[patient], "N0")
      } else if (nodes.patient==1){
        TNM.8.i[patient] <- paste0(TNM.8.i[patient], "N1a")
      } else if (nodes.patient==2 | nodes.patient==3){
        TNM.8.i[patient] <- paste0(TNM.8.i[patient], "N2a")
      } else if (nodes.patient>3){
        TNM.8.i[patient] <- paste0(TNM.8.i[patient], "N3a")
      }

      # M stage for all patients in our data set
      TNM.8.i[patient] <- paste0(TNM.8.i[patient], "M0")

      # determine AJCC.8.i
      if (TNM.8.i[patient]=="T1aN0M0"){
        AJCC.8.i[patient] <- 1.1
      } else if (TNM.8.i[patient]=="T1bN0M0" | TNM.8.i[patient]=="T2aN0M0"){
        AJCC.8.i[patient] <- 1.2
      } else if (TNM.8.i[patient]=="T2bN0M0" | TNM.8.i[patient]=="T3aN0M0"){
        AJCC.8.i[patient] <- 2.1
      } else if (TNM.8.i[patient]=="T3bN0M0" | TNM.8.i[patient]=="T4aN0M0"){
        AJCC.8.i[patient] <- 2.2
      } else if (TNM.8.i[patient]=="T4bN0M0"){
        AJCC.8.i[patient] <- 2.3
      } else if (TNM.8.i[patient]=="T1aN1aM0" | TNM.8.i[patient]=="T1bN1aM0" | TNM.8.i[patient]=="T2aN1aM0"
                 | TNM.8.i[patient]=="T1aN2aM0" | TNM.8.i[patient]=="T1bN2aM0" | TNM.8.i[patient]=="T2aN2aM0"){
        AJCC.8.i[patient] <- 3.1
      } else if (TNM.8.i[patient]=="T2bN1aM0" | TNM.8.i[patient]=="T3aN1aM0"){
        AJCC.8.i[patient] <- 3.2
      } else if (TNM.8.i[patient]=="T4bN3aM0"){
        AJCC.8.i[patient] <- 3.4
      } else {
        AJCC.8.i[patient] <- 3.3
      }

    }

    # save TNM.8 and AJCC.8 for each imputated data set
    TNM.8 <- cbind(TNM.8, TNM.8.i)
    AJCC.8 <- cbind(AJCC.8, AJCC.8.i)

    # compute C-index
    rc <- Hmisc::rcorr.cens(-AJCC.8.i, S.Rec.5) # add a minus to revert class order: from worst to best
    cindex.AJCC.8[i]<-rc["C Index"]
    cindex.AJCC.8.se[i]<-rc["S.D."]/2

    # compute C-index for only positive sentinel node patients
    rc.pos <- Hmisc::rcorr.cens(-AJCC.8.i[imputation.i$SNstatus=="Positive"], S.Rec.5[imputation.i$SNstatus=="Positive",])
    cindex.AJCC.8.pos[i]<-rc.pos["C Index"]
    cindex.AJCC.8.se.pos[i]<-rc.pos["S.D."]/2

    # compute C-index for only negative sentinel node patients
    rc.neg <- Hmisc::rcorr.cens(-AJCC.8.i[imputation.i$SNstatus=="Negative"], S.Rec.5[imputation.i$SNstatus=="Negative",])
    cindex.AJCC.8.neg[i]<-rc.neg["C Index"]
    cindex.AJCC.8.se.neg[i]<-rc.neg["S.D."]/2
  }
  # combine C-index for each imputed data set
  cindex.mi.AJCC.8<-Rubin.combine(cindex.AJCC.8,cindex.AJCC.8.se)
  cindex.mi.AJCC.8.pos<-Rubin.combine(cindex.AJCC.8.pos,cindex.AJCC.8.se.pos)
  cindex.mi.AJCC.8.neg<-Rubin.combine(cindex.AJCC.8.neg,cindex.AJCC.8.se.neg)

  # save C-index and confidence interval
  utils::write.table(rbind(c("All patients:", sprintf('%.2f', cindex.mi.AJCC.8$est),
                          paste0("[", sprintf('%.2f', cindex.mi.AJCC.8$est+qnorm(.025)*cindex.mi.AJCC.8$se),
                                 "; ", sprintf('%.2f', cindex.mi.AJCC.8$est+qnorm(.975)*cindex.mi.AJCC.8$se), "]")),
                    c("Positive SN patients:", sprintf('%.2f', cindex.mi.AJCC.8.pos$est),
                          paste0("[", sprintf('%.2f', cindex.mi.AJCC.8.pos$est+qnorm(.025)*cindex.mi.AJCC.8.pos$se),
                                 "; ", sprintf('%.2f', cindex.mi.AJCC.8.pos$est+qnorm(.975)*cindex.mi.AJCC.8.pos$se), "]")),
                    c("Negative SN patients", sprintf('%.2f', cindex.mi.AJCC.8.neg$est),
                          paste0("[", sprintf('%.2f', cindex.mi.AJCC.8.neg$est+qnorm(.025)*cindex.mi.AJCC.8.neg$se),
                                 "; ", sprintf('%.2f', cindex.mi.AJCC.8.neg$est+qnorm(.975)*cindex.mi.AJCC.8.neg$se), "]"))),
              file=paste0('Z:/Project Melanoom/PaperMelanoma/Results/AJCC.8.Cindex.', type, '.txt'), 
              row.names=FALSE, col.names=FALSE, sep=",")

  # table of number of AJCC.7 and AJCC.8
  table.AJCC.7 <- c()
  table.AJCC.8 <- c()
  for (i in 1:m){
    table.AJCC.7 <- cbind(table.AJCC.7, as.numeric(table(AJCC.7[,i])))
    table.AJCC.8 <- cbind(table.AJCC.8, as.numeric(table(AJCC.8[,i])))
  }

  # print(sort(unique(AJCC.7[,1])))
  # print(sort(unique(AJCC.8[,1])))

  if (type=="training"){
    table.AJCC.7.8 <- cbind(c('1.1', '1.2', '2.1', '2.2', '2.3', '3.1', '3.2', '3.3', '3.4'),
                          c(round(rowMeans(table.AJCC.7), 0), 0),
                          round(rowMeans(table.AJCC.8)))
  } else if (type=="validation"){
    table.AJCC.7.8 <- cbind(c('1.1', '1.2', '2.1', '2.2', '2.3', '3.1', '3.2', '3.3', '3.4'),
                          c(round(rowMeans(table.AJCC.7), 0), 0),
                          c(0, round(rowMeans(table.AJCC.8), 0)))
  }
  utils::write.table(table.AJCC.7.8,
              file=paste0('Z:/Project Melanoom/PaperMelanoma/Results/table.AJCC.7.and.8.', type, '.txt'),
              row.names=FALSE, col.names=FALSE, sep=",")
}
load('Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.RData')
calculate.AJCC.7.8(imputed.data=imputed.training.data, S.Rec.5=S.Rec.5.training, S.MSM.5=S.MSM.5.training, type='training')
load('Z:/Project Melanoom/PaperMelanoma/Data/validation/validation.data.RData')
calculate.AJCC.7.8(imputed.data=imputed.validation.data, S.Rec.5=S.Rec.5.validation, S.MSM.5=S.MSM.5.validation, type='validation')
```