### 2022-11-29 Code/6_FitModels.Rmd

---
title: "Fit models"
author: "Carolien C.H.M. Maas"
date: "14-11-2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fit models

### Load packages, functions, and data 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# erase memory
rm(list = ls(all.names = TRUE))

# for reproducibility
set.seed(100)

# load libraries
library(Hmisc)
library(stats)
library(rms)
library(utils)
```

# Full model for recurrence
```{r, eval=TRUE}
load('Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.RData')
options(datadist='dd.training')
options(digits=8)

form.full.Rec.5 <- S.Rec.5.training ~
  SNstatus +
  Sex +
  Age.SN +
  Ulceration +
  Loc_CAT +
  Histology +
  log(Breslow) +
  multiple.fields +
  Tot_SNs_neg +
  Tot_SNs_pos +
  log(Rdamcrit) +
  Dewar +
  SNstatus:Sex +
  SNstatus:Age.SN +
  SNstatus:Ulceration +
  SNstatus:Loc_CAT +
  SNstatus:Histology +
  SNstatus:log(Breslow) +
  SNstatus:multiple.fields
f.mi.full.Rec.5 <- Hmisc::fit.mult.impute(form.full.Rec.5, cph, xtrans=imputed.training.data,
                                 data=training.data.to.be.imputed,n.impute=m, pr=FALSE,
                                 fit.reps=TRUE, y=TRUE, x=TRUE, se.fit=TRUE)
stats::anova(f.mi.full.Rec.5)
summary(f.mi.full.Rec.5, Tot_SNs_pos=c(0, 1))
```

# Final model for recurrence selected by backward selection with p = 0.01
```{r, eval=TRUE}
c.Breslow <- mean(log(last.imputed.training.data.set$Breslow))
form.BS.Rec.5 <- S.Rec.5.training ~
  SNstatus +
  Age.SN +
  Ulceration +
  Loc_CAT +
  I(log(Breslow)-c.Breslow) +
  log(Rdamcrit) +
  SNstatus:I(log(Breslow)-c.Breslow)
f.mi.BS.Rec.5 <- Hmisc::fit.mult.impute(form.BS.Rec.5,cph,xtrans=imputed.training.data,data=training.data.to.be.imputed,n.impute=m,
                           pr=FALSE,fit.reps=TRUE,y=TRUE,x=TRUE,se.fit=TRUE)
# backward selection
# round(stats::anova(f.mi.BS.Rec.5)[order(stats::anova(f.mi.BS.Rec.5)[, 'P']),], 3)
summary(f.mi.BS.Rec.5, Breslow=exp(c.Breslow))
coef(f.mi.BS.Rec.5)
stats::anova(f.mi.BS.Rec.5)
```

# Refitted full model for melanoma specific mortality
```{r, eval=TRUE}
form.full.MSM.5.refit <- S.MSM.5.training ~
  SNstatus +
  Sex +
  Age.SN +
  Ulceration +
  Loc_CAT +
  Histology +
  log(Breslow) +
  multiple.fields +
  Tot_SNs_neg +
  Tot_SNs_pos +
  log(Rdamcrit) +
  Dewar +
  SNstatus:Sex +
  SNstatus:Age.SN +
  SNstatus:Ulceration +
  SNstatus:Loc_CAT +
  SNstatus:Histology +
  SNstatus:log(Breslow) +
  SNstatus:multiple.fields
f.mi.full.MSM.5.refit <- Hmisc::fit.mult.impute(form.full.MSM.5.refit,cph,xtrans=imputed.training.data,data=training.data.to.be.imputed,n.impute=m,
                                 pr=FALSE,fit.reps=TRUE,y=TRUE,x=TRUE,se.fit=TRUE)
summary(f.mi.full.MSM.5.refit, Tot_SNs_pos=c(0, 1))
stats::anova(f.mi.full.MSM.5.refit)
```

# Refitted final model for melanoma specific mortality
```{r, eval=TRUE}
form.BS.MSM.5.refit <- S.MSM.5.training ~
  SNstatus +
  Age.SN +
  Ulceration +
  Loc_CAT +
  I(log(Breslow)-c.Breslow) +
  log(Rdamcrit) +
  SNstatus:I(log(Breslow)-c.Breslow)
f.mi.BS.MSM.5.refit <- Hmisc::fit.mult.impute(form.BS.MSM.5.refit,cph,xtrans=imputed.training.data,data=training.data.to.be.imputed,n.impute=m,
                                  pr=FALSE,fit.reps=TRUE,y=TRUE,x=TRUE,se.fit=TRUE)
summary(f.mi.BS.MSM.5.refit, Breslow=exp(c.Breslow))
stats::anova(f.mi.BS.MSM.5.refit)
```

# Calibrated final model for melanoma specific mortality
Fit disease specific mortality to the linear predictor of recurrence
```{r, eval=TRUE}
lp.BS <- f.mi.BS.Rec.5$linear
f.mi.BS.MSM.5 <- rms::cph(S.MSM.5.training~lp.BS,y=TRUE,x=TRUE)
MSM.cal.fact <- stats::coef(f.mi.BS.MSM.5)
MSM.cal.fact
stats::confint(f.mi.BS.MSM.5)
utils::write.table(paste0(sprintf("%.2f", stats::coef(f.mi.BS.MSM.5)), " [",
            sprintf("%.2f", as.numeric(stats::confint(f.mi.BS.MSM.5))[1]), "; ",
            sprintf("%.2f", as.numeric(stats::confint(f.mi.BS.MSM.5))[2]), "]"),
            file='Z:/Project Melanoom/PaperMelanoma/Results/calibration.MSS.txt', 
            row.names=FALSE, col.names=FALSE, sep=",")

save(list=c("dd.training", "training.data.to.be.imputed", "m", "imputed.training.data", "last.imputed.training.data.set", 
            "horizon", "S.Rec.training", "S.MSM.training", "S.Rec.5.training", "S.MSM.5.training",
            "f.mi.full.Rec.5", "f.mi.BS.Rec.5", "f.mi.full.MSM.5.refit", "f.mi.BS.MSM.5.refit", "f.mi.BS.MSM.5", "MSM.cal.fact", "c.Breslow", 
            "form.full.Rec.5", "form.BS.Rec.5", "form.full.MSM.5.refit", "form.BS.MSM.5.refit"),
     file='Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.Rdata', compress=TRUE)
load('Z:/Project Melanoom/PaperMelanoma/Data/validation/validation.data.RData')
save(list=c("dd.validation", "validation.data.to.be.imputed", "m", "imputed.validation.data", "last.imputed.validation.data.set", "horizon", 
            "S.Rec.5.validation", "S.MSM.5.validation", "f.mi.BS.Rec.5"),
     file=paste0('Z:/Project Melanoom/PaperMelanoma/Data/validation/validation.data.Rdata'), compress=TRUE)
```