### 2022-12-19 Code/10_AJCC.Rmd

---
title: "Calculate AJCC"
author: "Carolien C.H.M. Maas"
date: "December 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Calculate AJCC

### Load packages, functions, and data 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# erase memory
rm(list = ls(all.names = TRUE))

# for reproducibility
set.seed(100)

# load libraries
library(PredictionTools)  # to plot calibration
library(compareC)         # compare AJCC to model
```
# Obtain C-index of AJCC7 and AJCC8
```{r, eval=TRUE}
calculate.AJCC.7.8 <- function(imputed.data=NULL, type=NULL, S.Rec.5=NULL){
  ####
  #### AJCC 7
  ####
  TNM.7 <- c()
  AJCC.7 <- c()
  cindex.AJCC.7 <- rep(0, m)
  cindex.AJCC.7.se <- rep(0, m)
  cindex.AJCC.7.pos <- rep(0, m)
  cindex.AJCC.7.se.pos <- rep(0, m)
  cindex.AJCC.7.neg <- rep(0, m)
  cindex.AJCC.7.se.neg <- rep(0, m)
  for (i in 1:m){
    imputation.i <- mice::complete(imputed.data,i)
    TNM.7.i <- rep("", nrow(imputation.i))
    AJCC.7.i <- rep(0, nrow(imputation.i))
    for (patient in 1:nrow(imputation.i)){
      breslow.patient <- imputation.i[patient, "Breslow"]
      ulceration.patient <- as.character(imputation.i[patient, "Ulceration"])
      mitosis.patient <- as.character(imputation.i[patient, "Mitosis"])
      nodes.patient <- imputation.i[patient, "Tot_SNs_pos"]

      snstatus.patient <- imputation.i[patient, "SNstatus"]

      # T stage
      if (breslow.patient <= 1){
        if (ulceration.patient=="No" & mitosis.patient=="No"){
          TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T1a")
        } else if (ulceration.patient=="Yes" | mitosis.patient=="Yes"){
          TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T1b")
        }
      } else if (breslow.patient>1 & breslow.patient <= 2){
        if (ulceration.patient=="No"){
          TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T2a")
        } else if (ulceration.patient=="Yes"){
          TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T2b")
        }
      } else if (breslow.patient>2 & breslow.patient <= 4){
        if (ulceration.patient=="No"){
          TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T3a")
        } else if (ulceration.patient=="Yes"){
          TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T3b")
        }
      } else if (breslow.patient>4){
        if (ulceration.patient=="No"){
          TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T4a")
        } else if (ulceration.patient=="Yes"){
          TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T4b")
        }
      }

      # N stage
      if (nodes.patient == 0){
        TNM.7.i[patient] <- paste0(TNM.7.i[patient], "N0")
      } else if (nodes.patient==1){
        TNM.7.i[patient] <- paste0(TNM.7.i[patient], "N1a")
      } else if (nodes.patient==2 | nodes.patient==3){
        TNM.7.i[patient] <- paste0(TNM.7.i[patient], "N2a")
      } else if (nodes.patient>3){
        TNM.7.i[patient] <- paste0(TNM.7.i[patient], "N3a")
      }

      # M stage for all patients in our data set
      TNM.7.i[patient] <- paste0(TNM.7.i[patient], "M0")

      # determine AJCC.7.i
      if (TNM.7.i[patient]=="T1aN0M0"){
        AJCC.7.i[patient] <- 1.1
      } else if (TNM.7.i[patient]=="T1bN0M0" | TNM.7.i[patient]=="T2aN0M0"){
        AJCC.7.i[patient] <- 1.2
      } else if (TNM.7.i[patient]=="T2bN0M0" | TNM.7.i[patient]=="T3aN0M0"){
        AJCC.7.i[patient] <- 2.1
      } else if (TNM.7.i[patient]=="T3bN0M0" | TNM.7.i[patient]=="T4aN0M0"){
        AJCC.7.i[patient] <- 2.2
      } else if (TNM.7.i[patient]=="T4bN0M0"){
        AJCC.7.i[patient] <- 2.3
      } else if (TNM.7.i[patient]=="T1aN1aM0" | TNM.7.i[patient]=="T2aN1aM0" | # N1a
                 TNM.7.i[patient]=="T3aN1aM0" | TNM.7.i[patient]=="T4aN1aM0" |
                 TNM.7.i[patient]=="T1aN2aM0" | TNM.7.i[patient]=="T2aN2aM0" | # N2a
                 TNM.7.i[patient]=="T3aN2aM0" | TNM.7.i[patient]=="T4aN2aM0"){
        AJCC.7.i[patient] <- 3.1
      } else if (TNM.7.i[patient]=="T1bN1aM0" | TNM.7.i[patient]=="T2bN1aM0" | # N1a
                 TNM.7.i[patient]=="T3bN1aM0" | TNM.7.i[patient]=="T4bN1aM0" |
                 TNM.7.i[patient]=="T1bN2aM0" | TNM.7.i[patient]=="T2bN2aM0" | # N2a
                 TNM.7.i[patient]=="T3bN2aM0" | TNM.7.i[patient]=="T4bN2aM0" |
                 TNM.7.i[patient]=="T1aN1bM0" | TNM.7.i[patient]=="T2aN1bM0" | # N1b
                 TNM.7.i[patient]=="T3aN1bM0" | TNM.7.i[patient]=="T4aN1bM0" |
                 TNM.7.i[patient]=="T1aN2bM0" | TNM.7.i[patient]=="T2aN2bM0" | # N2b
                 TNM.7.i[patient]=="T3aN2bM0" | TNM.7.i[patient]=="T4aN2bM0" |
                 TNM.7.i[patient]=="T1aN2cM0" | TNM.7.i[patient]=="T2aN2cM0" | # N2c
                 TNM.7.i[patient]=="T3aN2cM0" | TNM.7.i[patient]=="T4aN2cM0"){
        AJCC.7.i[patient] <- 3.2
      } else {
        AJCC.7.i[patient] <- 3.3
      }
    }

    # save TNM.7 and AJCC.7 for each imputated data set
    TNM.7 <- cbind(TNM.7, TNM.7.i)
    AJCC.7 <- cbind(AJCC.7, AJCC.7.i)

    # compute C-index
    rc <- Hmisc::rcorr.cens(-AJCC.7.i, S.Rec.5) # add a minus to revert class order: from worst to best
    cindex.AJCC.7[i]<-rc["C Index"]
    cindex.AJCC.7.se[i]<-rc["S.D."]/2

    # compute C-index for only positive sentinel node patients
    rc.pos <- Hmisc::rcorr.cens(-AJCC.7.i[imputation.i$SNstatus=="Positive"], S.Rec.5[imputation.i$SNstatus=="Positive",])
    cindex.AJCC.7.pos[i]<-rc.pos["C Index"]
    cindex.AJCC.7.se.pos[i]<-rc.pos["S.D."]/2

    # compute C-index for only negative sentinel node patients
    rc.neg <- Hmisc::rcorr.cens(-AJCC.7.i[imputation.i$SNstatus=="Negative"], S.Rec.5[imputation.i$SNstatus=="Negative",])
    cindex.AJCC.7.neg[i]<-rc.neg["C Index"]
    cindex.AJCC.7.se.neg[i]<-rc.neg["S.D."]/2
  }

  # combine C-index for each imputed data set
  cindex.mi.AJCC.7<-Rubin.combine(cindex.AJCC.7,cindex.AJCC.7.se)
  cindex.mi.AJCC.7.pos<-Rubin.combine(cindex.AJCC.7.pos,cindex.AJCC.7.se.pos)
  cindex.mi.AJCC.7.neg<-Rubin.combine(cindex.AJCC.7.neg,cindex.AJCC.7.se.neg)

  # save C-index and confidence interval
  AJCC.7.table <- rbind(c("All patients:", sprintf("%.2f", cindex.mi.AJCC.7$est),
                paste0("[", sprintf("%.2f", cindex.mi.AJCC.7$est+qnorm(.025)*cindex.mi.AJCC.7$se),
                        "; ", sprintf("%.2f", cindex.mi.AJCC.7$est+qnorm(.975)*cindex.mi.AJCC.7$se), "]")),
              c("Negative SN patients:", sprintf("%.2f", cindex.mi.AJCC.7.neg$est),
                    paste0("[", sprintf("%.2f", cindex.mi.AJCC.7.neg$est+qnorm(.025)*cindex.mi.AJCC.7.neg$se),
                           "; ", sprintf("%.2f", cindex.mi.AJCC.7.neg$est+qnorm(.975)*cindex.mi.AJCC.7.neg$se), "]")),
              c("Positive SN patients:", sprintf("%.2f", cindex.mi.AJCC.7.pos$est),
                    paste0("[", sprintf("%.2f", cindex.mi.AJCC.7.pos$est+qnorm(.025)*cindex.mi.AJCC.7.pos$se),
                           "; ", sprintf("%.2f", cindex.mi.AJCC.7.pos$est+qnorm(.975)*cindex.mi.AJCC.7.pos$se), "]")))

  ####
  #### AJCC8
  ####
  TNM.8 <- c()
  AJCC.8 <- c()
  cindex.AJCC.8 <- rep(0, m)
  cindex.AJCC.8.se <- rep(0, m)
  cindex.AJCC.8.pos <- rep(0, m)
  cindex.AJCC.8.se.pos <- rep(0, m)
  cindex.AJCC.8.neg <- rep(0, m)
  cindex.AJCC.8.se.neg <- rep(0, m)
  for (i in 1:m){
    imputation.i <- mice::complete(imputed.data,i)
    TNM.8.i <- rep("", nrow(imputation.i))
    AJCC.8.i <- rep(0, nrow(imputation.i))
    for (patient in 1:nrow(imputation.i)){
      breslow.patient <- imputation.i[patient, "Breslow"]
      ulceration.patient <- as.character(imputation.i[patient, "Ulceration"])
      mitosis.patient <- as.character(imputation.i[patient, "Mitosis"])
      nodes.patient <- imputation.i[patient, "Tot_SNs_pos"]

      # T stage
      if (breslow.patient <= 1){
        if (ulceration.patient=="No" & breslow.patient < 0.8){
          TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T1a")
        } else {
          TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T1b")
        }
      } else if (breslow.patient>1 & breslow.patient <= 2){
        if (ulceration.patient=="No"){
          TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T2a")
        } else if (ulceration.patient=="Yes"){
          TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T2b")
        }
      } else if (breslow.patient>2 & breslow.patient <= 4){
        if (ulceration.patient=="No"){
          TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T3a")
        } else if (ulceration.patient=="Yes"){
          TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T3b")
        }
      } else if (breslow.patient>4){
        if (ulceration.patient=="No"){
          TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T4a")
        } else if (ulceration.patient=="Yes"){
          TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T4b")
        }
      }

      # N stage
      if (nodes.patient == 0){
        TNM.8.i[patient] <- paste0(TNM.8.i[patient], "N0")
      } else if (nodes.patient==1){
        TNM.8.i[patient] <- paste0(TNM.8.i[patient], "N1a")
      } else if (nodes.patient==2 | nodes.patient==3){
        TNM.8.i[patient] <- paste0(TNM.8.i[patient], "N2a")
      } else if (nodes.patient>3){
        TNM.8.i[patient] <- paste0(TNM.8.i[patient], "N3a")
      }

      # M stage for all patients in our data set
      TNM.8.i[patient] <- paste0(TNM.8.i[patient], "M0")

      # determine AJCC.8.i
      if (TNM.8.i[patient]=="T1aN0M0"){
        AJCC.8.i[patient] <- 1.1
      } else if (TNM.8.i[patient]=="T1bN0M0" | TNM.8.i[patient]=="T2aN0M0"){
        AJCC.8.i[patient] <- 1.2
      } else if (TNM.8.i[patient]=="T2bN0M0" | TNM.8.i[patient]=="T3aN0M0"){
        AJCC.8.i[patient] <- 2.1
      } else if (TNM.8.i[patient]=="T3bN0M0" | TNM.8.i[patient]=="T4aN0M0"){
        AJCC.8.i[patient] <- 2.2
      } else if (TNM.8.i[patient]=="T4bN0M0"){
        AJCC.8.i[patient] <- 2.3
      } else if (TNM.8.i[patient]=="T1aN1aM0" | TNM.8.i[patient]=="T1bN1aM0" | TNM.8.i[patient]=="T2aN1aM0"
                 | TNM.8.i[patient]=="T1aN2aM0" | TNM.8.i[patient]=="T1bN2aM0" | TNM.8.i[patient]=="T2aN2aM0"){
        AJCC.8.i[patient] <- 3.1
      } else if (TNM.8.i[patient]=="T2bN1aM0" | TNM.8.i[patient]=="T3aN1aM0"){
        AJCC.8.i[patient] <- 3.2
      } else if (TNM.8.i[patient]=="T4bN3aM0"){
        AJCC.8.i[patient] <- 3.4
      } else {
        AJCC.8.i[patient] <- 3.3
      }

    }

    # save TNM.8 and AJCC.8 for each imputated data set
    TNM.8 <- cbind(TNM.8, TNM.8.i)
    AJCC.8 <- cbind(AJCC.8, AJCC.8.i)

    # compute C-index
    rc <- Hmisc::rcorr.cens(-AJCC.8.i, S.Rec.5) # add a minus to revert class order: from worst to best
    cindex.AJCC.8[i]<-rc["C Index"]
    cindex.AJCC.8.se[i]<-rc["S.D."]/2

    # compute C-index for only positive sentinel node patients
    rc.pos <- Hmisc::rcorr.cens(-AJCC.8.i[imputation.i$SNstatus=="Positive"], S.Rec.5[imputation.i$SNstatus=="Positive",])
    cindex.AJCC.8.pos[i]<-rc.pos["C Index"]
    cindex.AJCC.8.se.pos[i]<-rc.pos["S.D."]/2

    # compute C-index for only negative sentinel node patients
    rc.neg <- Hmisc::rcorr.cens(-AJCC.8.i[imputation.i$SNstatus=="Negative"], S.Rec.5[imputation.i$SNstatus=="Negative",])
    cindex.AJCC.8.neg[i]<-rc.neg["C Index"]
    cindex.AJCC.8.se.neg[i]<-rc.neg["S.D."]/2
  }
  # combine C-index for each imputed data set
  cindex.mi.AJCC.8<-Rubin.combine(cindex.AJCC.8,cindex.AJCC.8.se)
  cindex.mi.AJCC.8.pos<-Rubin.combine(cindex.AJCC.8.pos,cindex.AJCC.8.se.pos)
  cindex.mi.AJCC.8.neg<-Rubin.combine(cindex.AJCC.8.neg,cindex.AJCC.8.se.neg)

  # save C-index and confidence interval
  AJCC.8.table <- rbind(c("All patients:", sprintf("%.2f", cindex.mi.AJCC.8$est),
                          paste0("[", sprintf("%.2f", cindex.mi.AJCC.8$est+qnorm(.025)*cindex.mi.AJCC.8$se),
                                 "; ", sprintf("%.2f", cindex.mi.AJCC.8$est+qnorm(.975)*cindex.mi.AJCC.8$se), "]")),
                    c("Negative SN patients", sprintf("%.2f", cindex.mi.AJCC.8.neg$est),
                          paste0("[", sprintf("%.2f", cindex.mi.AJCC.8.neg$est+qnorm(.025)*cindex.mi.AJCC.8.neg$se),
                                 "; ", sprintf("%.2f", cindex.mi.AJCC.8.neg$est+qnorm(.975)*cindex.mi.AJCC.8.neg$se), "]")),
                    c("Positive SN patients:", sprintf("%.2f", cindex.mi.AJCC.8.pos$est),
                          paste0("[", sprintf("%.2f", cindex.mi.AJCC.8.pos$est+qnorm(.025)*cindex.mi.AJCC.8.pos$se),
                                 "; ", sprintf("%.2f", cindex.mi.AJCC.8.pos$est+qnorm(.975)*cindex.mi.AJCC.8.pos$se), "]")))

  # table of number of AJCC.7 and AJCC.8
  table.AJCC.7 <- c()
  table.AJCC.8 <- c()
  for (i in 1:m){
    table.AJCC.7 <- cbind(table.AJCC.7, as.numeric(table(AJCC.7[,i])))
    table.AJCC.8 <- cbind(table.AJCC.8, as.numeric(table(AJCC.8[,i])))
  }

  # print(sort(unique(AJCC.7[,1])))
  # print(sort(unique(AJCC.8[,1])))

  if (type=="training"){
    AJCC.7.8.count <- cbind(c("1.1", "1.2", "2.1", "2.2", "2.3", "3.1", "3.2", "3.3", "3.4"),
                          c(round(rowMeans(table.AJCC.7), 0), 0),
                          round(rowMeans(table.AJCC.8)))
  } else if (type=="validation"){
    AJCC.7.8.count <- cbind(c("1.1", "1.2", "2.1", "2.2", "2.3", "3.1", "3.2", "3.3", "3.4"),
                          c(round(rowMeans(table.AJCC.7), 0), 0),
                          c(0, round(rowMeans(table.AJCC.8), 0)))
  }
  return(list(AJCC.7=AJCC.7, AJCC.7.table=AJCC.7.table,
              AJCC.8=AJCC.8, AJCC.8.table=AJCC.8.table, AJCC.7.8.count=AJCC.7.8.count))
}
load("Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.RData")
train.AJCC <- calculate.AJCC.7.8(imputed.data=imputed.training.data, S.Rec.5=S.Rec.5.training, type="training")
load("Z:/Project Melanoom/PaperMelanoma/Data/validation/validation.data.RData")
validation.AJCC <- calculate.AJCC.7.8(imputed.data=imputed.validation.data, S.Rec.5=S.Rec.5.validation, type="validation")
```

# Compare C-index of AJCC and developed model
```{r, eval=TRUE}
for (edition in c(7, 8)){
  for (patient.type in c("All", "Negative", "Positive")){
    model.C <- c()
    se.C <- c()
    diff <- c()
    se <- c()
    for (i in 1:m){
      # compute linear predictor
      lp.i <- predict(f.mi.BS.Rec.5,
                    newdata=mice::complete(imputed.validation.data, i),
                    type="lp")

      if (patient.type=="All"){
        lp.i <- lp.i
        S <- S.Rec.5.validation
        AJCC <- eval(parse(text=paste0("validation.AJCC$AJCC.", edition, "[, i]")))
      } else if (patient.type=="Positive"){
        lp.i <- lp.i[validation.data.to.be.imputed$SNstatus=="Positive"]
        S <- S.Rec.5.validation[validation.data.to.be.imputed$SNstatus=="Positive",]
        AJCC <- eval(parse(text=paste0("validation.AJCC$AJCC.", edition, "[validation.data.to.be.imputed$SNstatus==\"Positive\", i]")))
      } else if (patient.type=="Negative"){
        lp.i <- lp.i[validation.data.to.be.imputed$SNstatus=="Negative"]
        S <- S.Rec.5.validation[validation.data.to.be.imputed$SNstatus=="Negative",]
        AJCC <- eval(parse(text=paste0("validation.AJCC$AJCC.", edition, "[validation.data.to.be.imputed$SNstatus==\"Negative\", i]")))
      }

      # compute predictions from model
      p.i <- 1-exp(-h0.Rec*exp(lp.i))

      # compare AJCC to model
      compare.i <- compareC::compareC(S[,1], S[,2], -p.i, -AJCC)

      model.C <- c(model.C, compare.i$est.c["Cxy"])
      se.C <- c(se.C, sqrt(compare.i$est.varCxy))

      # save differences and SE of difference
      diff <- c(diff, compare.i$est.diff_c)
      se <- c(se, sqrt(compare.i$est.vardiff_c))
    }
    assign(paste0("diff.", patient.type, ".", edition), Rubin.combine(diff, se))
    assign(paste0("model.", patient.type), Rubin.combine(model.C, se.C))
  }
}

# save in Table
compare.AJCC.model <- rbind(c("All patients:", 
                sprintf("%.2f", diff.All.7$est),
                paste0("[", sprintf("%.2f", diff.All.7$est+qnorm(.025)*diff.All.7$se),
                        "; ", sprintf("%.2f", diff.All.7$est+qnorm(.975)*diff.All.7$se), "]"),
                sprintf("%.2f", diff.All.8$est),
                paste0("[", sprintf("%.2f", diff.All.8$est+qnorm(.025)*diff.All.8$se),
                        "; ", sprintf("%.2f", diff.All.8$est+qnorm(.975)*diff.All.8$se), "]")),
              c("Negative SN patients:", 
                sprintf("%.2f", diff.Negative.7$est),
                paste0("[", sprintf("%.2f", diff.Negative.7$est+qnorm(.025)*diff.Negative.7$se),
                       "; ", sprintf("%.2f", diff.Negative.7$est+qnorm(.975)*diff.Negative.7$se), "]"),
                sprintf("%.2f", diff.Negative.8$est),
                paste0("[", sprintf("%.2f", diff.Negative.8$est+qnorm(.025)*diff.Negative.8$se),
                       "; ", sprintf("%.2f", diff.Negative.8$est+qnorm(.975)*diff.Negative.8$se), "]")),
              c("Positive SN patients:", 
                sprintf("%.2f", diff.Positive.7$est),
                paste0("[", sprintf("%.2f", diff.Positive.7$est+qnorm(.025)*diff.Positive.7$se),
                       "; ", sprintf("%.2f", diff.Positive.7$est+qnorm(.975)*diff.Positive.7$se), "]"),
                sprintf("%.2f", diff.Positive.8$est),
                paste0("[", sprintf("%.2f", diff.Positive.8$est+qnorm(.025)*diff.Positive.8$se),
                       "; ", sprintf("%.2f", diff.Positive.8$est+qnorm(.975)*diff.Positive.8$se), "]")))

CIndex.val <- rbind(c("All patients:",
                sprintf("%.2f", model.All$est),
                paste0("[", sprintf("%.2f", model.All$est+qnorm(.025)*model.All$se), 
                      "; ", sprintf("%.2f", model.All$est+qnorm(.975)*model.All$se), "]")),
                c("Negative SN patients:",
                sprintf("%.2f", model.Negative$est), 
                paste0("[", sprintf("%.2f", model.Negative$est+qnorm(.025)*model.Negative$se), 
                      "; ", sprintf("%.2f", model.Negative$est+qnorm(.975)*model.Negative$se), "]")),
                c("Positve SN patients:",
                sprintf("%.2f", model.Positive$est), 
                paste0("[", sprintf("%.2f", model.Positive$est+qnorm(.025)*model.Positive$se), 
                      "; ", sprintf("%.2f", model.Positive$est+qnorm(.975)*model.Positive$se), "]")))
```

# Write to Excel
```{r, eval=TRUE}
utils::write.table(cbind(train.AJCC$AJCC.7.table, rep("", 3), train.AJCC$AJCC.8.table[,-1]),
              file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/AJCC.train.txt"),
              row.names=FALSE, col.names=FALSE, sep=",")
utils::write.table(cbind(CIndex.val, rep("", 3),
                   validation.AJCC$AJCC.7.table[, -1], rep("", 3), 
                   compare.AJCC.model[, 2:3], rep("", 3),
                   validation.AJCC$AJCC.8.table[,-1], rep("", 3),
                   compare.AJCC.model[, 4:5]),
              file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/AJCC.validation.txt"),
              row.names=FALSE, col.names=FALSE, sep=",")
utils::write.table(cbind(train.AJCC$AJCC.7.8.count, rep("", 3), validation.AJCC$AJCC.7.8.count[,-1]),
              file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/AJCC.7.8.count.txt"),
              row.names=FALSE, col.names=FALSE, sep=",")
```