### 2022-12-19 Code/6_CreateTablesandPlots.Rmd

---
title: "Create Tables"
author: "Carolien C.H.M. Maas"
date: "December 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Create Tables

### Load packages, functions, and data 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# erase memory
rm(list = ls(all.names = TRUE))

# for reproducibility
set.seed(100)

# load libraries
library(foreign)
library(mice)
library(survminer)
library(prodlim)
library(rms)
library(Hmisc)
library(png)
library(PredictionTools)
source("Z:/Project Melanoom/PaperMelanoma/Code/PaperMelanomaFunctions.R")
```

### Create variables with results
##### Put hazard ratio"s and anova Chi-square statistics of positive and negative SN patients in Table of final recurrence model and refitted MSM model
```{r, eval=TRUE}
load("Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.RData")
options(datadist="dd.training")

for (model in c("Rec.5", "MSM.5.refit")){
  # load model
  f.mi.full <- eval(parse(text=paste0("f.mi.full.", model)))

  # Coefficients with CI
  # Negative results
  neg.results <- summary(f.mi.full, SNstatus="Negative", Loc_CAT="arm")
  neg.table <- rbind(rep(NA, 3), # SN status
                     exp(neg.results[c("Sex - male:female",
                                       "Age.SN",
                                       "Ulceration - Yes:No",
                                       "Loc_CAT - leg:arm",
                                       "Loc_CAT - trunk:arm",
                                       "Loc_CAT - head & neck:arm",
                                       "Histology - NM:SSM",
                                       "Histology - ALM:SSM",
                                       "Histology - other:SSM",
                                       "Breslow",
                                       "multiple.fields - Multiple SN fields:Other",
                                       "Tot_SNs_neg"), c("Effect", "Lower 0.95", "Upper 0.95")]),
                     matrix(NA, nrow=6, ncol=3)) # Tot_SNs_pos, Rdamcrit, Combined, Parenchymal, Multifocal, Extensive
  # add CI
  neg.table <- cbind(sprintf("%.2f", neg.table[, 1]),
                     paste0("[", sprintf("%.2f", neg.table[, 2]), "; ", sprintf("%.2f", neg.table[, 3]), "]"))

  # Positive results
  pos.results <- summary(f.mi.full, SNstatus="Positive", Loc_CAT="arm", Tot_SNs_pos=c(0, 1))
  pos.table <- rbind(exp(neg.results["SNstatus - Positive:Negative", c("Effect", "Lower 0.95", "Upper 0.95")]),
                     exp(pos.results[c("Sex - male:female",
                                       "Age.SN",
                                       "Ulceration - Yes:No",
                                       "Loc_CAT - leg:arm",
                                       "Loc_CAT - trunk:arm",
                                       "Loc_CAT - head & neck:arm",
                                       "Histology - NM:SSM",
                                       "Histology - ALM:SSM",
                                       "Histology - other:SSM",
                                       "Breslow",
                                       "multiple.fields - Multiple SN fields:Other",
                                       "Tot_SNs_neg",
                                       "Tot_SNs_pos",
                                       "Rdamcrit",
                                       "Dewar - combined:subcap",
                                       "Dewar - parenchymal:subcap",
                                       "Dewar - multifocal:subcap",
                                       "Dewar - extensive:subcap"), c("Effect", "Lower 0.95", "Upper 0.95")]))
  # add CI
  pos.table <- cbind(sprintf("%.2f", pos.table[, 1]),
                     paste0("[", sprintf("%.2f", pos.table[, 2]), "; ", sprintf("%.2f", pos.table[, 3]), "]"))

  # combine postive with negative
  full.table <- cbind(neg.table, pos.table)

  # add reference rows
  final.table <- rbind(full.table[1:4,],
                       rep(NA, 4),     # location
                       c("(ref)", NA, "(ref)", NA), # ref arm
                       full.table[5:7,],   # other locations
                       rep(NA, 4),     # histology
                       c("(ref)", NA, "(ref)", NA), # ref SMM
                       full.table[8:15,],
                       rep(NA, 4),     # Dewar
                       c(rep(NA, 2), "(ref)", NA), # ref subcap
                       full.table[16:nrow(full.table),]) # other Dewar

  # add Chi-square statistics
  anova.out <- anova(f.mi.full)[, "Chi-Square"]
  anova.main <- c(anova.out[c("SNstatus  (Factor+Higher Order Factors)",
                              "Sex  (Factor+Higher Order Factors)",
                              "Age.SN  (Factor+Higher Order Factors)",
                              "Ulceration  (Factor+Higher Order Factors)",
                              "Loc_CAT  (Factor+Higher Order Factors)")],
                  rep(NA, 4),
                  anova.out["Histology  (Factor+Higher Order Factors)"],
                  rep(NA, 4),
                  anova.out[c("Breslow  (Factor+Higher Order Factors)",
                              "multiple.fields  (Factor+Higher Order Factors)",
                              "Tot_SNs_neg",
                              "Tot_SNs_pos",
                              "Rdamcrit",
                              "Dewar")],
                  rep(NA, 5))
  anova.int <- c(anova.out[c("TOTAL INTERACTION",
                             "SNstatus * Sex  (Factor+Higher Order Factors)",
                             "SNstatus * Age.SN  (Factor+Higher Order Factors)",
                             "SNstatus * Ulceration  (Factor+Higher Order Factors)",
                             "SNstatus * Loc_CAT  (Factor+Higher Order Factors)")],
                 rep(NA, 4),
                 anova.out["SNstatus * Histology  (Factor+Higher Order Factors)"],
                 rep(NA, 4),
                 anova.out[c("SNstatus * Breslow  (Factor+Higher Order Factors)",
                             "SNstatus * multiple.fields  (Factor+Higher Order Factors)")],
                 rep(NA, 9))
  final.table <- cbind(final.table, sprintf("%.0f", anova.main), sprintf("%.0f", anova.int))

  colnames(final.table) <- c("HR.neg", "CI.neg", "HR.pos", "CI.pos", "Main", "Int")
  assign(paste0("final.full.table.", model), as.data.frame(final.table))
}
```

##### Put hazard ratio"s and anova Chi-square statistics of positive and negative SN patients in Table of final recurrence model and refitted MSM model
```{r, eval=TRUE}
for (model in c("Rec.5", "MSM.5.refit")){
  f.mi.BS <- eval(parse(text=paste0("f.mi.BS.", model)))

  # Coefficients with CI
  # Negative results
  neg.results.BS <- summary(f.mi.BS, SNstatus="Negative", Breslow=exp(c.Breslow), Loc_CAT="arm")
  neg.table.BS <- rbind(rep(NA, 3), # SN status
                        exp(neg.results.BS[c("Age.SN",
                                             "Ulceration - Yes:No",
                                             "Loc_CAT - leg:arm",
                                             "Loc_CAT - trunk:arm",
                                             "Loc_CAT - head & neck:arm",
                                             "Breslow"), c("Effect", "Lower 0.95", "Upper 0.95")]),
                        rep(NA, 3)) # Rdamcrit
  # add CI
  neg.table.BS <- cbind(sprintf("%.2f", neg.table.BS[, 1]),
                        paste0("[", sprintf("%.2f", neg.table.BS[, 2]), "; ", sprintf("%.2f", neg.table.BS[, 3]), "]"))

  # Positive results
  pos.results.BS <- summary(f.mi.BS, SNstatus="Positive", Breslow=exp(c.Breslow), Loc_CAT="arm")
  pos.table.BS <- rbind(exp(neg.results.BS["SNstatus - Positive:Negative", c("Effect", "Lower 0.95", "Upper 0.95")]),
                        exp(pos.results.BS[c("Age.SN",
                                             "Ulceration - Yes:No",
                                             "Loc_CAT - leg:arm",
                                             "Loc_CAT - trunk:arm",
                                             "Loc_CAT - head & neck:arm",
                                             "Breslow",
                                             "Rdamcrit"), c("Effect", "Lower 0.95", "Upper 0.95")]))
  # add CI
  pos.table.BS <- cbind(sprintf("%.2f", pos.table.BS[, 1]),
                        paste0("[", sprintf("%.2f", pos.table.BS[, 2]), "; ", sprintf("%.2f", pos.table.BS[, 3]), "]"))

  # combine postive with negative
  BS.table <- cbind(neg.table.BS, pos.table.BS)

  # add reference rows
  BS.table <- rbind(BS.table[1:3,],
                    rep(NA, 4),
                    c("(ref)", NA, "(ref)", NA),
                    BS.table[4:nrow(BS.table),])

  # add Chi-square statistics
  anova.out <- anova(f.mi.BS)[, "Chi-Square"]
  anova.main <- c(anova.out[c("SNstatus  (Factor+Higher Order Factors)",
                              "Age.SN",
                              "Ulceration",
                              "Loc_CAT")],
                  rep(NA, 4),
                  anova.out[c("Breslow  (Factor+Higher Order Factors)",
                              "Rdamcrit")])
  anova.int <- c(rep(NA, 8),
                 anova.out["SNstatus * Breslow  (Factor+Higher Order Factors)"],
                 NA)
  BS.table <- cbind(BS.table, sprintf("%.0f", anova.main), sprintf("%.0f", anova.int))

  colnames(BS.table) <- c("HR.neg", "CI.neg", "HR.pos", "CI.pos", "Main", "Int")
  assign(paste0("final.BS.table.", model), as.data.frame(BS.table))
}
```

# Calculate optimism using bootstrap using full model and doing backward selection for each bootstrap
```{r, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(100)
slope<-NULL
cindex.orig<-NULL
opt<-NULL
cindex.B<-NULL
for (i in 1:m){
  v<-validate(f.mi.full.Rec.5$fits[[i]],method="boot",bw=TRUE,rule="p",sls=0.05,B=100,pr=FALSE,type="individual")
  slope<-c(slope,v["Slope","test"]) # slope of test
  cindex.orig<-c(cindex.orig,(v["Dxy","index.orig"]+1)/2) # original C-index
  opt<-c(opt,v["Dxy","optimism"]/2) # optimism C-index
  cindex.B<-c(cindex.B,(v["Dxy","index.corrected"]+1)/2) # index corrected C-index
}
```

# Show optimism
```{r, eval=TRUE}
mean(cindex.orig)
optimism <- mean(opt)
optimism
mean(cindex.B)
mean(cindex.orig)-optimism
```

# Add C-index to model for Recurrence and MSM refit Table
```{r, eval=TRUE}
for (model in list("Rec.5", "MSM.5.refit")){
  if (model == "Rec.5"){
    # plot predict
    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/plot.predict.Rec.png"))
    plot.predict <- plot(Predict(f.mi.full.Rec.5))
    show(plot.predict)
    grDevices::dev.off()

    # plot predict Breslow
    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/plot.predict.Breslow.Rec.png"))
    plot.Breslow <- plot(Predict(f.mi.full.Rec.5, Breslow=0:20, SNstatus), ~ Breslow, groups="SNstatus",
                         adj.subtitle=FALSE, lwd=2, labels=FALSE,
                         par.settings=list(superpose.line=list(lty=c(1, 2), col=c("red", "blue"))),
                         key=list(cex=.8,corner = c(0.1, .9),lines=list(col=c("red","blue"), lty=c(1,2), lwd=2),text=list(c("Negative","Positive"))))
    show(plot.Breslow)
    grDevices::dev.off()
  }

  # discrimination full and backward selected model
  for (type in c("full", "BS")){
    # load model
    f.mi <- eval(parse(text=paste0("f.mi.", type, ".", model)))

    # load outcome
    if (model=="Rec.5"){
      S <- S.Rec.5.training
    }
    else if (model=="MSM.5.refit"){
      S <- S.MSM.5.training
    }

    # obtain predictions for each imputed data set
    p <- matrix(NA, nrow(training.data.to.be.imputed), m)
    for (i in 1:m){
      # predicted lp for imputation i
      lp.i <- predict(f.mi,newdata=mice::complete(imputed.training.data,i), type="lp")
      # baseline hazard
      f.basehaz <- survival::basehaz(f.mi,TRUE)
      h0.i <- f.basehaz$hazard[f.basehaz$time==max(f.basehaz$time[f.basehaz$time<=horizon])]
      # probabilities
      p.i <- fun.event(lp=lp.i, h0=h0.i)
      p[, i] <- p.i
    }

    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/all.centers.", model, ".", type, ".png"))
    out <- PredictionTools::val.surv.mi(p,S,g=5,main="All centers",
                       time=NULL,lim=c(0,1),dist=TRUE)
    grDevices::dev.off()

    final.full.table <- eval(parse(text=paste0("final.", type, ".table.", model)))
    final.full.table["C-index", "HR.neg"] <- sprintf("%.2f", out$cindex-optimism)
    final.full.table["C-index", "CI.neg"] <- paste0("[", sprintf("%.2f", out$cindex.lower-optimism),
                                                   "; ", sprintf("%.2f", out$cindex.upper-optimism), "]")

    assign(paste0("final.", type, ".table.", model), final.full.table)

    if (type=="BS"&model=="Rec.5"){
      CIndex.BS.Rec.5.All <- c(sprintf("%.2f", out$cindex-optimism), 
                              paste0("[", sprintf("%.2f", out$cindex.lower-optimism), 
                                     "; ", sprintf("%.2f", out$cindex.upper-optimism), "]"))
    }
  }
}
```

# Add the C-index for calibrated MSM model to Table for recurrence
```{r, eval=TRUE}
for (type in c("full", "BS")){
  f.mi <- eval(parse(text=paste0("f.mi.", type, ".Rec.5")))
  cindex.MSM.refit <- rep(0, m)
  cindex.se.MSM.refit <- rep(0, m)
  for (i in 1:m){
    rc <- Hmisc::rcorr.cens(-predict(f.mi,newdata=mice::complete(imputed.training.data,i), type="lp"), S.MSM.5.training)
    cindex.MSM.refit[i]<-rc["C Index"]
    cindex.se.MSM.refit[i]<-rc["S.D."]/2
  }
  cindex.mi.MSM.refit<-Rubin.combine(cindex.MSM.refit,cindex.se.MSM.refit)

  final.full.table <- eval(parse(text=paste0("final.", type, ".table.Rec.5")))
  final.full.table["C-index MSM", "HR.neg"] <- sprintf("%.2f", cindex.mi.MSM.refit$est-optimism)
  final.full.table["C-index MSM", "CI.neg"] <- paste0("[", sprintf("%.2f", cindex.mi.MSM.refit$est+qnorm(.025)*cindex.mi.MSM.refit$se-optimism),
                                                 "; ", sprintf("%.2f", cindex.mi.MSM.refit$est+qnorm(.975)*cindex.mi.MSM.refit$se-optimism), "]")

  assign(paste0("final.", type, ".table.Rec.5"), final.full.table)
}
```

# Save dataframes of result Tables
```{r, eval=TRUE}
# replace NA by blank
for (df in c("final.full.table.Rec.5", "final.full.table.MSM.5.refit", "final.BS.table.Rec.5", "final.BS.table.MSM.5.refit")){
  final.df <- eval(parse(text=df))
  final.df[final.df=="[NA; NA]"] <- NA
  final.df[final.df=="NA"] <- NA
  final.df[is.na(final.df)] <- ""
  assign(df, final.df)
}

# set rownames
final.full.table.Rec.5 <- as.data.frame(final.full.table.Rec.5)
row.names(final.full.table.Rec.5) <- c("Positive SN status", "Male",
                                    "Age iqr", "Ulceration", "Location",
                                    "Arm", "Leg", "Trunk", "Head and neck",
                                    "Histology", "SSM", "NM", "ALM", "Other",
                                    "Breslow iqr", "Multiple fields", 
                                    "Total number of negative nodes",
                                    "Total number of positive nodes",
                                    "SN tumour burden iqr", 
                                    "Location metastasis in lymph node",
                                    "Subcap", "Combined", "Parenchymal",
                                    "Multifocal", "Extensive",
                                    rownames(final.full.table.Rec.5)[(nrow(final.full.table.Rec.5)-1):nrow(final.full.table.Rec.5)])
final.full.table.MSM.5.refit <- as.data.frame(final.full.table.MSM.5.refit)
row.names(final.full.table.MSM.5.refit) <- c("Positive SN status", "Male",
                                          "Age", "Ulceration", "Location",
                                          "Arm", "Leg", "Trunk", "Head and neck",
                                          "Histology", "SSM", "NM", "ALM", "Other",
                                          "Breslow", "Multiple fields", 
                                          "Total number of negative nodes",
                                          "Total number of positive nodes",
                                          "SN tumour burden", 
                                          "Location metastasis in lymph node",
                                          "Subcap", "Combined", "Parenchymal",
                                          "Multifocal", "Extensive",
                                          row.names(final.full.table.MSM.5.refit)[nrow(final.full.table.MSM.5.refit)])
final.BS.table.Rec.5 <- as.data.frame(final.BS.table.Rec.5)
row.names(final.BS.table.Rec.5) <- c("Positive SN status", "Age", "Ulceration",
                                   "Location", "Arm", "Leg", "Trunk", 
                                   "Head and neck",
                                   "Breslow", "SN tumour burden", 
                                   row.names(final.BS.table.Rec.5)[(nrow(final.BS.table.Rec.5)-1):nrow(final.BS.table.Rec.5)])
final.BS.table.MSM.5.refit <- as.data.frame(final.BS.table.MSM.5.refit)
row.names(final.BS.table.MSM.5.refit) <- c("Positive SN status", "Age", "Ulceration",
                                   "Location", "Arm", "Leg", "Trunk", 
                                   "Head and neck",
                                   "Breslow", "SN tumour burden", 
                                   row.names(final.BS.table.MSM.5.refit)[nrow(final.BS.table.MSM.5.refit)])

# save tables
write.table(final.full.table.Rec.5,
            file="Z:/Project Melanoom/PaperMelanoma/Results/final.full.table.Rec.txt", 
            sep=",")
write.table(final.full.table.MSM.5.refit,
            file="Z:/Project Melanoom/PaperMelanoma/Results/final.full.table.MSM.refit.txt", 
            sep=",")
write.table(final.BS.table.Rec.5,
            file="Z:/Project Melanoom/PaperMelanoma/Results/final.BS.table.Rec.txt", 
            sep=",")
write.table(final.BS.table.MSM.5.refit,
            file="Z:/Project Melanoom/PaperMelanoma/Results/final.BS.table.MSM.refit.txt", 
            sep=",")

# show tables
final.full.table.Rec.5
final.full.table.MSM.5.refit
final.BS.table.Rec.5
final.BS.table.MSM.5.refit
```

# C-index of final model for recurrence stratified by SN status
```{r, eval=TRUE}
for (arm in c("Positive", "Negative")){
  f.mi.stratified <- eval(parse(text=paste0("f.mi.BS.Rec.5")))
  S.5.subset <- S.Rec.5.training[last.imputed.training.data.set$SNstatus==arm,]

  cindex.stratified <- rep(0, m)
  cindex.se.stratified <- rep(0, m)
  for (i in 1:m){
    subset.imputed.training.data <- as.data.frame(mice::complete(imputed.training.data,i)[last.imputed.training.data.set$SNstatus==arm,])
    rc <- Hmisc::rcorr.cens(-predict(f.mi.stratified,newdata=subset.imputed.training.data, type="lp"), S.5.subset)
    cindex.stratified[i]<-rc["C Index"]
    cindex.se.stratified[i]<-rc["S.D."]/2
  }
  cindex.mi.stratified<-Rubin.combine(cindex.stratified,cindex.se.stratified)

  assign(paste0("CIndex.BS.Rec.5.", arm), 
         c(sprintf("%.2f", cindex.mi.stratified$est), 
           paste0("[", sprintf("%.2f", cindex.mi.stratified$est+qnorm(.025)*cindex.mi.stratified$se), 
                  "; ", sprintf("%.2f", cindex.mi.stratified$est+qnorm(.975)*cindex.mi.stratified$se), "]")))
}

CIndex.Rec.Table <- rbind(CIndex.BS.Rec.5.All, CIndex.BS.Rec.5.Positive, CIndex.BS.Rec.5.Negative)
CIndex.Rec.Table
utils::write.table(CIndex.Rec.Table,
              file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/CIndex.train.txt"),
              row.names=FALSE, col.names=FALSE, sep=",")
```
# Plot cross-validated calibration of all centers combined and all centers seperate for recurrence and calibrated melanoma specific mortality model
```{r, eval=TRUE}
Centers <- sort(as.character(unique(training.data.to.be.imputed$Center)))
Centers.anonimized <- c("A", "B", "C", "D")
cuts <- 5
for (model in c("Rec.5", "MSM.5")){
  if (model == "Rec.5"){
    S.model <- S.Rec.5.training
    fact <- 1
  }
  if (model == "MSM.5"){
    S.model <- S.MSM.5.training
    fact <- MSM.cal.fact
  }

  for (type in c("full", "BS")){
    for (j in 1:4){
      S.j<-S.model[training.data.to.be.imputed$Center==Centers[j],]
      S.notj<-S.model[training.data.to.be.imputed$Center!=Centers[j],]
      S.Rec.5.notj<-S.Rec.5.training[training.data.to.be.imputed$Center!=Centers[j],]

      form.model<-eval(parse(text=paste0("form.", type, ".Rec.5")))
      form.notj<-update(form.model, S.Rec.5.notj  ~ . )
      f.mi.Rec.notj<-Hmisc::fit.mult.impute(form.notj,cph,xtrans=imputed.training.data,data=training.data.to.be.imputed,
                                 n.impute=m,pr=FALSE,fit.reps=TRUE,y=TRUE,x=TRUE,se.fit=TRUE,
                                 sub=training.data.to.be.imputed$Center!=Centers[j])

      if (model=="Rec.5"){
        f.basehaz.notj<-survival::basehaz(f.mi.Rec.notj,TRUE)
        h0.notj<-f.basehaz.notj$hazard[f.basehaz.notj$time==max(f.basehaz.notj$time[f.basehaz.notj$time<=horizon])]
      } else if (model=="MSM.5"){
        lp.notj<-f.mi.Rec.notj$linear
        f.MSM.notj<-rms::cph(S.notj~lp.notj,y=TRUE,x=TRUE)

        f.basehaz.notj<-basehaz(f.MSM.notj,TRUE)
        h0.notj<-f.basehaz.notj$hazard[f.basehaz.notj$time==max(f.basehaz.notj$time[f.basehaz.notj$time<=horizon])]
      }

      for (select.value in c("Positive", "Negative", "All")){
        # identify those patients that are selected
        if (select.value == "All"){
          select.patients <- TRUE
        } else if (select.value != "All"){
          select.patients <- training.data.to.be.imputed[training.data.to.be.imputed$Center==Centers[j],]$SNstatus==select.value
        }

        # calibration with metrics
        S <- S.j[select.patients,]
        p <- matrix(NA, nrow(S), m)
        for (i in 1:m){
          lp.i<-predict(f.mi.Rec.notj,newdata=mice::complete(imputed.training.data,i)[training.data.to.be.imputed$Center==Centers[j],], type="lp")

          if (model=="MSM.5"){
            lp.i<-predict(f.MSM.notj,newdata=lp.i)
          }

          if (select.value != "All"){
            lp.i <- lp.i[select.patients]
          }

          # baseline hazard
          if (model=="Rec.5"){
            f.basehaz<-basehaz(f.mi.Rec.notj,TRUE)
          }
          else if (model=="MSM.5"){
            f.basehaz<-basehaz(f.MSM.notj,TRUE)
          }
          h0.i <- f.basehaz$hazard[f.basehaz$time==max(f.basehaz$time[f.basehaz$time<=horizon])]

          # probabilities
          p.i <- fun.event(lp=fact*lp.i, h0=h0.i)
          p[, i] <- p.i
        }

        # omit the observation with predicted probability = 1
        omit <- unique(c(which(p[,1]==1), which(p[,2]==1), which(p[,3]==1), which(p[,4]==1), which(p[,5]==1)))
        if (length(omit)>0){
          cat("Omit:", omit, "\n")
          p <- p[-omit,]
          S <- S[-omit,]
        }

        png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/", model, ".", select.value, ".", type, ".", Centers[j], ".png"))
        out <- PredictionTools::val.surv.mi(p, S, g=5, main=paste("Center", Centers.anonimized[j]),
                           time=NULL,lim=c(0,1),dist=TRUE)
        grDevices::dev.off()
      }
    }
  }
}

# combine calibration plots into one
for (select.value in c("All", "Positive", "Negative")){
  for (y in c("Rec.5", "MSM.5")){
    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/", y, ".", select.value, ".png"), width=16, height=16, units="cm", res=300)
    par(mar=rep(0, 4))
    layout(matrix(1:4, ncol=2, byrow=TRUE))
    Centers.val <- rep(1:4, 3)
    for (i in 1:4){
      plot(NA, xlim=0:1, ylim=0:1, xaxt="n", yaxt="n", bty="n")
      img <- readPNG(paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/", y, ".", select.value, ".BS.", Centers[Centers.val[i]], ".png"))
      rasterImage(img, 0, 0, 1, 1)
    }
    grDevices::dev.off()
  }
}
```