### 2022-12-19 Code/8_CreateNomogram.Rmd

---
title: "Create Nomogram"
author: "Carolien C.H.M. Maas"
date: "December 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Create Nomogram

### Load packages, functions, and data 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# erase memory
rm(list = ls(all.names = TRUE))

# for reproducibility
set.seed(100)

# load libraries
library(rms)
library(survival)
source("Z:/Project Melanoom/PaperMelanoma/Code/PaperMelanomaFunctions.R")
```

# Nomogram
```{r, eval=TRUE}
load("Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.RData")
options(datadist="dd.training")

# Limits of Age, Breslow, Rdamcrit
Age.class <- c(20, 50, 80)
Breslow.class <- c(0.1, 0.3, 0.6, 1, 2, 3, 7)
Rdamcrit.class <- c(0.01, 0.1, 0.3, 1, 3, 10)

# initial nomogram
nom.0 <- rms::nomogram(f.mi.BS.Rec.5,maxscale=10,
                  Breslow=Breslow.class, Rdamcrit=Rdamcrit.class, Age.SN=Age.class)

# baseline hazard of recurrence averaged over all imputed data sets
h0.Rec.all <- c()
for (i in 1:m){
  f.Rec.basehaz.i <- basehaz(cph(form.BS.Rec.5, data=mice::complete(imputed.training.data, i), se.fit=TRUE, x=TRUE, y=TRUE))
  h0.Rec.i <- f.Rec.basehaz.i$hazard[f.Rec.basehaz.i$time==max(f.Rec.basehaz.i$time[f.Rec.basehaz.i$time<=horizon])]
  h0.Rec.all <- c(h0.Rec.all, h0.Rec.i)
}
h0.Rec <- mean(h0.Rec.all)

# lp range Recurrence
rc.Rec <- (1/attributes(nom.0)$info$sc)
int.Rec <- attributes(nom.0)$info$Intercept
lp.range <- int.Rec+rc.Rec*(0:10)*2

# nomogram
nom <- rms::nomogram(f.mi.BS.Rec.5, maxscale=10,
                Breslow=Breslow.class, Rdamcrit=Rdamcrit.class, Age.SN=Age.class,
                lp=TRUE, fun.at=fun.event(lp=lp.range, h0=h0.Rec))
attributes(nom)$names[attributes(nom)$names=="Age.SN"]<-"Age"
attributes(nom)$names[attributes(nom)$names=="Loc_CAT"]<-"Location"
attributes(nom)$names[attributes(nom)$names=="Rdamcrit"]<-"Tumor burden"
attributes(nom)$names[attributes(nom)$names=="Breslow (SNstatus=Positive)"]<-"Breslow + positive SN"
attributes(nom)$names[attributes(nom)$names=="Breslow (SNstatus=Negative)"]<-"Breslow + negative SN"
# save width: 1000, height: 800
par(mar = rep(0, 4))
plot(nom, total.sep.page=TRUE, col.grid=gray(c(0.8, 0.95)))
nom.0
nom

save(list=c("dd.training", "training.data.to.be.imputed", "m", "imputed.training.data", 
            "last.imputed.training.data.set", "horizon", "S.Rec.5.training", "S.MSM.5.training",
            "f.mi.full.Rec.5", "f.mi.BS.Rec.5", "f.mi.full.MSM.5.refit", "f.mi.BS.MSM.5.refit", "MSM.cal.fact", "c.Breslow", 
            "form.full.Rec.5", "form.BS.Rec.5", "form.full.MSM.5.refit", "form.BS.MSM.5.refit", 
            "rc.Rec", "int.Rec", "h0.Rec"),
     file="Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.Rdata", compress=TRUE)
```