### 2023-09-13 Code/13_Check_Pooled_Data.Rmd
Update

---
title: "Fit models and plots"
author: "Carolien C.H.M. Maas"
date: "January 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fit models

### Load packages, functions, and data 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# erase memory
rm(list = ls(all.names = TRUE))

# for reproducibility
set.seed(100)

# load libraries
library(mice)
library(dplyr)
library(rms)
```

# Check pooled data
```{r, eval=TRUE}
load("Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.RData")

# set datadist
options(datadist="dd.training")

# model based on training data
c.Breslow.all <- c()
for (i in 1:m){
  c.Breslow.all <- c(c.Breslow.all, mean(log(mice::complete(imputed.training.data, i)$Breslow)))
}
c.Breslow <- mean(c.Breslow.all)

# prepare data
load("Z:/Project Melanoom/PaperMelanoma/Data/validation/validation.data.RData")
# combine two raw (without imputation) data sets
pooled.data.to.be.imputed <- rbind(training.data.to.be.imputed, validation.data.to.be.imputed)
# add continent
continent.df <- data.frame(Continent=c(rep("Europe", nrow(complete(imputed.training.data, 5))), 
                                     rep("Australia", nrow(complete(imputed.validation.data, 5)))))

# combine two imputed data sets into one mids object
imputed.pooled.data <- mice::rbind(imputed.training.data, imputed.validation.data)
# add continent variable
imputed.pooled.data <- mice::cbind(imputed.pooled.data, continent.df)

# define the outcome based on last imputed data set
last.imputed.pooled.data.set <- rbind(mice::complete(imputed.training.data, m), mice::complete(imputed.validation.data, m))
S.Rec.training.validation <- survival::Surv(exp(last.imputed.pooled.data.set$logRec.time), last.imputed.pooled.data.set$Recurrence=="recurred")
S.Rec.5.training.validation <- S.Rec.training.validation
S.Rec.5.training.validation[S.Rec.training.validation[,1]>5, 1] <- 5 # set time to 5
S.Rec.5.training.validation[S.Rec.training.validation[,1]>5, 2] <- 0 # set censoring
```

```{r, eval=TRUE}
# model is based on training data
# plus a variable for continent
# model is fitted on pooled data
form.BS.Rec.5.pooled.continent <- S.Rec.5.training.validation ~
  SNstatus +
  Age.SN +
  Ulceration +
  Loc_CAT +
  I(log(Breslow)-c.Breslow) +
  log(Rdamcrit) +
  SNstatus:I(log(Breslow)-c.Breslow) +
  Continent*(SNstatus + Age.SN + Ulceration + Loc_CAT + I(log(Breslow)-c.Breslow) + log(Rdamcrit))

# fit model
f.mi.full.pooled.continent <- Hmisc::fit.mult.impute(form.BS.Rec.5.pooled.continent, cph, xtrans=imputed.pooled.data,
                                 data=pooled.data.to.be.imputed, n.impute=m, pr=FALSE,
                                 fit.reps=TRUE, fitargs=list(y=TRUE, x=TRUE, se.fit=TRUE))
anova(f.mi.full.pooled.continent)
```
```{r, eval=TRUE}
# model is based on training data
# plus a variable for continent
# model is fitted on pooled data
form.BS.Rec.5.pooled <- S.Rec.5.training.validation ~
  SNstatus +
  Age.SN +
  Ulceration +
  Loc_CAT +
  I(log(Breslow)-c.Breslow) +
  log(Rdamcrit) +
  SNstatus:I(log(Breslow)-c.Breslow) +
  Continent

# fit model
f.mi.full.pooled <- Hmisc::fit.mult.impute(form.BS.Rec.5.pooled, cph, xtrans=imputed.pooled.data,
                                 data=pooled.data.to.be.imputed, n.impute=m, pr=FALSE,
                                 fit.reps=TRUE, fitargs=list(y=TRUE, x=TRUE, se.fit=TRUE))
summary(f.mi.full.pooled, Continent="Europe")
anova(f.mi.full.pooled)
```

```{r, eval=TRUE}
# Coefficients with CI
# Negative results
neg.results.BS <- summary(f.mi.full.pooled, SNstatus="Negative", Breslow=exp(c.Breslow), Loc_CAT="arm", Continent="Europe")
neg.table.BS <- rbind(rep(NA, 3), # SN status
                      exp(neg.results.BS[c("Age.SN",
                                           "Ulceration - Yes:No",
                                           "Loc_CAT - leg:arm",
                                           "Loc_CAT - trunk:arm",
                                           "Loc_CAT - head & neck:arm",
                                           "Breslow"), c("Effect", "Lower 0.95", "Upper 0.95")]),
                      rep(NA, 3), # Rdamcrit
                      exp(neg.results.BS["Continent - Australia:Europe", c("Effect", "Lower 0.95", "Upper 0.95")])) 
# add CI
neg.table.BS <- cbind(sprintf("%.2f", neg.table.BS[, 1]),
                      paste0("[", sprintf("%.2f", neg.table.BS[, 2]), "; ", sprintf("%.2f", neg.table.BS[, 3]), "]"))

# Positive results
pos.results.BS <- summary(f.mi.full.pooled, SNstatus="Positive", Breslow=exp(c.Breslow), Loc_CAT="arm", Continent="Europe")
pos.table.BS <- rbind(exp(neg.results.BS["SNstatus - Positive:Negative", c("Effect", "Lower 0.95", "Upper 0.95")]),
                      exp(pos.results.BS[c("Age.SN",
                                           "Ulceration - Yes:No",
                                           "Loc_CAT - leg:arm",
                                           "Loc_CAT - trunk:arm",
                                           "Loc_CAT - head & neck:arm",
                                           "Breslow",
                                           "Rdamcrit",
                                           "Continent - Australia:Europe"), c("Effect", "Lower 0.95", "Upper 0.95")]))
# add CI
pos.table.BS <- cbind(sprintf("%.2f", pos.table.BS[, 1]),
                      paste0("[", sprintf("%.2f", pos.table.BS[, 2]), "; ", sprintf("%.2f", pos.table.BS[, 3]), "]"))

# combine postive with negative
BS.table <- cbind(neg.table.BS, pos.table.BS)

BS.table <- rbind(BS.table[1:3,],
                    rep(NA, 4),
                    c("(ref)", NA, "(ref)", NA),
                    BS.table[4:nrow(BS.table),])

# add blanks
BS.table[BS.table=="[NA; NA]"] <- NA
BS.table[BS.table=="NA"] <- NA
BS.table[is.na(BS.table)] <- ""

# write to Excel
write.table(BS.table,
            file="Z:/Project Melanoom/PaperMelanoma/Results/pooled.final.full.table.Rec.txt",
            row.names=FALSE,
            sep=",")
```