### 2023-09-13 Code/4_TestNonlinearity.Rmd

---
title: "Test non-linearity"
author: "Carolien C.H.M. Maas"
date: "January 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Test non-linearity

### Load packages, functions, and data 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# erase memory
rm(list = ls(all.names = TRUE))

# for reproducibility
set.seed(100)

# load libraries
library(foreign)
library(mice)
library(survminer)
library(prodlim)
library(rms)
library(Hmisc)
library(png)
# remotes::install_github("CHMMaas/PredictionTools")
library(PredictionTools)  # to plot calibration
# import helper functions
source("Z:/Project Melanoom/PaperMelanoma/Code/PaperMelanomaFunctions.R")
```

# Test non-lineairities
### Test if Age should be modelled non-linearly
```{r, eval=TRUE}
load("Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.RData")
options(datadist="dd.training")

form.linear.full.Rec <- S.Rec.5.training ~
  SNstatus +
  Sex +
  Age.SN +
  Ulceration +
  Loc_CAT +
  Histology +
  Breslow +
  multiple.fields +
  Tot_SNs_neg +
  Tot_SNs_pos +
  Rdamcrit +
  Dewar +
  SNstatus:Sex +
  SNstatus:Age.SN +
  SNstatus:Ulceration +
  SNstatus:Loc_CAT +
  SNstatus:Histology +
  SNstatus:Breslow +
  SNstatus:multiple.fields
form.nonlinear.age.full.Rec <- S.Rec.5.training ~
  SNstatus +
  Sex +
  rcs(Age.SN) +
  Ulceration +
  Loc_CAT +
  Histology +
  Breslow +
  multiple.fields +
  Tot_SNs_neg +
  Tot_SNs_pos +
  Rdamcrit +
  Dewar +
  SNstatus:Sex +
  SNstatus:rcs(Age.SN) +
  SNstatus:Ulceration +
  SNstatus:Loc_CAT +
  SNstatus:Histology +
  SNstatus:Breslow +
  SNstatus:multiple.fields
linear.full.Rec <- Hmisc::fit.mult.impute(form.linear.full.Rec, cph, 
                                          xtrans=imputed.training.data,
                                          data=training.data.to.be.imputed,
                                          n.impute=m, pr=FALSE,
                                          fit.reps=TRUE, 
                                          fitargs=list(y=TRUE, x=TRUE, 
                                                       se.fit=TRUE))
nonlinear.age.full.Rec <- Hmisc::fit.mult.impute(form.nonlinear.age.full.Rec, cph, xtrans=imputed.training.data,
                                                 data=training.data.to.be.imputed,n.impute=m, pr=FALSE,
                                                 fit.reps=TRUE, 
                                                 fitargs=list(y=TRUE, x=TRUE, 
                                                       se.fit=TRUE))
# distribution of Age
hist(last.imputed.training.data.set$Age.SN)
# linear model
linear.full.Rec$loglik[2]
# using log(Age) gives slightly better log likelihood
nonlinear.age.full.Rec$loglik[2]
```

### Test if Breslow should be modelled non-linearly
```{r, eval=TRUE}
form.nonlinear.Breslow.full.Rec <- S.Rec.5.training ~
  SNstatus +
  Sex +
  Age.SN +
  Ulceration +
  Loc_CAT +
  Histology +
  rcs(Breslow) +
  multiple.fields +
  Tot_SNs_neg +
  Tot_SNs_pos +
  Rdamcrit +
  Dewar +
  SNstatus:Sex +
  SNstatus:Age.SN +
  SNstatus:Ulceration +
  SNstatus:Loc_CAT +
  SNstatus:Histology +
  SNstatus:rcs(Breslow) +
  SNstatus:multiple.fields
nonlinear.Breslow.full.Rec <- Hmisc::fit.mult.impute(form.nonlinear.Breslow.full.Rec, cph, xtrans=imputed.training.data,
                                                     data=training.data.to.be.imputed,n.impute=m, pr=FALSE,
                                                     fit.reps=TRUE, 
                                                     fitargs=list(y=TRUE, x=TRUE, 
                                                       se.fit=TRUE))
# distribution of Breslow
hist(last.imputed.training.data.set$Breslow)
# linear model
linear.full.Rec$loglik[2]
# using log(Breslow) gives better log likelihood
nonlinear.Breslow.full.Rec$loglik[2]
# test
anova(nonlinear.Breslow.full.Rec)
```

### Test if Rdamcrit should be modelled non-linearly
```{r, eval=TRUE}
form.nonlinear.Rdamcrit.full.Rec <- S.Rec.5.training ~
  SNstatus +
  Sex +
  Age.SN +
  Ulceration +
  Loc_CAT +
  Histology +
  Breslow +
  multiple.fields +
  Tot_SNs_neg +
  Tot_SNs_pos +
  rcs(Rdamcrit) +
  Dewar +
  SNstatus:Sex +
  SNstatus:Age.SN +
  SNstatus:Ulceration +
  SNstatus:Loc_CAT +
  SNstatus:Histology +
  SNstatus:Breslow +
  SNstatus:multiple.fields
nonlinear.Rdamcrit.full.Rec <- Hmisc::fit.mult.impute(form.nonlinear.Rdamcrit.full.Rec, cph, xtrans=imputed.training.data,
                                                      data=training.data.to.be.imputed,n.impute=m, pr=FALSE,
                                                      fit.reps=TRUE, 
                                                      fitargs=list(y=TRUE, x=TRUE, 
                                                       se.fit=TRUE))
# distribution of Rdamcrit
hist(last.imputed.training.data.set$Rdamcrit)
# linear model
linear.full.Rec$loglik[2]
# using log(Rdamcrit) gives better log likelihood
nonlinear.Rdamcrit.full.Rec$loglik[2]
```