### 2023-10-06 Code/AdditionalSensitivityAnalysis_CheckYears.Rmd

---
title: "Sensitivity Analysis"
author: "Carolien C.H.M. Maas"
date: "January 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fit models

### Load packages, functions, and data
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# erase memory
rm(list = ls(all.names = TRUE))

# for reproducibility
set.seed(100)

library(ggplot2)
library(rms)
```

# Load data 
```{r, eval=TRUE}
load("Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.Rdata")
original.training.data <- foreign::read.spss(file="Z:/Project Melanoom/PaperMelanoma/Data/training/Dataset Predictiemodel.sav",
                                    to.data.frame=TRUE, 
                                    origin="01-01-1970")
load("Z:/Project Melanoom/PaperMelanoma/Data/validation/validation.data.Rdata")
original.validation.data <- foreign::read.spss(file="Z:/Project Melanoom/PaperMelanoma/Data/validation/Validation_cohort_MIA.sav",
                                               to.data.frame=TRUE, 
                                               origin="01-01-1970")
```
# Inclusion development data
```{r, eval=TRUE}
sel.train.data <- original.training.data[original.training.data$EXCLUSIE!="Exclusion", ]
sel.train.data <- sel.train.data[!is.na(sel.train.data$Last_FU),]
sel.train.data$Rec.time <- (sel.train.data$Date_1st_Rec-sel.train.data$SNdate)/(24*60*60)/365.25
sel.train.data <- sel.train.data[!(sel.train.data$Rec.time<0&!is.na(sel.train.data$Rec.time)),]
```
# Describe years
```{r, eval=TRUE}
dates.train <- as.POSIXct(sel.train.data$SNdate, origin="1582-10-15")
dates.val <- as.POSIXct(original.validation.data$SNdate, origin="1582-10-15")
Years.train <- as.numeric(format(dates.train, format="%Y"))
Years.train[Years.train==2014] <- 2013
Years.val <- as.numeric(format(dates.val, format="%Y"))
hist(Years.train, main="Development")
hist(Years.val, main="Validation")
```
# Make plot
```{r, eval=TRUE}
dats <- rbind(data.frame(Years=Years.train, Data="Development"),
              data.frame(Years=Years.val, Data="Validation"))
ggplot(dats, aes(Years, fill=Data))+
  geom_histogram(alpha=1, position="dodge", bins=50)+
  ylab("Frequency")+xlab("Calendar year of sentinel node procedure")+
  theme_set(theme_bw())
```
# Add Years to model linearly
```{r, eval=TRUE}
Years.train.stand <- Years.train-1993

form.full.Rec.5.Years <- S.Rec.5.training ~
  SNstatus +
  Sex +
  Age.SN +
  Ulceration +
  Loc_CAT +
  Histology +
  log(Breslow) +
  multiple.fields +
  Tot_SNs_neg +
  Tot_SNs_pos +
  log(Rdamcrit) +
  Dewar +
  SNstatus:Sex +
  SNstatus:Age.SN +
  SNstatus:Ulceration +
  SNstatus:Loc_CAT +
  SNstatus:Histology +
  SNstatus:log(Breslow) +
  SNstatus:multiple.fields +
  Years.train.stand
f.mi.full.Rec.5.Years <- Hmisc::fit.mult.impute(form.full.Rec.5.Years, cph, 
                                          xtrans=imputed.training.data,
                                          data=training.data.to.be.imputed,
                                          n.impute=m, pr=FALSE, fit.reps=TRUE, 
                                          fitargs=list(y=TRUE, x=TRUE, 
                                                       se.fit=TRUE))

form.BS.Rec.5.Years <- S.Rec.5.training ~
  SNstatus +
  Age.SN +
  Ulceration +
  Loc_CAT +
  I(log(Breslow)-c.Breslow) +
  log(Rdamcrit) +
  SNstatus:I(log(Breslow)-c.Breslow)+
  Years.train.stand
f.mi.BS.Rec.5.Years <- Hmisc::fit.mult.impute(form.BS.Rec.5.Years,cph,
                                        xtrans=imputed.training.data,
                                        data=training.data.to.be.imputed,
                                        n.impute=m, pr=FALSE,fit.reps=TRUE,
                                        fitargs=list(y=TRUE,x=TRUE,se.fit=TRUE))
```

# Melanoma-specific survival
```{r, eval=TRUE}
form.BS.MSM.5.refit.Years <- S.MSM.5.training ~
  SNstatus +
  Age.SN +
  Ulceration +
  Loc_CAT +
  I(log(Breslow)-c.Breslow) +
  log(Rdamcrit) +
  SNstatus:I(log(Breslow)-c.Breslow)+
  Years.train.stand
f.mi.BS.MSM.5.refit.Years <- Hmisc::fit.mult.impute(form.BS.MSM.5.refit.Years,cph,
                                              xtrans=imputed.training.data,
                                              data=training.data.to.be.imputed,
                                              n.impute=m,
                                              pr=FALSE,fit.reps=TRUE,
                                              fitargs=list(y=TRUE,x=TRUE,
                                                           se.fit=TRUE))
```
# Add Years to non-linear
```{r, eval=TRUE}
form.full.Rec.5.Years.nonlinear <- S.Rec.5.training ~
  SNstatus +
  Sex +
  Age.SN +
  Ulceration +
  Loc_CAT +
  Histology +
  log(Breslow) +
  multiple.fields +
  Tot_SNs_neg +
  Tot_SNs_pos +
  log(Rdamcrit) +
  Dewar +
  SNstatus:Sex +
  SNstatus:Age.SN +
  SNstatus:Ulceration +
  SNstatus:Loc_CAT +
  SNstatus:Histology +
  SNstatus:log(Breslow) +
  SNstatus:multiple.fields +
  rcs(Years.train.stand)
f.mi.full.Rec.5.Years.nonlinear <- Hmisc::fit.mult.impute(form.full.Rec.5.Years.nonlinear, cph, 
                                          xtrans=imputed.training.data,
                                          data=training.data.to.be.imputed,
                                          n.impute=m, pr=FALSE, fit.reps=TRUE, 
                                          fitargs=list(y=TRUE, x=TRUE, 
                                                       se.fit=TRUE))

form.BS.Rec.5.Years.nonlinear <- S.Rec.5.training ~
  SNstatus +
  Age.SN +
  Ulceration +
  Loc_CAT +
  I(log(Breslow)-c.Breslow) +
  log(Rdamcrit) +
  SNstatus:I(log(Breslow)-c.Breslow)+
  rcs(Years.train.stand)
f.mi.BS.Rec.5.Years.nonlinear <- Hmisc::fit.mult.impute(form.BS.Rec.5.Years.nonlinear,cph,
                                        xtrans=imputed.training.data,
                                        data=training.data.to.be.imputed,
                                        n.impute=m, pr=FALSE,fit.reps=TRUE,
                                        fitargs=list(y=TRUE,x=TRUE,se.fit=TRUE))

form.BS.MSM.5.refit.Years.nonlinear <- S.MSM.5.training ~
  SNstatus +
  Age.SN +
  Ulceration +
  Loc_CAT +
  I(log(Breslow)-c.Breslow) +
  log(Rdamcrit) +
  SNstatus:I(log(Breslow)-c.Breslow)+
  rcs(Years.train.stand)
f.mi.BS.MSM.5.refit.Years.nonlinear <- Hmisc::fit.mult.impute(form.BS.MSM.5.refit.Years.nonlinear,cph,
                                              xtrans=imputed.training.data,
                                              data=training.data.to.be.imputed,
                                              n.impute=m,
                                              pr=FALSE,fit.reps=TRUE,
                                              fitargs=list(y=TRUE,x=TRUE,
                                                           se.fit=TRUE))
```
# Make Table
```{r,eval=TRUE}
results.Years <- c()
for (outcome in c("Rec.5", "MSM.5.refit")){
  if (outcome=="Rec.5"){
    models <- c("full", "BS")
  } else{
      models <- c("BS")
  }

  for (model in models){
    form <- eval(parse(text=paste0("f.mi.", model, ".", outcome)))
    results.Years <- rbind(results.Years, c(paste(model, outcome, "- no years"), 
                                            sprintf("%.1f", AIC(form)), "", ""))

    for (linearity in c("Years", "Years.nonlinear")){
        form.years <- eval(parse(text=paste0("f.mi.", model, ".", outcome, ".", linearity)))
        LR.Years <- lrtest(form, form.years)
        results.Years <- rbind(results.Years, c(paste(model, outcome, "-", linearity),
                                                sprintf("%.1f", AIC(form.years)), 
                                                sprintf("%.1f", as.numeric(LR.Years$stats["L.R. Chisq"])), 
                                                sprintf("%.3f", as.numeric(LR.Years$stats["P"]))))
    }
  }
}
openxlsx::write.xlsx(as.data.frame(results.Years),
            rowNames=FALSE,
            file="Z:/Project Melanoom/PaperMelanoma/Results/Review/SA.Years.xlsx")
```