### 2023-11-23 Code/0_PrepareValidationData.Rmd
Final commit

---
title: "Prepare Validation Data"
author: "Carolien C.H.M. Maas"
date: "November 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Validation data preparation

### Load packages, functions, and data 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# erase memory
rm(list = ls(all.names = TRUE))

# for reproducibility
set.seed(100)

# load libraries
library(foreign)
library(rms)
source("Z:/Project Melanoom/PaperMelanoma/Code/PaperMelanomaFunctions.R")

# load data file
original.validation.data <- foreign::read.spss(file="Z:/Project Melanoom/PaperMelanoma/Data/validation/Validation_cohort_MIA.sav",
                                               to.data.frame=TRUE, origin="01-01-1970")
cat("Number of patients in data set:", nrow(original.validation.data), "\n")

# remove space after Negative
levels(original.validation.data$SNstatus)[levels(original.validation.data$SNstatus)=="Negative "] <- "Negative"

# exclude 4 patients without Status (don"t know if they alive or dead)
cat("Number of patients without Status:", sum(original.validation.data$Status=="Unknown              "), "\n")
sel.val.data <- original.validation.data[original.validation.data$Status!="Dead, MCC            ",]
# exculde 1 patient with cause of death MCC
sel.val.data <- sel.val.data[sel.val.data$Status!="Unknown              ",]
cat("Number of included patients:", nrow(sel.val.data))
```

### Recode
#### Use age at SN instead of age at diagnosis
```{r, eval=TRUE}
sel.val.data$Age.SN <- floor((sel.val.data$SNdate-sel.val.data$DOB)/(24*60*60)/365.25)
```

#### Location
Merge location other with trunk
```{r, eval=TRUE}
sel.val.data$Loc_CAT[sel.val.data$Loc_CAT=="other"] <- "trunk"
```

Merge LMM other with other
```{r, eval=TRUE}
sel.val.data$Histology[sel.val.data$Histology=="LLM"] <- "other" # LLM and LMM is the same
```

#### Define multiple fields
```{r, eval=TRUE}
sel.val.data$multiple.fields <- as.numeric(!is.na(sel.val.data$Tot_nr_SNs)&sel.val.data$Tot_nr_SNs>1)
```

#### Number of positive and negative nodes
Check: there are no negative patients that have positive nodes
```{r, eval=TRUE}
sum(!is.na(sel.val.data$Tot_SNs_pos)&sel.val.data$Tot_SNs_pos>0&sel.val.data$SNstatus=="Negative")
```

Create number of negative nodes removed
```{r, eval=TRUE}
sel.val.data$Tot_nr_SNs <- sel.val.data$Tot_SNs_neg + sel.val.data$Tot_SNs_pos
```

607 positive SN patients have negative nodes
```{r, eval=TRUE}
sum(!is.na(sel.val.data$Tot_SNs_neg)&sel.val.data$Tot_SNs_neg>0&sel.val.data$SNstatus=="Positive")
```

#### SN tumour burden
43 patients with positive SN have SN tumour burden of exactly 1
```{r, eval=TRUE}
sum(sel.val.data$Rdamcrit==1&sel.val.data$SNstatus=="Positive"&!is.na(sel.val.data$Rdamcrit))
```

Set Rdamcrit of negative SN patients equal to 0, SNstatus will capture the effect of these patients
```{r, eval=TRUE}
sel.val.data$Rdamcrit[sel.val.data$SNstatus=="Negative"] <- 1 # since log(1) = 0
```

#### Location of metastasis
Set Dewar for negative patients to subcap to merge them with the reference category, SN status will capture the effect of Negative patients
```{r, eval=TRUE}
sel.val.data$Dewar <- sel.val.data$Subcap
sel.val.data$Dewar[sel.val.data$SNstatus=="Negative"] <- "subcap"

# 339 patients have subcap with positive SN
sum(sel.val.data$SNstatus=="Positive"&sel.val.data$Dewar=="subcap"&!is.na(sel.val.data$Dewar))
```
#### Recurrence 
Set recurrence indicator
```{r, eval=TRUE}
# Recurrence if patient has Date_1st_Rec, Alive with melanoma, or dead with melanoma
sel.val.data$Recurrence <- as.numeric(!is.na(sel.val.data$Date_1st_Rec)) # n=1352
# sel.val.data$Recurrence <- as.numeric(sel.val.data$Status=="Alive with Melanoma  " | sel.val.data$Status=="Dead, melanoma       " | !is.na(sel.val.data$Date_1st_Rec)) # n=1352
sel.val.data$Recurrence <- factor(sel.val.data$Recurrence, labels=c("no recurrence", "recurred"))
```

Define time since diagnosis in years (follow-up time)
```{r, eval=TRUE}
sel.val.data$FU.time <- (sel.val.data$Last_FU-sel.val.data$SNdate)/(24*60*60)/365.25
# median: 4.1629021
```

Define recurrence time in years
```{r, eval=TRUE}
sel.val.data$Recurrence.time <- (sel.val.data$Date_1st_Rec-sel.val.data$SNdate)/(24*60*60)/365.25
```

Check: no patients with Rec date before SN date
```{r, eval=TRUE}
sum(!(!(sel.val.data$Recurrence.time<0&!is.na(sel.val.data$Recurrence.time))))
```
# Define death by any cause
```{r, eval=TRUE}
sel.val.data$dead <- as.numeric(stringr::str_detect(sel.val.data$Status, "Dead"))
sel.val.data$dead <- factor(sel.val.data$dead, labels=c("No", "Yes"))
``` 
<!-- Set recurrence time equal to follow-up time for those without recurrence -->
<!-- ```{r, eval=TRUE} -->
<!-- sel.val.data$Recurrence.time[sel.val.data$Recurrence=="no recurrence"|is.na(sel.val.data$SNdate)]<-sel.val.data$FU.time[sel.val.data$Recurrence=="no recurrence"|is.na(sel.val.data$SNdate)] -->

<!-- sel.val.data$Recurrence.time[sel.val.data$Recurrence=="recurrence"&is.na(sel.val.data$Date_1st_Rec)]<-NA -->
<!-- table(sel.val.data$Status, sel.val.data$Recurrence) -->
<!-- ``` -->
Create composite Recurrence
```{r, eval=TRUE}
# event = 1 if recurrence or death, whichever occurs first
sel.val.data$Rec.composite <- ifelse(sel.val.data$Recurrence=="recurred" | sel.val.data$dead=="Yes", 1, 0)

# if patient has recurrence, set time to Recurrence.time
# if patient has no recurrence, but died, set time to FU.time
# if patient has no recurrence, and did not die, set to FU.time
sel.val.data$Rec.composite.time <- ifelse(sel.val.data$Recurrence=="recurred", 
       sel.val.data$Recurrence.time,
       sel.val.data$FU.time)
```
#### Melanoma specific mortality (MSM)
```{r, eval=TRUE}
table(sel.val.data$Status, sel.val.data$dead)
sel.val.data$MSM <- as.numeric(sel.val.data$Status=="Dead, melanoma       ")
```
#### Mitosis
Define Mitosis as Yes/No
```{r, eval=TRUE}
sel.val.data$Mitosis <- as.numeric(sel.val.data$Mitosis)
```

#### Delete unused levels
```{r, eval=TRUE}
sel.val.data <- droplevels(sel.val.data)
sel.val.data$Center <- 5
sel.val.data$Center <- factor(sel.val.data$Center, labels = c("Australia"))
```

#### Set levels
```{r, eval=TRUE}
sel.val.data$Ulceration <- factor(sel.val.data$Ulceration, labels=c("No", "Yes"))

sel.val.data$multiple.fields <- factor(sel.val.data$multiple.fields, labels=c("Multiple SN fields", "Other"))

sel.val.data$MSM <- factor(sel.val.data$MSM, labels=c("Other", "Death by melanoma"))
```

Relation between outcome and covariates
```{r, eval=TRUE}
sel.val.data$logRec.composite.time <- log(sel.val.data$Rec.composite.time+1/365.25)
sel.val.data$logFU.time <- log(sel.val.data$FU.time+1/365.25)
```

# Multiple imputation
```{r, eval=TRUE}
variables.to.be.imputed <- c("Center", "SNstatus", "Sex", "Age.SN", "Ulceration",
                             "Loc_CAT", "Histology", "Breslow", "multiple.fields",
                             "Tot_SNs_neg", "Tot_SNs_pos", "Rdamcrit", "Dewar",
                             "Mitosis", "Status",
                             "Rec.composite", "logRec.composite.time", 
                             "MSM", "logFU.time")
validation.data.to.be.imputed <- sel.val.data[, variables.to.be.imputed]
missings.table(validation.data.to.be.imputed)
```

Save distributions of all variables, necessary for Harrell summaries
```{r, eval=TRUE}
dd.validation<-rms::datadist(validation.data.to.be.imputed)
options(datadist="dd.validation")
options(digits=8)
save(list=c("dd.validation", "validation.data.to.be.imputed"),file="Z:/Project Melanoom/PaperMelanoma/Data/validation/validation.data.Rdata", compress=TRUE)
```