### 2023-12-11 Code/12_Validation.Rmd

---
title: "Validate model"
author: "Carolien C.H.M. Maas"
date: "November 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Create calibration plots

### Load packages, functions, and data 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# erase memory
rm(list = ls(all.names = TRUE))

# for reproducibility
set.seed(100)

# load libraries
library(mice)
library(survival)
library(PredictionTools)
source("Z:/Project Melanoom/PaperMelanoma/Code/PaperMelanomaFunctions.R")
```

# Training data
```{r, eval=TRUE}
# load data
load("Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.RData")

# predictions on training data
p.Rec.5.training <- matrix(NA, nrow(training.data.to.be.imputed), m)
p.MSM.5.training <- matrix(NA, nrow(training.data.to.be.imputed), m)
for (i in 1:m){
  lp.train.i <- predict(f.mi.BS.Rec.5,
                      newdata=mice::complete(imputed.training.data, i),
                      type="lp")
  p.Rec.5.training[, i] <- fun.event(h0=h0.Rec, lp=lp.train.i)
  p.MSM.5.training[, i] <- fun.event(h0=h0.MSM, lp=MSM.cal.fact*lp.train.i)
}
```

# Validation data
```{r, eval=TRUE}
load("Z:/Project Melanoom/PaperMelanoma/Data/validation/validation.data.RData")
p.Rec.5.validation <- matrix(NA, nrow(validation.data.to.be.imputed), m)
p.MSM.5.validation <- matrix(NA, nrow(validation.data.to.be.imputed), m)
for (i in 1:m){
  lp.val.i <- predict(f.mi.BS.Rec.5,
                      newdata=mice::complete(imputed.validation.data, i),
                      type="lp")
  p.Rec.5.validation[, i] <- 1-exp(-h0.Rec*exp(lp.val.i))
  p.MSM.5.validation[, i] <- 1-exp(-h0.MSM*exp(MSM.cal.fact*lp.val.i))
}
```

```{r, eval=TRUE}
# calibration curve for recurrence
# TODO: validation based on dd.training or dd.validation?
for (data.type in c("training", "validation")){
  for (outcome in c("Rec.5", "MSM.5")){
    options(datadist=paste0("dd.", data.type))

    S <- eval(parse(text=paste0("S.", outcome, ".", data.type)))
    p <- eval(parse(text=paste0("p.", outcome, ".", data.type)))
    data <- eval(parse(text=paste0(data.type, ".data.to.be.imputed")))

    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/", data.type, ".", outcome, ".All.png"))
    par(mar=rep(0, 4))
    PredictionTools::val.surv.mi(p, S,
                                 g=5, main="All patients", time=NULL, 
                                 lim=c(0,1), dist=TRUE, CI.metrics=TRUE,
                                 optimism.C=ifelse(data.type=="training", optimism, 0))
    grDevices::dev.off()

    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/", data.type, ".", outcome, ".Positive.png"))
    par(mar=rep(0, 4))
    PredictionTools::val.surv.mi(p[data$SNstatus=="Positive",],
                                 S[data$SNstatus=="Positive",],
                                 g=5, main="Patients with positive SN", 
                                 time=NULL, lim=c(0,1), dist=TRUE, CI.metrics=TRUE,
                                 optimism.C=ifelse(data.type=="training", optimism, 0))
    grDevices::dev.off()

    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/", data.type, ".", outcome, ".Negative.png"))
    par(mar=rep(0, 4))
    PredictionTools::val.surv.mi(p[data$SNstatus=="Negative",],
                                 S[data$SNstatus=="Negative",],
                                 g=5, main="Patients with negative SN", 
                                 time=NULL, lim=c(0,1), dist=TRUE, CI.metrics=TRUE,
                                 optimism.C=ifelse(data.type=="training", optimism, 0))
    grDevices::dev.off()

    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/", data.type, ".", outcome, ".Neg.Pos.png"), 
        width=32, height=16, units="cm", res=600)
    par(mar=rep(0, 4))
    layout(matrix(1:2, ncol=2, byrow=TRUE))
    for (select.value in c("Positive", "Negative")){
      plot(NA, xlim=0:1, ylim=0:1, xaxt="n", yaxt="n", bty="n")
      img <- png::readPNG(paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/", data.type, ".", outcome, ".", select.value, ".png"))
      rasterImage(img, 0, 0, 1, 1)
    }
    grDevices::dev.off()

    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/", data.type, ".", outcome, ".png"), width=32, height=45, units="cm", res=600)
    par(mar=rep(0, 4))
    layout(matrix(1:2, ncol=1, nrow=2, byrow=TRUE), heights=c(2, 1), widths=c(2, 2))
    for (select.value in c("All", "Neg.Pos")){
      plot(NA, xlim=0:1, ylim=0:1, xaxt="n", yaxt="n", bty="n")
      img <- png::readPNG(paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/", data.type, ".", outcome, ".", select.value, ".png"))
      rasterImage(img, 0, 0, 1, 1)
    }
    grDevices::dev.off()
  }
}
```