### 2024-01-19 Code/07_CreateCrossValidatedCalibrationPlots.Rmd

---
title: "Create calibration plots"
author: "Carolien C.H.M. Maas"
date: "November 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Create calibration plots

### Load packages, functions, and data 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# erase memory
rm(list = ls(all.names = TRUE))

# for reproducibility
set.seed(100)

# load libraries
library(Hmisc)
library(survival)
library(rms)
library(mice)
library(grDevices)
# remotes::install_github("CHMMaas/PredictionTools")
library(PredictionTools)  # to plot calibration
source("Z:/Project Melanoom/PaperMelanoma/Code/PaperMelanomaFunctions.R")
```

# Cross-validation recurrence 5-years
```{r, eval=TRUE}
load("Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.RData")
options(datadist="dd.training")

Centers <- sort(as.character(unique(training.data.to.be.imputed$Center)))
Centers.anonimized <- c("A", "B", "C", "D")
cuts <- 5
for (model in c("Rec.5", "MSM.5")){
  if (model == "Rec.5"){
    S.full <- S.Rec.training
    S.model <- S.Rec.5.training
  }
  if (model == "MSM.5"){
    S.full <- S.MSM.training
    S.model <- S.MSM.5.training
  }

  for (type in c("full", "BS")){
    for (j in 1:4){
      S.j<-S.model[training.data.to.be.imputed$Center==Centers[j],]
      S.full.j<-S.full[training.data.to.be.imputed$Center==Centers[j],]
      S.notj<-S.model[training.data.to.be.imputed$Center!=Centers[j],]
      S.Rec.5.notj<-S.Rec.5.training[training.data.to.be.imputed$Center!=Centers[j],]

      form.model<-eval(parse(text=paste0("form.", type, ".Rec.5")))
      form.notj<-update(form.model, S.Rec.5.notj  ~ . )
      f.mi.Rec.notj<-Hmisc::fit.mult.impute(form.notj,cph,
                                            xtrans=imputed.training.data,
                                            data=training.data.to.be.imputed,
                                            n.impute=m,pr=FALSE,fit.reps=TRUE,
                                            fitargs=list(y=TRUE, x=TRUE, se.fit=TRUE),
                                            sub=training.data.to.be.imputed$Center!=Centers[j])

      for (select.value in c("Positive", "Negative", "All")){
        # identify those patients that are selected
        if (select.value == "All"){
          select.patients <- TRUE
        } else if (select.value != "All"){
          select.patients <- training.data.to.be.imputed[training.data.to.be.imputed$Center==Centers[j],]$SNstatus==select.value
        }

        # calibration with metrics
        # S.full <- S.full.j[select.patients,]
        S <- S.full.j[select.patients,]
        p <- matrix(NA, nrow(S), m)
        for (i in 1:m){
          data.i <- mice::complete(imputed.training.data, i)
          lp.i <- predict(f.mi.Rec.notj,newdata=data.i[training.data.to.be.imputed$Center==Centers[j],], type="lp")

          if (model=="MSM.5"){
            lp.notj<-f.mi.Rec.notj$linear
            f.MSM.notj<-rms::cph(S.notj~lp.notj, data=data.i[which(data.i$Center!=Centers[j]),],y=TRUE,x=TRUE)
            lp.i<- predict(f.MSM.notj,newdata=lp.i)
          }

          if (select.value != "All"){
            lp.i <- lp.i[select.patients]
          }

          # baseline hazard
          if (model=="Rec.5"){
            f.basehaz.i <- basehaz(cph(form.notj, data=data.i[which(data.i$Center!=Centers[j]),], se.fit=TRUE, x=TRUE, y=TRUE))
          }
          else if (model=="MSM.5"){
            f.basehaz.i <- basehaz(f.MSM.notj)
          }
          h0.i <- f.basehaz.i$hazard[f.basehaz.i$time==max(f.basehaz.i$time[f.basehaz.i$time<=horizon])]

          # probabilities
          p.i <- fun.event(lp=lp.i, h0=h0.i)
          p[, i] <- p.i
        }

        # omit the observation with predicted probability = 1
        # omit <- unique(c(which(p[,1]==1), which(p[,2]==1), which(p[,3]==1), which(p[,4]==1), which(p[,5]==1)))
        # if (length(omit)>0){
        #   cat("Omit:", omit, "\n")
        #   p <- p[-omit,]
        #   S.full <- S.full[-omit,]
        # }

        png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/", model, ".", select.value, ".", type, ".", Centers[j], ".png"),
            width=1800, height=1800, units="px", res=300)
        out <- PredictionTools::val.surv.mi(p, S, g=5, 
                                            main=paste("Center", Centers.anonimized[j]),
                                            time=5, lim=c(0,1), dist=TRUE,
                                            CI.metrics=TRUE, n.iter=40)
        grDevices::dev.off()
      }
    }
  }
}

# combine calibration plots into one
for (select.value in c("All", "Positive", "Negative")){ 
  for (y in c("Rec.5", "MSM.5")){
    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/", y, ".", select.value, ".png"), width=16, height=16, units="cm", res=300)
    par(mar=rep(0, 4))
    layout(matrix(1:4, ncol=2, byrow=TRUE))
    Centers.val <- rep(1:4, 3)
    for (i in 1:4){
      plot(NA, xlim=0:1, ylim=0:1, xaxt="n", yaxt="n", bty="n")
      img <- png::readPNG(paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/", y, ".", select.value, ".BS.", Centers[Centers.val[i]], ".png"))
      rasterImage(img, 0, 0, 1, 1)
    }
    grDevices::dev.off()
  }
}
```