### 2024-01-19 Code/09_CreateRiskScoreHistogram.Rmd

---
title: "Create histogram with risk scores"
author: "Carolien C.H.M. Maas"
date: "November 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Create calibration plots

### Load packages, functions, and data 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# erase memory
rm(list = ls(all.names = TRUE))

# for reproducibility
set.seed(100)

# load libraries
library(grDevices)
library(stats)
library(survival)
source("Z:/Project Melanoom/PaperMelanoma/Code/PaperMelanomaFunctions.R")
```

# Risk score distribution
```{r, eval=TRUE}
load("Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.RData")
options(datadist="dd.training")

# baseline hazard of MSM
h0.MSM.all <- c()
lp.MSM.all <- c()
for (i in 1:m){
  lp.i<-predict(f.mi.BS.Rec.5,
                newdata=mice::complete(imputed.training.data,i), 
                type="lp")
  lp.MSM.all <- cbind(lp.MSM.all, lp.i)
  f.MSM.basehaz.i <- basehaz(rms::cph(S.MSM.5.training~lp.i,y=TRUE,x=TRUE))
  h0.MSM.i <- f.MSM.basehaz.i$hazard[f.MSM.basehaz.i$time==max(f.MSM.basehaz.i$time[f.MSM.basehaz.i$time<=horizon])]
  h0.MSM.all <- c(h0.MSM.all, h0.MSM.i)
}
h0.MSM <- mean(h0.MSM.all)
lp.MSM <- rowMeans(lp.MSM.all)

# TODO: Is this correct?
lp.MSM.all[3,]

# lp range MSM
rc.MSM <- MSM.cal.fact*rc.Rec
int.MSM <- MSM.cal.fact*int.Rec
max.range <- 17
points.range <- 0:max.range
lp.points.Rec.5 <- int.Rec+rc.Rec*points.range
lp.points.MSM.5 <- int.MSM+rc.MSM*points.range

# Recurrence lp scores
Breslow.trunc <- last.imputed.training.data.set$Breslow
Breslow.trunc[last.imputed.training.data.set$Breslow<0.1] <- .1
Breslow.trunc[last.imputed.training.data.set$Breslow>7] <- 7
lp.Rec <- predict(f.mi.BS.Rec.5,newdata=data.frame(SNstatus=last.imputed.training.data.set$SNstatus,
                           Age.SN=last.imputed.training.data.set$Age.SN,
                           Ulceration=last.imputed.training.data.set$Ulceration,
                           Loc_CAT=last.imputed.training.data.set$Loc_CAT,
                           Breslow=Breslow.trunc,
                           Rdamcrit=last.imputed.training.data.set$Rdamcrit))
mean(lp.Rec)
score.Rec <- round((lp.Rec-int.Rec)/rc.Rec, 0)
max.breaks <- max.range+1
h.Rec <- hist(score.Rec, 
              plot=FALSE,
              breaks=0:max.breaks,
              right=FALSE, 
              beside=TRUE)

# score table
data.frame(table(score.Rec)/length(score.Rec))
```

# Plot of risk score distribution
```{r, eval=TRUE, message=FALSE}
h.plot <- h.Rec
h.plot$density <- 100*h.Rec$density
p.Rec <- fun.event(lp=lp.points.Rec.5, h0=h0.Rec)
p.MSM <- fun.event(lp=lp.points.MSM.5, h0=h0.MSM)
x.lim <- c(0, max.range+1)
y.lim <- c(0, 100)
y.lim.hist <- c(0, 25)
png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Risk.distribution.png"),
    width=1800, height=1800, units="px", res=300)
par(mar = c(5,5,2,5))
plot(h.plot,freq=FALSE,axes=FALSE, xlab=NA, xlim=x.lim, ylab=NA,ylim=y.lim.hist,main=NA,col="white")
abline(h=(0:30)*1,col="light grey",lty=2)
axis(side = 4)
mtext(side = 4, line = 3, "Risk score distribution (%)")
par(new = TRUE)
plot(h.plot,freq=FALSE,axes=FALSE, xlab=NA, xlim=x.lim, ylab=NA,ylim=y.lim.hist,main=NA,col="white")
par(new = TRUE)
plot(points.range,p.Rec*100,type="b",lwd=1,xlim=x.lim,ylim=y.lim,pch=15, xlab="Risk score",ylab="Predicted 5-year risk (%)")
points(points.range,p.MSM*100,type="b",lwd=1,xlim=x.lim,ylim=y.lim,pch=16)
legend(x=0,y=100,legend=c("Recurrence or death","Melanoma-specific mortality"),pch=c(15,16),cex=.9,bg="white")
grDevices::dev.off()
```

```{r, eval=TRUE, echo=FALSE}
# table of risk distribution
cbind(points.range, round(p.Rec*100, 1), round(p.MSM*100, 1))[c(1, 9, 10, 12, 13),]
```
# Save for Rshiny app
```{r, eval=TRUE}
coef.Rec <- stats::coef(f.mi.BS.Rec.5)
center.Rec <- f.mi.BS.Rec.5$center

save(rc.Rec, int.Rec, f.mi.BS.Rec.5, MSM.cal.fact, max.range,
     file="Z:/Project Melanoom/PaperMelanoma/Results/model.Rdata", 
     compress=TRUE)

save(list=c("dd.training", "training.data.to.be.imputed", "m", 
            "imputed.training.data", "last.imputed.training.data.set", 
            "horizon", 
            "S.Rec.training", "S.MSM.training",
            "S.Rec.5.training", "S.MSM.5.training",
            "f.mi.full.Rec.5", "f.mi.BS.Rec.5", 
            "f.mi.full.MSM.5.refit", "f.mi.BS.MSM.5.refit", 
            "MSM.cal.fact", "c.Breslow", "optimism",
            "form.full.Rec.5", "form.BS.Rec.5", 
            "form.full.MSM.5.refit", "form.BS.MSM.5.refit", 
            "rc.Rec", "int.Rec", "h0.Rec", "lp.points.Rec.5",
            "rc.MSM", "rc.Rec", "h0.MSM", "lp.points.MSM.5",
            "lp.Rec", "lp.MSM", "max.breaks", "h.Rec",
            "h.plot", "y.lim.hist", "x.lim", "coef.Rec", "center.Rec"),
     file="Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.Rdata", compress=TRUE)
```

# Risk distribution and scores with ggplot
```{r, eval=TRUE}
library(ggplot2)
last.imputed.training.data.set$score.Rec <- score.Rec # TODO: over 5 imputations?

plot <- ggplot(data=data.frame(), aes(x=0:max.range, y=0:100))+
  geom_bar(data=last.imputed.training.data.set, 
           aes(x=score.Rec, y=(..count..)/sum(..count..)*400, fill=SNstatus))+
  geom_line(aes(x=points.range, y=p.Rec*100))+
  geom_point(aes(x=points.range, y=p.Rec*100, alpha="Recurrence or death"), shape=15)+
  geom_line(aes(x=points.range, y=p.MSM*100))+
  geom_point(aes(x=points.range, y=p.MSM*100, alpha="Melanoma-specific mortality"), shape=19)+
  scale_y_continuous(limits=c(0, 100), minor_breaks=seq(0, 100, 5),
                     sec.axis=sec_axis(~./4, name="Risk score distribution (%)"))+
  ylab("Predicted 5-year risk (%)")+
  xlab("Risk score")+
  theme_bw()+
  scale_fill_discrete(name=NULL,
                      labels=c("Negative SN status", "Positive SN status"))+
  scale_alpha_manual(name=NULL, 
                     values=c(15, 19),
                     breaks=c("Recurrence or death", "Melanoma-specific mortality"),
                     guide=guide_legend(override.aes=list(linetype=c(0, 1), 
                                                          shape=c(15, 19), 
                                                          color="black")))+
  theme(legend.position=c(0.25, 0.80),
        legend.title=element_blank())
ggsave(file="Z:/Project Melanoom/PaperMelanoma/Results/Risk.distribution.pdf",
       width=15, height=15, units="cm", dpi=300)
```