### 2024-01-19 Code/10_AJCC.Rmd

---
title: "Calculate AJCC"
author: "Carolien C.H.M. Maas"
date: "November 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Calculate AJCC

### Load packages, functions, and data 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# erase memory
rm(list = ls(all.names = TRUE))

# for reproducibility
set.seed(100)

# load libraries
library(PredictionTools)  # to plot calibration
library(compareC)         # compare AJCC to model
library(survival)
```
# Obtain C-index of AJCC7 and AJCC8
```{r, eval=TRUE}
calculate.AJCC.7.8 <- function(imputed.data=NULL, type=NULL, 
                               S.Rec.5=NULL, S.Rec=NULL){
  ####
  #### AJCC 7
  ####
  TNM.7 <- c()
  AJCC.7 <- c()
  for (i in 1:m){
    # imputed data set i
    imputation.i <- mice::complete(imputed.data, i)

    # calcualte TNM and AJCC for each imputation
    TNM.7.i <- rep("", nrow(imputation.i))
    AJCC.7.i <- rep(0, nrow(imputation.i))
    for (patient in 1:nrow(imputation.i)){
      breslow.patient <- imputation.i[patient, "Breslow"]
      ulceration.patient <- as.character(imputation.i[patient, "Ulceration"])
      if (type=="training"){
        mitosis.patient <- as.character(imputation.i[patient, "Mitosis"])
      } else{
        mitosis.patient <- ifelse(imputation.i[patient, "Mitosis"]<1, "Yes", "No")
      }
      nodes.patient <- imputation.i[patient, "Tot_SNs_pos"]

      snstatus.patient <- imputation.i[patient, "SNstatus"]

      # T stage
      if (breslow.patient <= 1){
        if (ulceration.patient=="No" & mitosis.patient=="No"){
          TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T1a")
        } else if (ulceration.patient=="Yes" | mitosis.patient=="Yes"){
          TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T1b")
        }
      } else if (breslow.patient>1 & breslow.patient <= 2){
        if (ulceration.patient=="No"){
          TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T2a")
        } else if (ulceration.patient=="Yes"){
          TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T2b")
        }
      } else if (breslow.patient>2 & breslow.patient <= 4){
        if (ulceration.patient=="No"){
          TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T3a")
        } else if (ulceration.patient=="Yes"){
          TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T3b")
        }
      } else if (breslow.patient>4){
        if (ulceration.patient=="No"){
          TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T4a")
        } else if (ulceration.patient=="Yes"){
          TNM.7.i[patient] <- paste0(TNM.7.i[patient], "T4b")
        }
      }

      # N stage
      if (nodes.patient == 0){
        TNM.7.i[patient] <- paste0(TNM.7.i[patient], "N0")
      } else if (nodes.patient==1){
        TNM.7.i[patient] <- paste0(TNM.7.i[patient], "N1a")
      } else if (nodes.patient==2 | nodes.patient==3){
        TNM.7.i[patient] <- paste0(TNM.7.i[patient], "N2a")
      } else if (nodes.patient>3){
        TNM.7.i[patient] <- paste0(TNM.7.i[patient], "N3a")
      }

      # M stage for all patients in our data set
      TNM.7.i[patient] <- paste0(TNM.7.i[patient], "M0")

      # determine AJCC.7.i
      if (TNM.7.i[patient]=="T1aN0M0"){
        AJCC.7.i[patient] <- 1.1
      } else if (TNM.7.i[patient]=="T1bN0M0" | TNM.7.i[patient]=="T2aN0M0"){
        AJCC.7.i[patient] <- 1.2
      } else if (TNM.7.i[patient]=="T2bN0M0" | TNM.7.i[patient]=="T3aN0M0"){
        AJCC.7.i[patient] <- 2.1
      } else if (TNM.7.i[patient]=="T3bN0M0" | TNM.7.i[patient]=="T4aN0M0"){
        AJCC.7.i[patient] <- 2.2
      } else if (TNM.7.i[patient]=="T4bN0M0"){
        AJCC.7.i[patient] <- 2.3
      } else if (TNM.7.i[patient]=="T1aN1aM0" | TNM.7.i[patient]=="T2aN1aM0" | # N1a
                 TNM.7.i[patient]=="T3aN1aM0" | TNM.7.i[patient]=="T4aN1aM0" |
                 TNM.7.i[patient]=="T1aN2aM0" | TNM.7.i[patient]=="T2aN2aM0" | # N2a
                 TNM.7.i[patient]=="T3aN2aM0" | TNM.7.i[patient]=="T4aN2aM0"){
        AJCC.7.i[patient] <- 3.1
      } else if (TNM.7.i[patient]=="T1bN1aM0" | TNM.7.i[patient]=="T2bN1aM0" | # N1a
                 TNM.7.i[patient]=="T3bN1aM0" | TNM.7.i[patient]=="T4bN1aM0" |
                 TNM.7.i[patient]=="T1bN2aM0" | TNM.7.i[patient]=="T2bN2aM0" | # N2a
                 TNM.7.i[patient]=="T3bN2aM0" | TNM.7.i[patient]=="T4bN2aM0" |
                 TNM.7.i[patient]=="T1aN1bM0" | TNM.7.i[patient]=="T2aN1bM0" | # N1b
                 TNM.7.i[patient]=="T3aN1bM0" | TNM.7.i[patient]=="T4aN1bM0" |
                 TNM.7.i[patient]=="T1aN2bM0" | TNM.7.i[patient]=="T2aN2bM0" | # N2b
                 TNM.7.i[patient]=="T3aN2bM0" | TNM.7.i[patient]=="T4aN2bM0" |
                 TNM.7.i[patient]=="T1aN2cM0" | TNM.7.i[patient]=="T2aN2cM0" | # N2c
                 TNM.7.i[patient]=="T3aN2cM0" | TNM.7.i[patient]=="T4aN2cM0"){
        AJCC.7.i[patient] <- 3.2
      } else {
        AJCC.7.i[patient] <- 3.3
      }
    }
    # save TNM.7 and AJCC.7 for each imputed data set
    TNM.7 <- cbind(TNM.7, TNM.7.i)
    AJCC.7 <- cbind(AJCC.7, AJCC.7.i)
  }

  ####
  #### AJCC8
  ####
  TNM.8 <- c()
  AJCC.8 <- c()
  for (i in 1:m){
    # imputed data set i
    imputation.i <- mice::complete(imputed.data,i)

    # calculate TNM and AJCC for each imputed data set i
    TNM.8.i <- rep("", nrow(imputation.i))
    AJCC.8.i <- rep(0, nrow(imputation.i))
    for (patient in 1:nrow(imputation.i)){
      breslow.patient <- imputation.i[patient, "Breslow"]
      ulceration.patient <- as.character(imputation.i[patient, "Ulceration"])
      mitosis.patient <- as.character(imputation.i[patient, "Mitosis"])
      nodes.patient <- imputation.i[patient, "Tot_SNs_pos"]

      # T stage
      if (breslow.patient <= 1){
        if (ulceration.patient=="No" & breslow.patient < 0.8){
          TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T1a")
        } else {
          TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T1b")
        }
      } else if (breslow.patient>1 & breslow.patient <= 2){
        if (ulceration.patient=="No"){
          TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T2a")
        } else if (ulceration.patient=="Yes"){
          TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T2b")
        }
      } else if (breslow.patient>2 & breslow.patient <= 4){
        if (ulceration.patient=="No"){
          TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T3a")
        } else if (ulceration.patient=="Yes"){
          TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T3b")
        }
      } else if (breslow.patient>4){
        if (ulceration.patient=="No"){
          TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T4a")
        } else if (ulceration.patient=="Yes"){
          TNM.8.i[patient] <- paste0(TNM.8.i[patient], "T4b")
        }
      }

      # N stage
      if (nodes.patient == 0){
        TNM.8.i[patient] <- paste0(TNM.8.i[patient], "N0")
      } else if (nodes.patient==1){
        TNM.8.i[patient] <- paste0(TNM.8.i[patient], "N1a")
      } else if (nodes.patient==2 | nodes.patient==3){
        TNM.8.i[patient] <- paste0(TNM.8.i[patient], "N2a")
      } else if (nodes.patient>3){
        TNM.8.i[patient] <- paste0(TNM.8.i[patient], "N3a")
      }

      # M stage for all patients in our data set
      TNM.8.i[patient] <- paste0(TNM.8.i[patient], "M0")

      # determine AJCC.8.i
      if (TNM.8.i[patient]=="T1aN0M0"){
        AJCC.8.i[patient] <- 1.1
      } else if (TNM.8.i[patient]=="T1bN0M0" | TNM.8.i[patient]=="T2aN0M0"){
        AJCC.8.i[patient] <- 1.2
      } else if (TNM.8.i[patient]=="T2bN0M0" | TNM.8.i[patient]=="T3aN0M0"){
        AJCC.8.i[patient] <- 2.1
      } else if (TNM.8.i[patient]=="T3bN0M0" | TNM.8.i[patient]=="T4aN0M0"){
        AJCC.8.i[patient] <- 2.2
      } else if (TNM.8.i[patient]=="T4bN0M0"){
        AJCC.8.i[patient] <- 2.3
      } else if (TNM.8.i[patient]=="T1aN1aM0" | TNM.8.i[patient]=="T1bN1aM0" | TNM.8.i[patient]=="T2aN1aM0"
                 | TNM.8.i[patient]=="T1aN2aM0" | TNM.8.i[patient]=="T1bN2aM0" | TNM.8.i[patient]=="T2aN2aM0"){
        AJCC.8.i[patient] <- 3.1
      } else if (TNM.8.i[patient]=="T2bN1aM0" | TNM.8.i[patient]=="T3aN1aM0"){
        AJCC.8.i[patient] <- 3.2
      } else if (TNM.8.i[patient]=="T4bN3aM0"){
        AJCC.8.i[patient] <- 3.4
      } else {
        AJCC.8.i[patient] <- 3.3
      }

    }
    # save TNM.8 and AJCC.8 for each imputed data set
    TNM.8 <- cbind(TNM.8, TNM.8.i)
    AJCC.8 <- cbind(AJCC.8, AJCC.8.i)
  }

  for (AJCC.nr in c(7, 8)){  
    AJCC.table <- c()
    for (select.value in c("All", "Positive", "Negative")){
      cindex <- rep(0, m)
      cindex.se <- rep(0, m)
      unoC <- rep(0, m)
      tdAUC <- rep(0, m)
      for (i in 1:m){
        # select the AJCC 7 or 8 for the ith imputed data set
        AJCC.i <- eval(parse(text=paste0("AJCC.", AJCC.nr, "[, i]")))

        # identify those patients that are selected
        if (select.value == "All"){
          select.patients <- TRUE
        } else if (select.value != "All"){
          select.patients <- imputation.i$SNstatus==select.value
        }

        # compute Harrell's C-index
        # add a minus to revert class order: from worst to best
        rc <- Hmisc::rcorr.cens(-AJCC.i[select.patients], S.Rec.5[select.patients,]) 
        cindex[i] <- rc["C Index"]
        cindex.se[i] <- rc["S.D."]/2

        # Uno's C-index
        Uno.data <- cbind(S.Rec.5[select.patients, 1], 
                          S.Rec.5[select.patients, 2], 
                          AJCC.i[select.patients])
        unoC[i] <- survC1::Est.Cval(mydata=Uno.data, 
                                    tau=horizon, 
                                    nofit=TRUE)$Dhat
        if (i == 1){
          unoC.boot <- c()
          n.iter <- 40
          for (itr in 1:n.iter){
            boot <- Uno.data[sample(1:nrow(Uno.data), nrow(Uno.data), replace=TRUE),]
            unoC <- survC1::Est.Cval(mydata=boot,
                                     tau=horizon,
                                     nofit=TRUE)$Dhat
            unoC.boot <- c(unoC.boot, unoC)
          }
          unoC.se <- PredictionTools::Rubin.combine(mean(unoC.boot), sd(unoC.boot))$se
        }

        # time-dependent ROC
        tdAUC.i <- timeROC::timeROC(T=S.Rec[select.patients, 1],
                                       delta=S.Rec[select.patients, 2],
                                       marker=AJCC.i[select.patients],
                                       times=horizon,
                                       cause=1, 
                                       iid=FALSE)

        # time-dependent ROC SE
        if (i == 1){
          tdAUC.i <- timeROC::timeROC(T=S.Rec[select.patients, 1],
                                       delta=S.Rec[select.patients, 2],
                                       marker=AJCC.i[select.patients],
                                       times=horizon,
                                       cause=1, 
                                       iid=TRUE)
          tdAUC.CI <- stats::confint(tdAUC.i)
          tdAUC.se <- as.numeric((as.numeric(tdAUC.CI$CI_AUC[2])/100-as.numeric(tdAUC.i$AUC[2]))/tdAUC.CI$C.alpha)
        }
        tdAUC[i] <- as.numeric(tdAUC.i$AUC[2])
      }
      cindex.mi<-Rubin.combine(cindex, cindex.se)

      # add to Table
      AJCC.table <- rbind(AJCC.table, 
                          c(paste(select.value, "patients"), sprintf("%.2f", cindex.mi$est),
                            paste0("[", sprintf("%.2f", cindex.mi$est+qnorm(.025)*cindex.mi$se),
                                    "; ", sprintf("%.2f", cindex.mi$est+qnorm(.975)*cindex.mi$se), "]"),
                            sprintf("%.2f", mean(unoC)),
                            paste0("[", sprintf("%.2f", mean(unoC)+qnorm(.025)*unoC.se),
                                    "; ", sprintf("%.2f", mean(unoC)+qnorm(.975)*unoC.se), "]"),
                            sprintf("%.2f", mean(tdAUC)),
                            paste0("[", sprintf("%.2f", mean(tdAUC)+qnorm(.025)*tdAUC.se),
                                    "; ", sprintf("%.2f", mean(tdAUC)+qnorm(.975)*tdAUC.se), "]")))
    }
    assign(paste0("AJCC.", AJCC.nr, ".table"), AJCC.table)
  }

  # table of number of AJCC.7 and AJCC.8
  table.AJCC.7 <- c()
  table.AJCC.8 <- c()
  for (i in 1:m){
    table.AJCC.7 <- cbind(table.AJCC.7, as.numeric(table(AJCC.7[,i])))
    table.AJCC.8 <- cbind(table.AJCC.8, as.numeric(table(AJCC.8[,i])))
  }

  if (type=="training"){
    AJCC.7.8.count <- cbind(c("1.1", "1.2", "2.1", "2.2", "2.3", "3.1", "3.2", "3.3", "3.4"),
                          c(round(rowMeans(table.AJCC.7), 0), 0),
                          round(rowMeans(table.AJCC.8)))
  } else if (type=="validation"){
    AJCC.7.8.count <- cbind(c("1.1", "1.2", "2.1", "2.2", "2.3", "3.1", "3.2", "3.3", "3.4"),
                          c(round(rowMeans(table.AJCC.7), 0), 0),
                          c(0, round(rowMeans(table.AJCC.8), 0)))
  }

  return(list(AJCC.7=AJCC.7, AJCC.7.table=AJCC.7.table,
              AJCC.8=AJCC.8, AJCC.8.table=AJCC.8.table, AJCC.7.8.count=AJCC.7.8.count))
}
load("Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.RData")
train.AJCC <- calculate.AJCC.7.8(imputed.data=imputed.training.data, 
                                 S.Rec.5=S.Rec.5.training, 
                                 S.Rec=S.Rec.training, type="training")
```

# Write to Excel
```{r, eval=TRUE}
AJCC.Cindex.table <- rbind(train.AJCC$AJCC.7.table[1,], 
                         train.AJCC$AJCC.8.table[1,],
                         train.AJCC$AJCC.7.table[2,], 
                         train.AJCC$AJCC.8.table[2,],
                         train.AJCC$AJCC.7.table[3,], 
                         train.AJCC$AJCC.8.table[3,])
utils::write.table(AJCC.Cindex.table,
              file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/AJCC.train.txt"),
              row.names=FALSE, col.names=FALSE, sep=",")
```