### 2024-03-11 Code/05_FitModels_save.Rmd

---
title: "Fit models and plots"
author: "Carolien C.H.M. Maas"
date: "November 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fit models

### Load packages, functions, and data 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# erase memory
rm(list = ls(all.names = TRUE))

# for reproducibility
set.seed(100)

# load libraries
library(Hmisc)
library(stats)
library(rms)
library(utils)
library(PredictionTools)
```

# Full model for recurrence
```{r, eval=TRUE}
load("Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.RData")
options(datadist="dd.training")

form.full.Rec.5 <- S.Rec.5.training ~
  SNstatus +
  Sex +
  Age.SN +
  Ulceration +
  Loc_CAT +
  Histology +
  log(Breslow) +
  multiple.fields +
  Tot_SNs_neg +
  Tot_SNs_pos +
  log(Rdamcrit) +
  Dewar +
  SNstatus:Sex +
  SNstatus:Age.SN +
  SNstatus:Ulceration +
  SNstatus:Loc_CAT +
  SNstatus:Histology +
  SNstatus:log(Breslow) +
  SNstatus:multiple.fields
f.mi.full.Rec.5 <- Hmisc::fit.mult.impute(form.full.Rec.5, cph, 
                                          xtrans=imputed.training.data,
                                          data=training.data.to.be.imputed,
                                          n.impute=m, pr=FALSE, fit.reps=TRUE, 
                                          fitargs=list(y=TRUE, x=TRUE, 
                                                       se.fit=TRUE))
stats::anova(f.mi.full.Rec.5)
summary(f.mi.full.Rec.5, Tot_SNs_pos=c(0, 1))
```

# Final model for recurrence selected by backward selection with p = 0.01
```{r, eval=TRUE}
# center Breslow by using all imputated data sets
c.Breslow.all <- c()
for (i in 1:m){
  c.Breslow.all <- c(c.Breslow.all, 
                     mean(log(mice::complete(imputed.training.data, i)$Breslow)))
}
c.Breslow <- mean(c.Breslow.all)

form.BS.Rec.5 <- S.Rec.5.training ~
  SNstatus +
  Age.SN +
  Ulceration +
  Loc_CAT +
  I(log(Breslow)-c.Breslow) +
  log(Rdamcrit) +
  SNstatus:I(log(Breslow)-c.Breslow)
f.mi.BS.Rec.5 <- Hmisc::fit.mult.impute(form.BS.Rec.5,cph,
                                        xtrans=imputed.training.data,
                                        data=training.data.to.be.imputed,
                                        n.impute=m, pr=FALSE,fit.reps=TRUE,
                                        fitargs=list(y=TRUE,x=TRUE,se.fit=TRUE))

# check Cox Proportional Hazards Assumption
PH.assumption <- survival::cox.zph(f.mi.BS.Rec.5)

# backward selection
# round(stats::anova(f.mi.BS.Rec.5)[order(stats::anova(f.mi.BS.Rec.5)[, "P"]),], 3)
summary(f.mi.BS.Rec.5, Breslow=exp(c.Breslow))
coef(f.mi.BS.Rec.5)
stats::anova(f.mi.BS.Rec.5)
```

# Refitted full model for melanoma specific mortality
```{r, eval=TRUE}
form.full.MSM.5.refit <- S.MSM.5.training ~
  SNstatus +
  Sex +
  Age.SN +
  Ulceration +
  Loc_CAT +
  Histology +
  log(Breslow) +
  multiple.fields +
  Tot_SNs_neg +
  Tot_SNs_pos +
  log(Rdamcrit) +
  Dewar +
  SNstatus:Sex +
  SNstatus:Age.SN +
  SNstatus:Ulceration +
  SNstatus:Loc_CAT +
  SNstatus:Histology +
  SNstatus:log(Breslow) +
  SNstatus:multiple.fields
f.mi.full.MSM.5.refit <- Hmisc::fit.mult.impute(form.full.MSM.5.refit,cph,
                                                xtrans=imputed.training.data,
                                                data=training.data.to.be.imputed,
                                                n.impute=m, pr=FALSE,
                                                fit.reps=TRUE,
                                                fitargs=list(y=TRUE,x=TRUE,
                                                             se.fit=TRUE))
summary(f.mi.full.MSM.5.refit, Tot_SNs_pos=c(0, 1))
stats::anova(f.mi.full.MSM.5.refit)
```

# Refitted final model for melanoma specific mortality
```{r, eval=TRUE}
form.BS.MSM.5.refit <- S.MSM.5.training ~
  SNstatus +
  Age.SN +
  Ulceration +
  Loc_CAT +
  I(log(Breslow)-c.Breslow) +
  log(Rdamcrit) +
  SNstatus:I(log(Breslow)-c.Breslow)
f.mi.BS.MSM.5.refit <- Hmisc::fit.mult.impute(form.BS.MSM.5.refit,cph,
                                              xtrans=imputed.training.data,
                                              data=training.data.to.be.imputed,
                                              n.impute=m,
                                              pr=FALSE,fit.reps=TRUE,
                                              fitargs=list(y=TRUE,x=TRUE,
                                                           se.fit=TRUE))
summary(f.mi.BS.MSM.5.refit, Breslow=exp(c.Breslow))
stats::anova(f.mi.BS.MSM.5.refit)
```

# Calibrated final model for melanoma specific mortality
Fit disease specific mortality to the linear predictor of recurrence
```{r, eval=TRUE}
# Combine calibration coefficients for all imputated data sets 
MSM.cal.fact.all <- c()
MSM.cal.fact.SE <- c()
for (i in 1:m){
  lp.i<-predict(f.mi.BS.Rec.5,
                newdata=mice::complete(imputed.training.data,i), 
                type="lp")
  f.mi.i <- rms::cph(S.MSM.5.training~lp.i,y=TRUE,x=TRUE) 
  MSM.cal.fact.all <- c(MSM.cal.fact.all, stats::coef(f.mi.i))
  MSM.cal.fact.SE <- c(MSM.cal.fact.SE, sqrt(diag(vcov(f.mi.i))))
}
MSM.cal.fact.combined <- Rubin.combine(MSM.cal.fact.all, MSM.cal.fact.SE)
MSM.cal.fact <- MSM.cal.fact.combined$est

utils::write.table(paste0(sprintf("%.2f", MSM.cal.fact), " [",
            sprintf("%.2f", MSM.cal.fact.combined$est+qnorm(.025)*MSM.cal.fact.combined$se), "; ",
            sprintf("%.2f", MSM.cal.fact.combined$est+qnorm(.975)*MSM.cal.fact.combined$se), "]"),
            file="Z:/Project Melanoom/PaperMelanoma/Results/calibration.MSS.txt", 
            row.names=FALSE, col.names=FALSE, sep=",")

load("Z:/Project Melanoom/PaperMelanoma/Data/validation/validation.data.RData")
save(list=c("dd.validation", "validation.data.to.be.imputed", "m", 
            "imputed.validation.data", "last.imputed.validation.data.set", "horizon", 
            "S.Rec.validation", "S.MSM.validation", 
            "S.Rec.5.validation", "S.MSM.5.validation", "f.mi.BS.Rec.5"),
     file=paste0("Z:/Project Melanoom/PaperMelanoma/Data/validation/validation.data.Rdata"), compress=TRUE)
```

# Redevelop model
```{r, eval=TRUE}
form.full.MSM.5.redev <- S.MSM.5.training ~
  SNstatus +
  Sex +
  Age.SN +
  Ulceration +
  # Loc_CAT +
  # Histology +
  log(Breslow) +
  # multiple.fields +
  # Tot_SNs_neg +
  # Tot_SNs_pos +
  log(Rdamcrit) +
  # Dewar +
  # SNstatus:Sex +
  SNstatus:Age.SN +
  # SNstatus:Ulceration +
  # SNstatus:Loc_CAT +
  # SNstatus:Histology +
  SNstatus:log(Breslow) 
  # SNstatus:multiple.fields
f.mi.full.MSM.5.redev <- Hmisc::fit.mult.impute(form.full.MSM.5.redev,cph,
                                                xtrans=imputed.training.data,
                                                data=training.data.to.be.imputed,
                                                n.impute=m,pr=FALSE,fit.reps=TRUE,
                                                fitargs=list(y=TRUE,x=TRUE,se.fit=TRUE))
# summary(f.mi.full.MSM.5.redev)
stats::anova(f.mi.full.MSM.5.redev)

# backward selection
# round(stats::anova(f.mi.full.MSM.5.redev)[order(stats::anova(f.mi.full.MSM.5.redev)[, "P"]),], 3)

# c-index
p.MSM.5.redev<- matrix(NA, nrow(training.data.to.be.imputed), m)
f.MSM.basehaz <- survival::basehaz(cph(form.full.MSM.5.redev, 
                             data=mice::complete(imputed.training.data, 1), 
                             se.fit=TRUE, x=TRUE, y=TRUE))
h0.MSM.redev <- f.MSM.basehaz$hazard[f.MSM.basehaz$time==max(f.MSM.basehaz$time[f.MSM.basehaz$time<=horizon])]
for (i in 1:m){
  lp.val.i <- predict(f.mi.full.MSM.5.redev,
                      newdata=mice::complete(imputed.training.data, i),
                      type="lp")
  p.MSM.5.redev[, i] <- 1-exp(-h0.MSM.redev*exp(lp.val.i))
}
PredictionTools::val.surv.mi(p.MSM.5.redev, S.MSM.training, g=5,
                             main="All patients", time=5, lim=c(0, 1),
                             dist=TRUE)
```

# Calculate baseline hazards
```{r, eval=TRUE}
# baseline hazard of recurrence averaged over all imputed data sets
h0.Rec.all <- c()
for (i in 1:m){
  f.Rec.basehaz.i <- survival::basehaz(cph(form.BS.Rec.5, data=mice::complete(imputed.training.data, i), se.fit=TRUE, x=TRUE, y=TRUE))
  h0.Rec.i <- f.Rec.basehaz.i$hazard[f.Rec.basehaz.i$time==max(f.Rec.basehaz.i$time[f.Rec.basehaz.i$time<=horizon])]
  h0.Rec.all <- c(h0.Rec.all, h0.Rec.i)
}
h0.Rec <- mean(h0.Rec.all)

# baseline hazard of MSM
h0.MSM.all <- c()
for (i in 1:m){
  lp.i<-predict(f.mi.BS.Rec.5,
                newdata=mice::complete(imputed.training.data,i), 
                type="lp")
  f.MSM.basehaz.i <- survival::basehaz(rms::cph(S.MSM.5.training~lp.i,y=TRUE,x=TRUE))
  h0.MSM.i <- f.MSM.basehaz.i$hazard[f.MSM.basehaz.i$time==max(f.MSM.basehaz.i$time[f.MSM.basehaz.i$time<=horizon])]
  h0.MSM.all <- c(h0.MSM.all, h0.MSM.i)
}
h0.MSM <- mean(h0.MSM.all)
```
# Save for webapp
```{r,eval=TRUE}
save(f.mi.BS.Rec.5, MSM.cal.fact, h0.Rec, h0.MSM, c.Breslow,
     file="Z:/Project Melanoom/PaperMelanoma/Results/model.Rdata", 
     compress=TRUE)
save(list=c("dd.training", "training.data.to.be.imputed", "m", 
            "imputed.training.data", "last.imputed.training.data.set", 
            "horizon", "S.Rec.training", "S.MSM.training",
            "S.Rec.5.training", "S.MSM.5.training",
            "f.mi.full.Rec.5", "f.mi.BS.Rec.5", "f.mi.full.MSM.5.refit", 
            "f.mi.BS.MSM.5.refit", "MSM.cal.fact", "c.Breslow", "h0.Rec", "h0.MSM",
            "form.full.Rec.5", "form.BS.Rec.5", "form.full.MSM.5.refit", "form.BS.MSM.5.refit"),
     file="Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.Rdata", compress=TRUE)
```