### 2024-03-11 Code/06_CreateTablesandPlots_save.Rmd

---
title: "Create Tables"
author: "Carolien C.H.M. Maas"
date: "November 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Create Tables

### Load packages, functions, and data 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# erase memory
rm(list = ls(all.names = TRUE))

# for reproducibility
set.seed(100)

# load libraries
library(foreign)
library(mice)
library(survminer)
library(prodlim)
library(rms)
library(Hmisc)
library(png)
# remotes::install_github("CHMMaas/PredictionTools")
library(PredictionTools)
source("Z:/Project Melanoom/PaperMelanoma/Code/PaperMelanomaFunctions.R")
```

### Create variables with results
##### Put hazard ratio"s and anova Chi-square statistics of positive and negative SN patients in Table of final recurrence model and refitted MSM model
```{r, eval=TRUE}
load("Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.RData")
options(datadist="dd.training")

for (model in c("Rec.5", "MSM.5.refit")){
  # load model
  f.mi.full <- eval(parse(text=paste0("f.mi.full.", model)))

  # Coefficients with CI
  # Negative results
  neg.results <- summary(f.mi.full, SNstatus="Negative", Loc_CAT="arm")
  neg.table <- rbind(rep(NA, 3), # SN status
                     exp(neg.results[c("Sex - male:female",
                                       "Age.SN",
                                       "Ulceration - Yes:No",
                                       "Loc_CAT - leg:arm",
                                       "Loc_CAT - trunk:arm",
                                       "Loc_CAT - head & neck:arm",
                                       "Histology - NM:SSM",
                                       "Histology - ALM:SSM",
                                       "Histology - other:SSM",
                                       "Breslow",
                                       "multiple.fields - Multiple SN fields:Other",
                                       "Tot_SNs_neg"), c("Effect", "Lower 0.95", "Upper 0.95")]),
                     matrix(NA, nrow=6, ncol=3)) # Tot_SNs_pos, Rdamcrit, Combined, Parenchymal, Multifocal, Extensive
  # add CI
  neg.table <- cbind(sprintf("%.2f", neg.table[, 1]),
                     paste0("[", sprintf("%.2f", neg.table[, 2]), "; ", 
                            sprintf("%.2f", neg.table[, 3]), "]"))

  # Positive results
  pos.results <- summary(f.mi.full, SNstatus="Positive", Loc_CAT="arm", Tot_SNs_pos=c(0, 1))
  pos.table <- rbind(exp(neg.results["SNstatus - Positive:Negative", c("Effect", "Lower 0.95", "Upper 0.95")]),
                     exp(pos.results[c("Sex - male:female",
                                       "Age.SN",
                                       "Ulceration - Yes:No",
                                       "Loc_CAT - leg:arm",
                                       "Loc_CAT - trunk:arm",
                                       "Loc_CAT - head & neck:arm",
                                       "Histology - NM:SSM",
                                       "Histology - ALM:SSM",
                                       "Histology - other:SSM",
                                       "Breslow",
                                       "multiple.fields - Multiple SN fields:Other",
                                       "Tot_SNs_neg",
                                       "Tot_SNs_pos",
                                       "Rdamcrit",
                                       "Dewar - combined:subcap",
                                       "Dewar - parenchymal:subcap",
                                       "Dewar - multifocal:subcap",
                                       "Dewar - extensive:subcap"), c("Effect", "Lower 0.95", "Upper 0.95")]))
  # add CI
  pos.table <- cbind(sprintf("%.2f", pos.table[, 1]),
                     paste0("[", sprintf("%.2f", pos.table[, 2]), "; ", sprintf("%.2f", pos.table[, 3]), "]"))

  # combine postive with negative
  full.table <- cbind(neg.table, pos.table)

  # add reference rows
  final.table <- rbind(full.table[1:4,],
                       rep(NA, 4),     # location
                       c("(ref)", NA, "(ref)", NA), # ref arm
                       full.table[5:7,],   # other locations
                       rep(NA, 4),     # histology
                       c("(ref)", NA, "(ref)", NA), # ref SMM
                       full.table[8:15,],
                       rep(NA, 4),     # Dewar
                       c(rep(NA, 2), "(ref)", NA), # ref subcap
                       full.table[16:nrow(full.table),]) # other Dewar

  # add Chi-square statistics
  anova.out <- anova(f.mi.full)[, "Chi-Square"]
  anova.main <- c(anova.out[c("SNstatus  (Factor+Higher Order Factors)",
                              "Sex  (Factor+Higher Order Factors)",
                              "Age.SN  (Factor+Higher Order Factors)",
                              "Ulceration  (Factor+Higher Order Factors)",
                              "Loc_CAT  (Factor+Higher Order Factors)")],
                  rep(NA, 4),
                  anova.out["Histology  (Factor+Higher Order Factors)"],
                  rep(NA, 4),
                  anova.out[c("Breslow  (Factor+Higher Order Factors)",
                              "multiple.fields  (Factor+Higher Order Factors)",
                              "Tot_SNs_neg",
                              "Tot_SNs_pos",
                              "Rdamcrit",
                              "Dewar")],
                  rep(NA, 5))
  anova.int <- c(anova.out[c("TOTAL INTERACTION",
                             "SNstatus * Sex  (Factor+Higher Order Factors)",
                             "SNstatus * Age.SN  (Factor+Higher Order Factors)",
                             "SNstatus * Ulceration  (Factor+Higher Order Factors)",
                             "SNstatus * Loc_CAT  (Factor+Higher Order Factors)")],
                 rep(NA, 4),
                 anova.out["SNstatus * Histology  (Factor+Higher Order Factors)"],
                 rep(NA, 4),
                 anova.out[c("SNstatus * Breslow  (Factor+Higher Order Factors)",
                             "SNstatus * multiple.fields  (Factor+Higher Order Factors)")],
                 rep(NA, 9))
  final.table <- cbind(final.table, sprintf("%.0f", anova.main), sprintf("%.0f", anova.int))

  colnames(final.table) <- c("HR.neg", "CI.neg", "HR.pos", "CI.pos", "Main", "Int")
  assign(paste0("final.full.table.", model), as.data.frame(final.table))
}
```

##### Put hazard ratio"s and anova Chi-square statistics of positive and negative SN patients in Table of final recurrence model and refitted MSM model
```{r, eval=TRUE}
for (model in c("Rec.5", "MSM.5.refit")){
  f.mi.BS <- eval(parse(text=paste0("f.mi.BS.", model)))

  # Coefficients with CI
  # Negative results
  neg.results.BS <- summary(f.mi.BS, SNstatus="Negative", Breslow=exp(c.Breslow), Loc_CAT="arm")
  pos.results.BS <- summary(f.mi.BS, SNstatus="Positive", Breslow=exp(c.Breslow), Loc_CAT="arm")
  table.BS <- rbind(exp(neg.results.BS["SNstatus - Positive:Negative", c("Effect", "Lower 0.95", "Upper 0.95")]),
                    exp(neg.results.BS[c("Age.SN",
                                             "Ulceration - Yes:No",
                                             "Loc_CAT - leg:arm",
                                             "Loc_CAT - trunk:arm",
                                             "Loc_CAT - head & neck:arm",
                                             "Breslow"), c("Effect", "Lower 0.95", "Upper 0.95")]),
                    exp(pos.results.BS[c("Breslow",
                                         "Rdamcrit"), c("Effect", "Lower 0.95", "Upper 0.95")]))
  # add CI
  BS.table <- cbind(sprintf("%.2f", table.BS[, 1]),
                        paste0("[", sprintf("%.2f", table.BS[, 2]), "; ", 
                               sprintf("%.2f", table.BS[, 3]), "]"))

  # add reference rows
  BS.table <- rbind(rep(NA, 2),             # SN status title
                    c("(ref)", NA),         # Negative SN status
                    BS.table[1:3,],         # Positive SN status, Age, Ulceration
                    rep(NA, 2),             # Location title
                    c("(ref)", NA),         # Upper limb ref
                    BS.table[4:6,],         # Lower limb, tunk, head&neck
                    rep(NA, 2),             # Breslow title
                    BS.table[7:9,])         # Breslow negative and positive, tumour burden

  # add Chi-square statistics
  anova.out <- anova(f.mi.BS)[, "Chi-Square"]
  anova.main <- c(anova.out[c("SNstatus  (Factor+Higher Order Factors)")], # SN status
                  NA,               # negative SN
                  NA,               # positive SN
                  anova.out[c("Age.SN",     # Age
                              "Ulceration", # Ulceration
                              "Loc_CAT")],  # Location
                  NA,               # upper limb
                  NA,               # lower limb
                  NA,               # trunk
                  NA,               # head & neck
                  NA,               # Breslow title
                  anova.out[c("Breslow  (Factor+Higher Order Factors)")], # Breslow negative
                  anova.out["SNstatus * Breslow  (Factor+Higher Order Factors)"], # Breslow positive
                  anova.out[c("Rdamcrit")])
  BS.table <- cbind(BS.table, sprintf("%.0f", anova.main))

  # add p-values
  p.out <- anova(f.mi.BS)[, "P"]
  p.main <- c(p.out[c("SNstatus  (Factor+Higher Order Factors)")], # SN status
                  NA,               # negative SN
                  NA,               # positive SN
                  p.out[c("Age.SN",     # Age
                              "Ulceration", # Ulceration
                              "Loc_CAT")],  # Location
                  NA,               # upper limb
                  NA,               # lower limb
                  NA,               # trunk
                  NA,               # head & neck
                  NA,               # Breslow title
                  p.out[c("Breslow  (Factor+Higher Order Factors)")], # Breslow negative
                  p.out["SNstatus * Breslow  (Factor+Higher Order Factors)"], # Breslow positive
                  p.out[c("Rdamcrit")])
  BS.table <- cbind(BS.table, sprintf("%.4f", p.main))

  colnames(BS.table) <- c("HR", "CI", "Imp.", "p-value")
  assign(paste0("final.BS.table.", model), as.data.frame(BS.table))
}
```

# Calculate optimism using bootstrap using full model and doing backward selection for each bootstrap
```{r, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(100)
slope<-NULL
cindex.orig<-NULL
opt<-NULL
cindex.B<-NULL
for (i in 1:m){
  v <- validate(f.mi.full.Rec.5$fits[[i]], method="boot", bw=TRUE, rule="p",
                sls=0.05, B=100, pr=FALSE, type="individual")
  slope <- c(slope, v["Slope","test"]) # slope of test
  cindex.orig <- c(cindex.orig, (v["Dxy","index.orig"]+1)/2) # original C-index
  opt <- c(opt, v["Dxy","optimism"]/2) # optimism C-index
  cindex.B <- c(cindex.B, (v["Dxy","index.corrected"]+1)/2) # index corrected C-index
}
```

# Show optimism
```{r, eval=TRUE}
mean(cindex.orig)
optimism <- mean(opt)
optimism
mean(cindex.B)
mean(cindex.orig)-optimism
```

# Add C-index to model for Recurrence and MSM refit Table
```{r, eval=TRUE}
for (model in list("Rec.5", "MSM.5.refit")){
  print(model)
  if (model == "Rec.5"){
    # plot predict
    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/plot.predict.Rec.png"),
        width=1800, height=1800, units="px", res=300)
    plot.predict <- plot(Predict(f.mi.full.Rec.5))
    show(plot.predict)
    grDevices::dev.off()

    # plot predict Breslow
    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/plot.predict.Breslow.Rec.png"),
        width=1800, height=1800, units="px", res=300)
    plot.Breslow <- plot(Predict(f.mi.full.Rec.5, Breslow=0:20, SNstatus), ~ Breslow, groups="SNstatus",
                         adj.subtitle=FALSE, lwd=2, labels=FALSE,
                         par.settings=list(superpose.line=list(lty=c(1, 2), 
                                                               col=c("red", "blue"))),
                         key=list(cex=.8,corner = c(0.1, .9),
                                  lines=list(col=c("red","blue"), 
                                             lty=c(1,2), lwd=2),
                                  text=list(c("Negative","Positive"))))
    show(plot.Breslow)
    grDevices::dev.off()
  }

  # discrimination full and backward selected model
  for (type in c("full", "BS")){
    print(type)
    # load model
    f.mi <- eval(parse(text=paste0("f.mi.", type, ".", model)))
    form.mi <- eval(parse(text=paste0("form.", type, ".", model)))

    # obtain predictions for each imputed data set
    p <- matrix(NA, nrow(training.data.to.be.imputed), m)
    for (i in 1:m){
      # baseline hazard
      # f.Rec.basehaz.i <- basehaz(f.mi, TRUE)
      f.Rec.basehaz.i <- survival::basehaz(cph(form.mi, 
                                               data=mice::complete(imputed.training.data, i), 
                                               se.fit=TRUE, x=TRUE, y=TRUE))
      h0.Rec.i <- f.Rec.basehaz.i$hazard[f.Rec.basehaz.i$time==max(f.Rec.basehaz.i$time[f.Rec.basehaz.i$time<=horizon])]

      # probabilities
      lp.i <- predict(f.mi,newdata=mice::complete(imputed.training.data, i), type="lp")
      p.i <- fun.event(lp=lp.i, h0=h0.Rec.i)
      p[, i] <- p.i
    }

    # load outcome
    if (model=="Rec.5"){
      S <- S.Rec.training
    }
    else if (model=="MSM.5.refit"){
      S <- S.MSM.training
    }

    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/all.centers.", model, ".", type, ".png"),
        width=1800, height=1800, units="px", res=300)
    out <- PredictionTools::val.surv.mi(p=p, y=S, g=5, time=5,
                                        main="All centers", CI.metrics=TRUE,
                                        lim=c(0,1), dist=TRUE, n.iter=40)
    grDevices::dev.off()

    final.full.table <- eval(parse(text=paste0("final.", type, ".table.", model)))
    final.full.table["Harrell's C-index", "HR"] <- sprintf("%.2f", out$HarrellC)
    final.full.table["Harrell's C-index", "CI"] <- paste0("[", sprintf("%.2f", out$HarrellC.lower),
                                                   "; ", sprintf("%.2f", out$HarrellC.upper), "]")
    final.full.table["Harrell's optimism-corrected C-index", "HR"] <- sprintf("%.2f", out$HarrellC-optimism)
    final.full.table["Harrell's optimism-corrected C-index", "CI"] <- paste0("[", sprintf("%.2f", out$HarrellC.lower-optimism),
                                                   "; ", sprintf("%.2f", out$HarrellC.upper-optimism), "]")
    final.full.table["Uno's C-index", "HR"] <- sprintf("%.2f", out$UnoC)
    final.full.table["Uno's C-index", "CI"] <- paste0("[", sprintf("%.2f", out$UnoC.lower),
                                                   "; ", sprintf("%.2f", out$UnoC.upper), "]")
    final.full.table["AUC", "HR"] <- sprintf("%.2f", out$AUC)
    final.full.table["AUC", "CI"] <- paste0("[", sprintf("%.2f", out$AUC.lower),
                                                   "; ", sprintf("%.2f", out$AUC.upper), "]")

    assign(paste0("final.", type, ".table.", model), final.full.table)

    if (type=="BS"&model=="Rec.5"){
      CIndex.BS.Rec.5.All <- c(sprintf("%.2f", out$HarrellC-optimism), 
                              paste0("[", sprintf("%.2f", out$HarrellC.lower-optimism), 
                                     "; ", sprintf("%.2f", out$HarrellC.upper-optimism), "]"))
    }
  }
}
```

# Add the C-index for calibrated MSM model to Table for recurrence
```{r, eval=TRUE}
for (type in c("full", "BS")){
  print(type)
  f.mi <- eval(parse(text=paste0("f.mi.", type, ".Rec.5")))
  HarrellC.MSM <- rep(0, m)
  HarrellC.MSM.se <- rep(0, m)
  UnoC.MSM <- rep(0, m)
  UnoC.MSM.se <- rep(0, m)
  AUC.MSM <- rep(0, m)
  AUC.MSM.se <- rep(0, m)
  for (i in 1:m){
    # linear predictor ith prediction
    lp.i <- predict(f.mi, newdata=mice::complete(imputed.training.data,i), type="lp")
    f.MSM.lp.i <- survival::coxph(S.MSM.5.training ~ lp.i)

    # Harrell's C-index
    rc.H <- survival::concordance(f.MSM.lp.i, timewt = "n")
    HarrellC.MSM[i] <- rc.H$concordance
    HarrellC.MSM.se[i] <- sqrt(rc.H$var)

    # Uno's C-index
    rc.U <- survival::concordance(f.MSM.lp.i, timewt = "n/G2")
    UnoC.MSM[i] <- rc.U$concordance
    UnoC.MSM.se[i] <- sqrt(rc.U$var)

    # time-dependent ROC
    AUC.MSM.i <- timeROC::timeROC(T=S.MSM.training[, 1],
                                   delta=S.MSM.training[, 2],
                                   marker=lp.i,
                                   times=horizon,
                                   cause=1,
                                   iid=TRUE)
    AUC.MSM[i] <- as.numeric(AUC.MSM.i$AUC[2])
    AUC.MSM.CI <- stats::confint(AUC.MSM.i)
    AUC.MSM.se[i] <- as.numeric((as.numeric(AUC.MSM.CI$CI_AUC[2])/100-as.numeric(AUC.MSM.i$AUC[2]))/AUC.MSM.CI$C.alpha)
  }
  HarrellC.mi.MSM <- Rubin.combine(HarrellC.MSM, HarrellC.MSM.se)
  Uno.mi.MSM <- Rubin.combine(UnoC.MSM, UnoC.MSM.se)
  AUC.mi.MSM <- Rubin.combine(AUC.MSM, AUC.MSM.se)

  final.full.table <- eval(parse(text=paste0("final.", type, ".table.Rec.5")))
  final.full.table["Harrell's C-index MSM", "HR"] <- sprintf("%.2f", HarrellC.mi.MSM$est)
  final.full.table["Harrell's C-index MSM", "CI"] <- paste0("[",
    sprintf("%.2f", HarrellC.mi.MSM$est+qnorm(.025)*HarrellC.mi.MSM$se),
    "; ", sprintf("%.2f", HarrellC.mi.MSM$est+qnorm(.975)*HarrellC.mi.MSM$se), "]")
  final.full.table["Harrell's optimism-corrected C-index MSM", "HR"] <- sprintf("%.2f", HarrellC.mi.MSM$est-optimism)
  final.full.table["Harrell's optimism-corrected C-index MSM", "CI"] <- paste0("[",
    sprintf("%.2f", HarrellC.mi.MSM$est+qnorm(.025)*HarrellC.mi.MSM$se-optimism),
    "; ", sprintf("%.2f", HarrellC.mi.MSM$est+qnorm(.975)*HarrellC.mi.MSM$se-optimism), "]")
  final.full.table["Uno's C-index MSM", "HR"] <- sprintf("%.2f", Uno.mi.MSM$est)
  final.full.table["Uno's C-index MSM", "CI"] <- paste0("[",
    sprintf("%.2f", Uno.mi.MSM$est+qnorm(.025)*Uno.mi.MSM$se),
    "; ", sprintf("%.2f", Uno.mi.MSM$est+qnorm(.975)*Uno.mi.MSM$se), "]")
  final.full.table["AUC MSM", "HR"] <- sprintf("%.2f", AUC.mi.MSM$est)
  final.full.table["AUC MSM", "CI"] <- paste0("[",
    sprintf("%.2f", AUC.mi.MSM$est+qnorm(.025)*AUC.mi.MSM$se),
    "; ", sprintf("%.2f", AUC.mi.MSM$est+qnorm(.975)*AUC.mi.MSM$se), "]")

  assign(paste0("final.", type, ".table.Rec.5"), final.full.table)
}
```

# Save dataframes of result Tables
```{r, eval=TRUE}
# replace NA by blank
for (df in c("final.full.table.Rec.5", "final.full.table.MSM.5.refit", 
             "final.BS.table.Rec.5", "final.BS.table.MSM.5.refit")){
  final.df <- eval(parse(text=df))
  final.df[final.df=="[NA; NA]"] <- NA
  final.df[final.df=="NA"] <- NA
  final.df[is.na(final.df)] <- ""
  assign(df, final.df)
}

# set rownames
final.full.table.Rec.5 <- as.data.frame(final.full.table.Rec.5)
row.names(final.full.table.Rec.5) <- c("Positive SN status", "Male",
                                    "Age iqr", "Ulceration", "Location",
                                    "Arm", "Leg", "Trunk", "Head and neck",
                                    "Histology", "SSM", "NM", "ALM", "Other",
                                    "Breslow iqr", "Multiple fields", 
                                    "Total number of negative nodes",
                                    "Total number of positive nodes",
                                    "SN tumour burden iqr", 
                                    "Location metastasis in lymph node",
                                    "Subcap", "Combined", "Parenchymal",
                                    "Multifocal", "Extensive",
                                    rownames(final.full.table.Rec.5)[(nrow(final.full.table.Rec.5)-7):nrow(final.full.table.Rec.5)])
final.full.table.MSM.5.refit <- as.data.frame(final.full.table.MSM.5.refit)
row.names(final.full.table.MSM.5.refit) <- c("Positive SN status", "Male",
                                          "Age", "Ulceration", "Location",
                                          "Arm", "Leg", "Trunk", "Head and neck",
                                          "Histology", "SSM", "NM", "ALM", "Other",
                                          "Breslow", "Multiple fields", 
                                          "Total number of negative nodes",
                                          "Total number of positive nodes",
                                          "SN tumour burden", 
                                          "Location metastasis in lymph node",
                                          "Subcap", "Combined", "Parenchymal",
                                          "Multifocal", "Extensive",
                                          row.names(final.full.table.MSM.5.refit)[(nrow(final.full.table.MSM.5.refit)-3):nrow(final.full.table.MSM.5.refit)])
final.BS.table.Rec.5 <- as.data.frame(final.BS.table.Rec.5)
row.names(final.BS.table.Rec.5) <- c("SN status",
                                     "Negative SN status",
                                     "Positive SN status", 
                                     "Age", 
                                     "Ulceration",
                                     "Location", 
                                     "Upper limb", 
                                     "Lower limb", 
                                     "Trunk", 
                                     "Head and neck",
                                     "Breslow",
                                     "Breslow (negative)",
                                     "Breslow (positive)",
                                     "SN tumour burden", 
                                   row.names(final.BS.table.Rec.5)[(nrow(final.BS.table.Rec.5)-7):nrow(final.BS.table.Rec.5)])
final.BS.table.MSM.5.refit <- as.data.frame(final.BS.table.MSM.5.refit)
row.names(final.BS.table.MSM.5.refit) <- c("SN status",
                                     "Negative SN status",
                                     "Positive SN status", 
                                     "Age", 
                                     "Ulceration",
                                     "Location", 
                                     "Upper limb", 
                                     "Lower limb", 
                                     "Trunk", 
                                     "Head and neck",
                                     "Breslow",
                                     "Breslow (negative)",
                                     "Breslow (positive)",
                                     "SN tumour burden",  
                                   row.names(final.BS.table.MSM.5.refit)[(nrow(final.BS.table.MSM.5.refit)-3):nrow(final.BS.table.MSM.5.refit)])

# save tables
write.table(final.full.table.Rec.5,
            file="Z:/Project Melanoom/PaperMelanoma/Results/final.full.table.Rec.txt", 
            sep=",")
write.table(final.full.table.MSM.5.refit,
            file="Z:/Project Melanoom/PaperMelanoma/Results/final.full.table.MSM.refit.txt", 
            sep=",")
write.table(final.BS.table.Rec.5,
            file="Z:/Project Melanoom/PaperMelanoma/Results/final.BS.table.Rec.txt", 
            sep=",")
write.table(final.BS.table.MSM.5.refit,
            file="Z:/Project Melanoom/PaperMelanoma/Results/final.BS.table.MSM.refit.txt", 
            sep=",")

# show tables
final.full.table.Rec.5
final.full.table.MSM.5.refit
final.BS.table.Rec.5
final.BS.table.MSM.5.refit
```

# C-index of final model for recurrence stratified by SN status
```{r, eval=FALSE}
for (arm in c("Positive", "Negative")){
  f.mi.stratified <- eval(parse(text=paste0("f.mi.BS.Rec.5")))
  S.5.subset <- S.Rec.5.training[last.imputed.training.data.set$SNstatus==arm,]

  cindex.stratified <- rep(0, m)
  cindex.se.stratified <- rep(0, m)
  for (i in 1:m){
    subset.imputed.training.data <- as.data.frame(mice::complete(imputed.training.data, i)[last.imputed.training.data.set$SNstatus==arm,])
    rc <- Hmisc::rcorr.cens(-predict(f.mi.stratified,
                                     newdata=subset.imputed.training.data, 
                                     type="lp"), 
                            S.5.subset)
    cindex.stratified[i] <- rc["C Index"]
    cindex.se.stratified[i] <- rc["S.D."]/2
  }
  cindex.mi.stratified <- Rubin.combine(cindex.stratified, cindex.se.stratified)

  assign(paste0("CIndex.BS.Rec.5.", arm),
         c(sprintf("%.2f", cindex.mi.stratified$est),
           paste0("[", sprintf("%.2f", cindex.mi.stratified$est+qnorm(.025)*cindex.mi.stratified$se),
                  "; ", sprintf("%.2f", cindex.mi.stratified$est+qnorm(.975)*cindex.mi.stratified$se), "]")))
}

# optimism-corrected
CIndex.Rec.Table <- rbind(CIndex.BS.Rec.5.All, CIndex.BS.Rec.5.Positive, CIndex.BS.Rec.5.Negative)
CIndex.Rec.Table
# utils::write.table(CIndex.Rec.Table,
#               file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/CIndex.train.txt"),
#               row.names=FALSE, col.names=FALSE, sep=",")
```

```{r,eval=TRUE}
save(list=c("dd.training", "training.data.to.be.imputed", "m", 
            "imputed.training.data", "last.imputed.training.data.set", 
            "horizon", 
            "S.Rec.training", "S.MSM.training",
            "S.Rec.5.training", "S.MSM.5.training",
            "f.mi.full.Rec.5", "f.mi.BS.Rec.5", "f.mi.full.MSM.5.refit", 
            "f.mi.BS.MSM.5.refit", "MSM.cal.fact", "h0.Rec", "h0.MSM",
            "c.Breslow", "optimism",
            "form.full.Rec.5", "form.BS.Rec.5", 
            "form.full.MSM.5.refit", "form.BS.MSM.5.refit"),
     file="Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.Rdata", compress=TRUE)
```