### 2024-03-11 Code/11_TestModel.Rmd

---
title: "Check model"
author: "Carolien C.H.M. Maas"
date: "November 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Create calibration plots

### Load packages, functions, and data 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# erase memory
rm(list = ls(all.names = TRUE))

# for reproducibility
set.seed(100)

# load libraries
library(rms)
library(survival)
source("Z:/Project Melanoom/PaperMelanoma/Code/PaperMelanomaFunctions.R")
```

### Check mean event rates of risk distribution
#### Recurrence
```{r, eval=TRUE}
load("Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.RData")

# h0 for only last imputed data set
f.Rec.basehaz.5 <- survival::basehaz(f.mi.BS.Rec.5,TRUE)
h0.Rec.5 <- f.Rec.basehaz.5$hazard[f.Rec.basehaz.5$time==max(f.Rec.basehaz.5$time[f.Rec.basehaz.5$time<=horizon])]

# lp for last imputed data set
lp.Rec<-predict(f.mi.BS.Rec.5,
              newdata=mice::complete(imputed.training.data,5), 
              type="lp")

# calibration factor
f.mi.5 <- rms::cph(S.MSM.5.training~lp.Rec,y=TRUE,x=TRUE) 
MSM.cal.fact.5 <- stats::coef(f.mi.5)

# sum(h.Rec$density*fun.event(lp=lp.points.Rec.5, h0=h0.Rec))
# similar to the 5-year survival probability, e.g. kaplan-meier at 5 years, i.e. survfit at 5 years:
SF<-survival::survfit(S.Rec.5.training~1,data=training.data.to.be.imputed)
100*(1-SF$surv[SF$time==max(SF$time[SF$time<=horizon])])
# average probability of recurrence
mean(fun.event(lp=lp.Rec, h0=h0.Rec))
```
# Test if model is correct
### Check calibration
Intercept of recurrence should be the same as when setting covariates to "zero" values minus the center
```{r, eval=TRUE}
load("Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.RData")
options(datadist="dd.training")

# int.Rec only for last imputed data set
int.Rec
lp.BS <- f.mi.BS.Rec.5$linear
f.mi.BS.MSM.5 <- rms::cph(S.MSM.5.training~lp.BS,y=TRUE,x=TRUE)
20*f.mi.BS.Rec.5$coef["Age.SN"]+
  (log(0.1)-c.Breslow)*f.mi.BS.Rec.5$coef["Breslow"]+
  log(0.01)*f.mi.BS.Rec.5$coef["Rdamcrit"]-
  f.mi.BS.Rec.5$center
```

#### MSM
```{r, eval=TRUE}
lp.BS <- f.mi.BS.Rec.5$linear
f.mi.BS.MSM.5 <- rms::cph(S.MSM.5.training~lp.BS,y=TRUE,x=TRUE)
f.MSM.basehaz.5 <- survival::basehaz(f.mi.BS.MSM.5,TRUE)
h0.MSM.5 <- f.MSM.basehaz.5$hazard[f.MSM.basehaz.5$time==max(f.MSM.basehaz.5$time[f.MSM.basehaz.5$time<=horizon])]

# sum(h.Rec$density*fun.event(lp=lp.points.MSM.5, h0=h0.MSM.5))
# equal to the 5-year survival probability, e.g. kaplan-meier at 5 years, i.e. survfit at 5 years:
SF<-survfit(S.MSM.5.training~1,data=training.data.to.be.imputed)
100*(1-SF$surv[SF$time==max(SF$time[SF$time<=horizon])])
# average probability of MSM
mean(fun.event(lp=MSM.cal.fact*lp.Rec, h0=h0.MSM))
```

# Check if predictions of new and old model align for negative SN patients
## New model
```{r, eval=TRUE}
lp.negative <- predict(f.mi.BS.Rec.5,
                       newdata=last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Negative",],
                       type="lp")
new.negativeSN.tool <- 1-exp(-h0.Rec*exp(lp.negative))
```

## Old model with new data
```{r, eval=TRUE}
rscript.negative <- function(breslow, ulceration, location) {
  S0 <- 0.85837329
  lp <- 0.79351989*log(breslow) +
    0.61912477*I(ulceration == 1) +
    0.29594843*I(location == 1) +
    0.42802046*I(location == 2) +
    0.86053044*I(location == 3) - 0.98860941
  S <- 100 - 100*S0^exp(lp)
  return(S)
}
old.negativeSN.tool <- rscript.negative(breslow=last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Negative", "Breslow"],
                                        ulceration=as.numeric(last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Negative", "Ulceration"])-1,
                                        location=as.numeric(last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Negative", "Loc_CAT"])-1)
```

## Compare
```{r, eval=TRUE}
plot(x=new.negativeSN.tool*100, y=old.negativeSN.tool,
     xlim=c(0, 100), ylim=c(0, 100),
     xlab="New predictions", ylab="Old predictions",
     main="Predicted recurrence of patients with negative SN status")
abline(a=0, b=1, col="red")
```

## Refitted old model new data
```{r, eval=TRUE}
refitted.old.negative <- rms::cph(S.Rec.5.training ~ log(Breslow)+Ulceration+Loc_CAT,
                                  data=last.imputed.training.data.set,
                                  y=TRUE, x=TRUE,
                                  subset=last.imputed.training.data.set$SNstatus=="Negative")
refitted.old.negative
f.Rec.basehaz.neg <- basehaz(refitted.old.negative,TRUE)
h0.neg <- f.Rec.basehaz.neg$hazard[f.Rec.basehaz.neg$time==max(f.Rec.basehaz.neg$time[f.Rec.basehaz.neg$time<=horizon])]
refit.old.negativeSN.tool <- fun.event(predict(refitted.old.negative),  h0.neg)
```

## Compare
```{r, eval=TRUE}
plot(x=new.negativeSN.tool*100, y=refit.old.negativeSN.tool*100,
     xlim=c(0, 100), ylim=c(0, 100),
     xlab="New predictions", ylab="Old predictions",
     main="Predicted recurrence of patients with negative SN status")
abline(a=0, b=1, col="red")
```

## Check mean recurrence probabilities
```{r, eval=TRUE}
SF<-survfit(S.Rec.5.training[last.imputed.training.data.set$SNstatus=="Negative"]~1)
100*(1-SF$surv[SF$time==max(SF$time[SF$time<=horizon])])
mean(new.negativeSN.tool)
mean(refit.old.negativeSN.tool)
mean(old.negativeSN.tool)
```

# Check if predictions of new and old model align for positive SN patients
## New model
```{r, eval=TRUE}
lp.positive <- predict(f.mi.BS.Rec.5,
                       newdata=last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Positive",],
                       type="lp")
new.positiveSN.tool <- 1-exp(-h0.Rec*exp(lp.positive))
```

## Old model with new data
```{r, eval=TRUE}
rscript.positive <- function(age, ulceration, tumor_burden, breslow) {
  S0 <- 0.51762362
  lp <- 0.56109945 * log(age) +
    0.23383770 * log(tumor_burden) +
    0.35489235 * log(breslow) +
    0.36562098 * I(ulceration == 1) - 2.709225
  S <- 100 - 100 * S0^exp(lp)
  return(S)
}
old.positiveSN.tool <- rscript.positive(age=last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Positive", "Age.SN"],
                                        ulceration=as.numeric(last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Positive", "Ulceration"])-1,
                                        breslow=last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Positive", "Breslow"],
                                        tumor_burden=as.numeric(last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Positive", "Rdamcrit"]))
```

## Compare
```{r, eval=TRUE}
plot(x=new.positiveSN.tool*100, y=old.positiveSN.tool,
     xlim=c(0, 100), ylim=c(0, 100),
     xlab="New predictions", ylab="Old predictions",
     main="Predicted recurrence of patients with positive SN status")
abline(a=0, b=1, col="red")
```

## Refitted old model new data
```{r, eval=TRUE}
refitted.old.positive <- rms::cph(S.Rec.5.training ~ log(Breslow)+Ulceration+Loc_CAT,
                                  data=last.imputed.training.data.set,
                                  y=TRUE, x=TRUE,
                                  subset=last.imputed.training.data.set$SNstatus=="Positive")
refitted.old.positive
f.Rec.basehaz.pos <- basehaz(refitted.old.positive,TRUE)
h0.pos <- f.Rec.basehaz.pos$hazard[f.Rec.basehaz.pos$time==max(f.Rec.basehaz.pos$time[f.Rec.basehaz.pos$time<=horizon])]
refit.old.positiveSN.tool <- fun.event(predict(refitted.old.positive),  h0.pos)
```

## Compare
```{r, eval=TRUE}
plot(x=new.positiveSN.tool*100, y=refit.old.positiveSN.tool*100,
     xlim=c(0, 100), ylim=c(0, 100),
     xlab="New predictions", ylab="Old predictions",
     main="Predicted recurrence of patients with positive SN status")
abline(a=0, b=1, col="red")
```

## Check mean recurrence probabilities
```{r, eval=TRUE}
SF<-survfit(S.Rec.5.training[last.imputed.training.data.set$SNstatus=="Positive"]~1)
100*(1-SF$surv[SF$time==max(SF$time[SF$time<=horizon])])
mean(new.positiveSN.tool)
mean(refit.old.positiveSN.tool)
mean(old.positiveSN.tool)
```

#### Example to check probabilities with app
```{r, eval=TRUE}
patient.nr <- 115 # 395 and 396 = Negative, 397 and 398 = Positive, 30 = Head & neck, 115 has large Breslow
print(training.data.to.be.imputed[patient.nr, c("SNstatus", "Age.SN", "Ulceration", "Loc_CAT", "Breslow", "Rdamcrit")])

# calculate linear predictor
lp.score.Rec.true <- predict(f.mi.BS.Rec.5)[patient.nr]
lp.score.Rec.try <- predict(f.mi.BS.Rec.5, newdata=data.frame(SNstatus=training.data.to.be.imputed[patient.nr,"SNstatus"],
                                                              Age.SN=training.data.to.be.imputed[patient.nr,"Age.SN"],
                                                              Ulceration=training.data.to.be.imputed[patient.nr,"Ulceration"],
                                                              Loc_CAT=training.data.to.be.imputed[patient.nr,"Loc_CAT"],
                                                              Breslow=training.data.to.be.imputed[patient.nr,"Breslow"],
                                                              Rdamcrit=training.data.to.be.imputed[patient.nr,"Rdamcrit"]),
                            type="lp")
p.Rec.example <- fun.event(lp=lp.score.Rec.try, h0=h0.Rec)
p.MSM.example <- fun.event(lp=lp.score.Rec.try, h0=h0.MSM)

# manual calculation
lp.score.Rec.manual <- coef.Rec["SNstatus=Positive"]*(as.numeric(training.data.to.be.imputed[patient.nr,"SNstatus"])-1)+
  coef.Rec["Age.SN"]*training.data.to.be.imputed[patient.nr,"Age.SN"]+
  coef.Rec["Ulceration=Yes"]*(as.numeric(training.data.to.be.imputed[patient.nr,"Ulceration"])-1)+
  coef.Rec["Loc_CAT=leg"]*as.numeric(training.data.to.be.imputed[patient.nr,"Loc_CAT"]=="leg")+
  coef.Rec["Loc_CAT=trunk"]*as.numeric(training.data.to.be.imputed[patient.nr,"Loc_CAT"]=="trunk")+
  coef.Rec["Loc_CAT=head & neck"]*as.numeric(training.data.to.be.imputed[patient.nr,"Loc_CAT"]=="head & neck")+
  coef.Rec["Breslow"]*(log(training.data.to.be.imputed[patient.nr,"Breslow"])-c.Breslow)+
  coef.Rec["Rdamcrit"]*log(training.data.to.be.imputed[patient.nr,"Rdamcrit"])+
  coef.Rec["SNstatus=Positive * Breslow"]*(log(training.data.to.be.imputed[patient.nr,"Breslow"])-c.Breslow)*(as.numeric(training.data.to.be.imputed[patient.nr,"SNstatus"])-1)

# lp check
lp.score.Rec.true
lp.score.Rec.try
lp.score.Rec.manual-center.Rec

# probability of recurrence with fun.event
fun.event(lp=lp.score.Rec.try, h0=h0.Rec)
# probability of recurrence manually
1-exp(-h0.Rec*exp(lp.score.Rec.manual-center.Rec))

# check calculation of lp of MSM
lp.score.MSM.true <- predict(f.mi.BS.MSM.5)[patient.nr]
lp.score.MSM.true
lp.MSM.try <- MSM.cal.fact*(lp.score.Rec.manual-center.Rec)
lp.MSM.try

# probability of MSM with fun.event
fun.event(lp=lp.MSM.try, h0=h0.MSM)
# probability of MSM manually
1-exp(-h0.MSM*exp(MSM.cal.fact*(lp.score.Rec.manual-center.Rec)))
```

### Histogram for negative SN status patients
```{r, eval=TRUE}
Breslow.trunc.neg <- last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Negative", "Breslow"]
Breslow.trunc.neg[last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Negative", "Breslow"]<0.1] <- .1
Breslow.trunc.neg[last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Negative", "Breslow"]>7] <- 7
lp.score.Rec.negative <- predict(f.mi.BS.Rec.5,newdata=data.frame(SNstatus=last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Negative", "SNstatus"],
                                                                  Age.SN=last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Negative", "Age.SN"],
                                                                  Ulceration=last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Negative", "Ulceration"],
                                                                  Loc_CAT=last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Negative", "Loc_CAT"],
                                                                  Breslow=Breslow.trunc.neg,
                                                                  Rdamcrit=last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Negative", "Rdamcrit"]))
mean(lp.score.Rec.negative)
# OK to use int.Rec and rc.Rec based on all patients?
score.Rec.neg <- round((lp.score.Rec.negative-int.Rec)/rc.Rec,0)
h.Rec.neg <- hist(score.Rec.neg,plot=FALSE,breaks=0:max.breaks,right=FALSE)
h.plot.neg <- h.Rec.neg
h.plot.neg$density <- 100*h.Rec.neg$density
```

### Histogram for positive SN status patients
```{r, eval=TRUE}
Breslow.trunc.pos <- last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Positive", "Breslow"]
Breslow.trunc.pos[last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Positive", "Breslow"]<0.1] <- .1
Breslow.trunc.pos[last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Positive", "Breslow"]>7] <- 7
lp.score.Rec.positive <- predict(f.mi.BS.Rec.5,newdata=data.frame(SNstatus=last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Positive", "SNstatus"],
                                                                  Age.SN=last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Positive", "Age.SN"],
                                                                  Ulceration=last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Positive", "Ulceration"],
                                                                  Loc_CAT=last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Positive", "Loc_CAT"],
                                                                  Breslow=Breslow.trunc.pos,
                                                                  Rdamcrit=last.imputed.training.data.set[last.imputed.training.data.set$SNstatus=="Positive", "Rdamcrit"]))
mean(lp.score.Rec.positive)
# OK to use int.Rec and rc.Rec based on all patients?
score.Rec.pos <- round((lp.score.Rec.positive-int.Rec)/rc.Rec,0)
h.Rec.pos <- hist(score.Rec.pos,plot=FALSE,breaks=0:max.breaks,right=FALSE)
h.plot.pos <- h.Rec.pos
h.plot.pos$density <- 100*h.Rec.pos$density

par(mfrow=c(1, 3))
plot(h.plot,freq=FALSE,axes=TRUE, xlab="Risk score", xlim=x.lim,ylim=y.lim.hist,
     ylab="Risk score distribution (%)", main="All patients",col="white")
plot(h.plot.neg,freq=FALSE,axes=TRUE, xlab="Risk score", xlim=x.lim,
     ylab="Risk score distribution (%)",ylim=y.lim.hist,main="Negative SN status",col="white")
plot(h.plot.pos,freq=FALSE,axes=TRUE, xlab="Risk score", xlim=x.lim,
     ylab="Risk score distribution (%)",ylim=y.lim.hist,main="Positive SN status",col="white")
```

### Test proportional hazards assumption
```{r, eval=TRUE}
test.prop<-survival::cox.zph(f.mi.full.Rec.5$fits[[1]])
test.prop
plot(test.prop)
```
### See if RFS always lower than MSS
```{r, eval=TRUE}
p.RFS <- fun.event(lp=predict(f.mi.BS.Rec.5), h0=h0.Rec)
p.MSS <- fun.event(lp=MSM.cal.fact*predict(f.mi.BS.Rec.5), h0=h0.MSM)
plot(p.RFS, p.MSS)
abline(a=0, b=1, col="red")
```